### 2017-06-07 abstract.tex
Initial ShareLaTeX Import

\begin{abstract}

Three dimensional electron microscopy (3DEM) is becoming a very data-intensive field in which vast amounts of experimental data are acquired at high speed. To manage such large-scale image data  projects, we have previously described  a modular workflow system called Scipion. We present here \scipionbox, an extension to Scipion for processing data in streaming. Streaming processing allows to overlap data acquisition with the first steps of digital image processing and therefore helps to: (1) detect problems at early stages, (2) save computing time and (3) provide users with a first evaluation of the data quality before the data acquisition is finished. This idea of processing the data while being acquired has been implemented in different laboratories mainly using custom-made scripts. The advantages of our \scipionbox solution are: (1) flexibility: users can customize their workflows from a  large collection of preloaded algorithms; (2) repeatability and traceability: workflows are stored and can be re-executed later, either with the same data or a different data set; (3) monitorization: there are special utilities that constantly check how the execution of the algorithms is going and inform about possible anomalies; (4) reports: we have developed several GUIs that are updated periodically and produce a graphical summary (e.g, CTF defocus values, system load etc). This summary may be  generated in HTML format that can be easily published though a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided. At present, \scipionbox has been installed and is in production mode in three facilities. And it is being considered in another two.

\textbf{Keywords:} electron microscopy

\end{abstract}

### 2017-06-07 acknowledgments.tex
Initial ShareLaTeX Import

\section{Acknowledgments}

UPDATE THIS acknowledgments

Javier Vargas is the recipient of a ‘‘Comfuturo” grant funded by
the Fundación General CSIC, C.O.S. Sorzano is the recipient of a
Ramón y Cajal fellowship from the Spanish Ministry of Economy
and Competitiveness (MINECO). The authors acknowledge financial
support from the MINECO through Grants AIC-A-2011-0638,
BFU2013-41249-P and BIO2013-44647-R, the Comunidad de
Madrid through Grant CAM (S2010/BMD-2305), the European
Union (EU) and Horizon 2020 through Grants EINFRA-2014-2:
EGI Engage (Proposal: 654142) and EINFRA-2015-1: West-Life
(Proposal: 675858).
This work was partially funded by Instruct, a component of the
European Strategy Forum on Research Infrastructures (ESFRI), and
supported by national member subscriptions.

    AFTER THIS LINE I JUST SAVED SOME TEXT THAT MAY BE USEFUL IN THE FUTURE DISREGARD THIS MATERIAL
   --------------------

\section{Getting Started}

\section{Project stability and funding. getting started}

Scipion and ScipionBox development is been done in the context of the Instruct Image Processing Cen-
ter (I2PC, http://i2pc.cnb.csic.es). Instruct is the European initiative for Strategic
Scientific Infrastructures (ESFRI) for Structural Biology. The I2PC started with the objective 
to provide efficient support for research projects in Europe demanding expertise in 
electron microscopy image processing. In order to support a large collection
of projects, the first activities of the I2PC were oriented to design and implement the
necessary software tools.

At present the I2PC has this money and this people

ScipionBox is part of the Scipion package which is provided 
freely as open source software. Online documentation
describing Scipion download and installation is available
at \url{http://scipion.cnb.csic.es/m/download_form/} and \url{https://github.com/I2PC/scipion/wiki/How-to-Install}, respectively.

%It was recognized the need for a platform that could provide
%a friendly GUI on top of advanced image processing algorithms and, at the same time,
%assure standardization, traceability and reproducibility.
\section{Things to mention if possible}
\begin{itemize}
 \item Cloud
\end{itemize}

### 2017-06-07 introduction.tex
Initial ShareLaTeX Import

\section{Introduction}

3D electron microscopy (3DEM)  can provide rich information about structural characteristics of macromolecular complexes. The field is under a major transformation due to the arrival of better microscopes [**REF], new large area detectors [**REFS ddd] and automatization [**REFs to data colection software as EPU, leginon, serialEm,etc...]  which makes possible that a single microscope generates high quality data sets on the order of terabytes per day [**REF] while working for several days without interruption. Major challenges faced in the field are: (1) an efficient management of these huge datasets and their corresponding image processing workflows; (2) a raising interest from scientists outside the field who might lack the skills of experienced microscopist and; (3) an the need for an effective transference of data and associated metadata in an increasingly distributed and collaborative environment. This situation places techniques that are automated and allow higher throughput in high demand.

We will focus in this article in the critical stage of the image data acquisition in 3DEM.  Our goal is  to develop software (\scipionbox) able to extend the functionality of the microscope data collecting programs. In this way, \scipionbox starts the processing of the movies as they are being acquired with the double aim of checking for possible data collection errors and execute a first user-tailored image processing workflow that provides users with an accurate measurement of the acquired data quality.

Accounting for major government investment in increasingly costly electron microscopes requires a large user  base. Clearly, the future of projects that require this high-end microscopes is through centralized microscopy facilities, in which automatic acquired data needs to be monitored by users and facility staff. \scipionbox has been designed with this scenario in mind in which the early detection of issues may save the data acquisition shift which otherwise need to be repeated after a new application that may be allowed many months later.

Many software packages have been developed to provide automated data collection, such as SerialEM (Mastronarde, 2005), Leginon (Suloway et al., 2005), UCSF-Tomography (Zheng et al., 2007), Tom Toolbox (Nickell et al.2005), EPU (Thermo Fisher Co.), Latitude in Digital Micrograph (Gatan Co.), or EMMenu
(TVIPS, Germany), etc. However, the next step,  which is to provide image processing while  data are being collected, is not so well served. In fact, in most places this is accomplished through home made scripts. Exceptions are UCSFImage4 [**REF] and Focus (Biayi et al 2017). As we will explain here, \scipionbox differs from these solutions in a larger flexibility both from the point of view of users and facility staff.  [**this parragraph need to be improved]

### 2017-06-07 overall.tex
Initial ShareLaTeX Import

\section{Overall System Design: Worflows, Protocols, Monitors, Consensus and Beyond}
\label{overall}
Just like snowflakes, no two \scipionbox installation are alike. So far, the system is running in 
three centralized facilities, and each installation has been tailored to the particular needs of each facility. The original design satisfied the requirements of our home facility (National Center for Biotechnology, Madrid, Spain), in which \scipionbox automatically fetches newly recorded movies from a network mounted disk. After this first installation, \scipionbox was deployed on XXXXX (a centralized facility located at Stockholm [**provide full identification @delarosa]) where a rigorous booking system is followed. There, \scipionbox was adapted to interface with this booking platform in order to produce better reports. Lately,  \scipionbox has been installed in a large Synchrotron  such as Diamond (Oxford, UK) where projects are handled by ISPyB (a customized lims -laboratory management information system- [**REF]) and data are saved on a distributed file system that heavily penalizes access to lists of files. In this environment, \scipionbox needs to constantly interchange information with ISPyB, and it cannot longer check if there is a new file in a particular storage device but must wait to be informed by the system that a new file has been created. All these particularities, that will be described in detail later, could only be handled thanks to a careful design that prioritizes flexibility at many levels.

\scipionbox uses the infrastructure provided by the Scipion project [Ref if not given already]
that takes care of the creation and execution of workflows. These workflows may be imported from a set of predefined ones or created by the final users even if they have no programing skills.

In most cases, after importing data, the first step will be to correct for beam-induced movement. For this task, \scipionbox allows you to choose between five different algorithms from three different labs (an exhaustive list containing the approximately 150 algorithms available in \scipionbox is at appendix XX). This multiplicity of possibilities is available at each different processing step although, if found overwhelming, may be ignored by choosing one of the default prepacked workflows.

In addition to offering algorithms (in Scipion lingo ``protocols'') we provide 
a set of ``monitors'', which are protocols that  check on the progress of other protocols. The monitors are designed to produce analysis plots, generate reports or raise alerts when some problems are detected. A monitor example is the CTF monitor, that checks the computed defocus values for
each micrograph and can raise an alert if the values are above or below certain thresholds.
Monitors were also designed in a modular way to be modified or extended. 
Monitors may send alert emails, continuously generate html files or interact with lims so that users and staff may easily follow the data acquisition and processing in-house or remotely. At the project end, a review of the processing history is produced by default.
%This feature was very important since we knew that difference facilities will have different needs and criteria to be fulfilled.

Another set of  specialized protocols are grouped under the name ``consensus''. For a given logical step (for example, particle picking) these protocols check if the datasets obtained from the same input data using different algorithm are identical. Continuing with the particle picking example, ``consensus picking'' will compare the particles selected for each one of the executed algorithms and report on the particles selected by most of them.

\scipionbox is a streaming system that runs on threads. This is a significant difference from a script, since it means that each module is run independently. For example, Do you want to add a new protocol or monitor in your workflow while it is running? No problem, do it. \scipionbox allows workflows to be  modified on the fly. The modification will not interfere with the already executed steps and may take advantage from all already computed data. Did something go wrong and the microscope stopped? By default \scipionbox will stop after 2 hours with no microscope activity, but if you wish to continue your acquisition after a stop just press the continue button [*+ scipion 1.1.2], there is no need to restart from the beginning.

Last but not least Scipion and therefore \scipionbox provides full data provenance. Data provenance is an important form of metadata that describes how a particular data set was generated, by detailing the steps, input data and parameters used in the computational process producing it. Data provenance guaranties reproducibility and traceability. Since provenance is solved by Scipion and not Scipionbox we will not discuss here how it is been achieved. We just note that, before choosing which data has to be stored, it is necessary to define how these data have to be structured so that they can be later recovered and understood. This abstract modeling of the data also simplifies the software interoperability offered by Scipion.

%Standardization
%User has access to all parameters used by algorithms (she may ignore them)
%Project can be imported in scipion without further ado.
%Gui vs programatico
%Export options?

### 2017-06-07 results.tex
Initial ShareLaTeX Import

\section{Results: Speed and performance}

%Our objective is to build a real time high performance pipeline for image processing  in electron microscopy that involves a continual input, process and output of data. 
Reliable software is a prerequisite for successful operation of a modern electron microscope facility but it is not longer enough since data processing speed is not a negligible aspect of the acquisition pipeline. Nowadays, image processing  in electron microscopy involves a continual input, process and output of data in a time period similar to the acquisition rate.

In order to understand the limitations of the streaming model for data processing we have tested and compared the speed performance of the different steps involved in the data acquisition and processing. The measured data provides values for guiding the design and implementation of new workflows and protocols.

\subsection{ TODO LIST:}

Prerequisites:
\begin{itemize}
   \item create protocol that mimic set of movie acquisition
   \item modify protocol system so it reports on IO
\end{itemize}

\noindent
TODO (requires microscope time):
\begin{itemize}
   \item Check samba maximum speed between microscope PC and scipion box (IDEA: use DiskSpd o similar with no load in the computers, https://blogs.technet.microsoft.com/josebda/2014/10/13/diskspd-powershell-and-storage-performance-measuring-iops-throughput-and-latency-for-both-local-disks-and-smb-file-shares/)

   \item Repeat above measurement with the  microscope working: a) alone, b) doing backup to local external usb disk , c) doing backup to local disk and another user downloading files, d) remote backup plus second user downloading files

   \item in all the following the backup must be running (at least at CNB)

   \item Motioncor execution (1 vs 2-3-4 threads). In one GPU and in 2 if possible

   \item Now add CTFfind - gCTF (comment on ultra fast programs as gctf)

   \item Add particle picking and extract if available

   \item All the time we need CPU, memory, GPU and IO plots

   \item Running things up to extract  how much available CPU, GPU time do we have?

   \item Can we say something about using ssd disk versus plain sata

\end{itemize}

### 2017-06-07 scipionbox.tex
Initial ShareLaTeX Import

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage[T1]{fontenc}
\usepackage[table]{xcolor} % Colors and tables
\usepackage{color, colortbl} % Define new colors and use in tables
\usepackage{setspace} % Set paragraph lines space
%\usepackage{tabu} % Other table package
\usepackage{url} % Other table package
\usepackage{listings}
\usepackage{rotating}
\usepackage[strict]{changepage}
\usepackage[caption=false]{subfig}
\usepackage{fancyhdr}
\usepackage{rotating}
\usepackage{xr}
\usepackage{xspace}
%\externaldocument[Supp-]{supplementary_material}

%\usepackage{lineno}
\usepackage[update,prepend]{epstopdf}
%\linenumbers
\definecolor{lightgreen}{rgb}{0.75,0.75,0.0}
%\usepackage{caption}
%\usepackage{endfloat}
%\DeclareCaptionFormat{empty}{}
%\captionsetup{format=empty}
\pagestyle{fancy}
\fancyhf{} % clear all header and footer fields
\fancyfoot[R]{\footnotesize  \thepage}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\def\scipionbox{\textit{ScipionBox}\xspace}
\def\scipion{\textit{Scipion}\xspace}
\newcommand{\protocol}[1]{\texttt{#1}}

%%%%%%%%%%%%%%%%%\def\baselinestretch{2}
%\def\PARstart#1#2{#1#2} % if draft, disable it
\def\degree{\ensuremath{^\circ}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\newcommand{\lyxaddress}[1]{
   \par {\raggedright #1
   \vspace{1.4em}
   \noindent\par}
}

\newcommand{\itemph}[1]{\item{\emph{#1}}}
\newcommand{\itprog}[1]{\item{\texttt{#1}}:}
\newcommand{\refig}[1]{Fig. \ref{#1}}
\newcommand{\seefig}[1]{(see \refig{#1})}
\definecolor{grey}{rgb}{0.9,0.9,0.9}
\renewcommand{\arraystretch}{1.3}

\lstset{ %
  language=XML,                % the language of the code
  basicstyle=\footnotesize{8pt},       % the size of the fonts that are used for the code
  %numbers=left,                   % where to put the line-numbers
  %numberstyle=\tiny\color{gray},  % the style that is used for the line-numbers
  %stepnumber=2,                   % the step between two line-numbers. If it's 1, each line 
                                  % will be numbered
  %numbersep=5pt,                  % how far the line-numbers are from the code
  backgroundcolor=\color{grey},      % choose the background color. You must add \usepackage{color}
  %showspaces=false,               % show spaces adding particular underscores
  %showstringspaces=false,         % underline spaces within strings
  %showtabs=false,                 % show tabs within strings adding particular underscores
  frame=single,                   % adds a frame around the code
  %rulecolor=\color{black},        % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. commens (green here))
  %tabsize=2,                      % sets default tabsize to 2 spaces
  %captionpos=b,                   % sets the caption-position to bottom
  %breaklines=true,                % sets automatic line breaking
  %breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
  %title=\lstname,                   % show the filename of files included with \lstinputlisting;
                                  % also try caption instead of title
  keywordstyle=\color{blue},          % keyword style
  %commentstyle=\color{dkgreen},       % comment style
  %stringstyle=\color{mauve},         % string literal style
  %escapeinside={\%*}{*)},            % if you want to add a comment within your code
  %morekeywords={*,...}               % if you want to add more keywords to the set
}

\begin{document}
\title{ScipionBox: automatic processing for }
\author{XXXXX XXXXX$^{(a)}$, 
        YYYYY YYYYY$^{(b)}$, \\
        zzzzzz$^{(m)}$,
        yyyyyyy$^{(b)}$
        }

%xxxx$^{(1)}$, xxxxx$^{(1,3)}$}
\maketitle

\lyxaddress{{\small \vspace{-0.5cm}$^{(a)}$Escuela Polit\'{e}cnica Superior, Universidad Aut\'{o}noma de Madrid, 28049 Cantoblanco, Madrid, Spain.}}
\lyxaddress{{\small$^{(b)}$ The National Resource for Automated Molecular Microscopy, The Scripps Research Institute, La Jolla, CA 92037, USA}}
\lyxaddress{{\small$^{(c)}$ Baylor College of Medicine, Houston, Texas 77030, USA}}
\lyxaddress{{\small$^{(d)}$ Lawrence Berkeley National Laboratory, Berkeley, CA 94720, USA}}
\lyxaddress{{\small$^{(e)}$ Howard Hughes Medical Institute, Columbia University, NY 10032, USA}}
\lyxaddress{{\small$^{(f)}$ Purdue University, Biological Sciences, IN 47907-2054, USA}}
\lyxaddress{{\small$^{(g)}$ Biocomputing Unit, National Center for Biotechnology (CSIC), C/ Darwin, 3, Campus Universidad Aut\'{o}noma, 28049 Cantoblanco,
Madrid, Spain.}}
\lyxaddress{{\small$^{(h)}$ Biozentrum, University of Basel, CH - 4058 Basel, Switzerland}}
\lyxaddress{{\small$^{(i)}$ MRC-LMB, Cambridge CB2 0QH, UK. 01223 267000, United Kingdom}}
\lyxaddress{{\small$^{(j)}$IMPMC, Sorbonne Universit\'{e}s - CNRS UMR 7590, UPMC Univ Paris 6, MNHN, IRD UMR 206, 75005 Paris, France.}} 
\lyxaddress{{\small$^{(k)}$Roosevelt University, Department of Biological, Chemical, and Physical Sciences, 1400 N. Roosevelt Blvd., Schaumburg, IL 60173, USA.}}
\lyxaddress{{\small$^{(l)}$Laboratory of Structural Biology Research, National Institute of Arthritis, Musculoskeletal and Skin Diseases,
National Institutes of Health, Bethesda, MD 20892, USA}}
\lyxaddress{{\small$^{(m)}$Massachusetts Institute of Technology, USA}}
%\lyxaddress{{\small$^{(14)}$Massachusetts Institute of Technology, USA}}
% 
% \lyxaddress{{\small \vspace{-0.5cm}$^{(3)}$Bioengineering Lab., Universidad CEU San Pablo, Campus Urb.
% Montepríncipe s/n, 28668 Boadilla del Monte, Madrid, Spain.}}
%

%This document provides a preliminary analysis of the data collected by the CTF Challenge (http://i2pc.cnb.csic.es/3dembenchmark/LoadHome.htm). Figures/tables/plots/etc are located at the document's end. 
\setcounter{figure}{2}%I do not know why but figure counter starts with -1
\textbf{Corresponding Author:} %Roberto Marabini, Ph= 34 91 5854510, email: roberto@cnb.csic.es\\
%Escuela Polit\'{e}cnica Superior\\
%Universidad Aut\'{o}noma de Madrid\\
%28049 Cantoblanco\\
%Madrid\\
%Spain

\newpage
\input{abstract.tex}
\input{introduction.tex}
\input{overall.tex}
\input{workflow.tex}
\input{results.tex}
\input{conclusions.tex}
\input{acknowledgments.tex}
\appendix

\bibliographystyle{elsart-harv}
\bibliography{large}

\clearpage
%\input{listOfFigures.tex}

\newpage
%tables
%\input{numberImagesDataSet.tex}
%%%%%\input{markersLegend.tex}
%\input{numberImagesUpload.tex}
%\newpage

\end{document}

### 2017-06-07 workflow.tex
Initial ShareLaTeX Import

\section{Processing with ScipionBox}

\scipionbox is a very flexible framework that allows to create many different workflows and to choose among several algorithms at each workflow-step. In this work we do not recommend a particular workflow 
over others since we realize that different specimens, microscopes or facilities may have different requirements. Nevertheless, since different use cases may clarify \scipionbox capabilities, in the following we describe the pipeline offered by default to the users of three different facilities where \scipionbox has been installed.

\subsection{CNB Facility}

CNB setup is a quite straight forward one. Data sets are recorded using EPU which runs in a dedicated Windows box. Data are not locally saved but they are stored in a Linux server through a Samba share. The Linux server has two Ethernet connections, a 10 Gb point-to-point link with the Windows box and a 1 Gb link with the local network. It also has 32 CPUs at 2.40GHz, 62 Gb of RAM memory and 1 \textit{Quadro M4000} GPU. \scipionbox is installed in the Linux server.

The first decision a user must face is the backup policy. The recommended strategy is either to attach an external hard disk to the server or provide a remote storage location. \scipionbox will transfer both raw and processed data as soon as available. In this way, access to the Linux server seldom overlaps between two users and once a microscope shift ends, users do not need to wait for their data to be transferred.

Once the image processing pipeline begins \scipionbox will start two independent protocols in parallel. The first protocol estimates the camera gain at each pixel for a few  movies ([**REF coss article][**ROB TODO: this protocol is not in streaming]). In the CNB setup the microscope camera is calibrated and internally performs the gain correction before writing the movies so the estimated gain image should be almost constant and equal to 1. The output of this protocol serves as a verification of the validity of the experimental gain image for the whole data set. The second protocol performs Scipion direct detector movie alignment. Currently \scipionbox offers  four  methods  for  DDD  movie  alignment: motioncorr (Li et al., 2013),
correlation (an unpublished CPU version of the motioncorr method developed in Xmipp),
unblur (Grant and Grigorieff, 2015) and optical flow (Abrishami et al., 2015). Within \scipionbox, movie frames can be aligned using any one of these approaches or any combination. Following the standard processing workflow, the resulting micrographs are screened to eliminate those with excessive astigmatism or drift. Scipion provides three CTF estimation methods:
CTFFIND (Rohou and Grigorieff, 2015), gCTF (Zhang, 2016) and Xmipp CTF estimation (Sorzano et al., 2007; Vargas et al., 2013). At this point the user may compute the CTFs using several algorithms and compare them with the protocol \protocol{CTF consensus} that report up to which frequency two estimation of the same micrograph are equivalent. Further automatic filtering based on CTF is available through the protocol \protocol{CTF filtering} that selects micrographs based on the astigmatism, resolution and defocus. Another criteria for CTF selection is iciness, that is,  the ratio of the intensity in the resolution band between 3.5 and 3.9 \AA, and between 80 and 30 \AA, which gives a good estimate of the ice crystal content of the vitreous specimen.

[**ROB TODO: I have not added the average protocol since I do not know exactly what it does, if somebody wants to contribute here go ahead]

Up to this point the image processing pipeline is quite solid and may be executed in streaming without user intervention. Now, it is time to select (pick) particles. We have recently ported the eight different options available at \scipion  to streaming (see XXXX for details[**delarosa up to which point this is true]). In this way,  particles may be selected as soon as any new micrograph is available. We notice here that some of the ported algorithms require a training step in which either user intervention or information from a previous selection is need. For those users that execute several picking algorithms, the protocol \protocol{picking consensus} may be useful. \protocol{picking consensus}  calculates the agreement between different particle-picking algorithms and produces a list with the particle selected by most of them. [**ROB this algorithm is not in streaming yet, volunteers?]

Two new protocols are under development and will join the described  pipeline in a near future. The first one will perform 2D classification and the second one will create a initial 3D map based on this classes. \scipion currently implements these protocols but these implementations are not adequate for real time processing.

As commented in section \ref{overall}, to complement the streaming process we have developed a set of monitors and report possibilities. By default \protocol{system}, \protocol{CTF} and \protocol{report} monitors are executed. \protocol{System} monitor reports on  CPU and GPU utilization, memory consumption and disk I/O [**ROB TODO: last one to be done]. \protocol{CTF} monitor plots the CTF defocus and astigmatism  and it will raise an alarm if these parameters are not within a given range. Finally the \protocol{report} protocol pushes all the created plots and summaries to a web server 
so this information can be accessed from outside.

The proposed pipeline offers quite flexibility to decide which particular algorithms will be used by a given user. At CNB (and collaborating institutions) we collect protocol usage statistics and sort the different algorithms by popularity.  This information is available at  \url{http://calm-shelf-73264.herokuapp.com/report_protocols/protocolTable/} and may help users to select   between the different algorithms.

\subsection{XX Facility}

\subsection{YY Facility}

%cat /proc/cpuinfo
%\begin{figure}[h]
%\centering
%     \includegraphics[width=50mm]{Images/to-be-done}
%     \caption{TODO: DROP? I do not know if an image showing CNB setup is needed this image should show CNB setup, microscope connected to control-PC which saves data in \scipionbox using samba. Show Ethernet speed. Soy connection to external disk/host.
%     Describe number GPUs, CPUs, SSD disks.
%     }
%     \label{cnbsetup}
%\end{figure}

%One of the most frequent requests by both beginners, and even experienced digital darkroom workers, is for detailed step-by-step instructions on how to perform various basic tasks in Photoshop.

%One of the key goals of ScipionBox is to guarantee interoperability between algorithms implemented in different software packages.  The possibility of choosing among several algorithms is provided at all levels of the workflow, in many cases adding the possibility of comparing the results of these methods to detect inconsistencies.  
%and do not need to be concerned with how the rest of the workflow is allocated.

%ScipionBox design is agnostic to a particular workflow or software packet.

%%Export options?

\subsection{What will do in future}
%Particle picking
%Extract particle
%2D classify (either classes or identify outliers)
%Initial model
%More complex reports
[** ROB This is already commented so this section disappears ]

### 2017-06-07 introduction.tex
small changes previous to first public release

@@
 The field is under a major transformation due to the arrival of better microscopes [**REF], new large area detectors [**REFS ddd] and automatization [**REFs to data colection software as EPU, leginon, serialEm,etc...]  which makes possible that a single microscope generates high quality data sets on the order of terabytes per day [**REF] while working for several days without interruption.
-Major challenges faced in the field are: (1) an efficient management of these huge datasets and their corresponding image processing workflows; (2) a raising interest from scientists outside the field who might lack the skills of experienced microscopist and; (3) an the need for an effective transference of data and associated metadata in an increasingly distributed and collaborative environment.
+Major challenges faced in the field are: (1) an efficient management of these huge datasets and their corresponding image processing workflows; (2) a raising interest from scientists outside the field who might lack the skills of experienced microscopist and; (3) the need for an effective transference of data and associated metadata in an increasingly distributed and collaborative environment.
 This situation places techniques that are automated and allow higher throughput in high demand.
@@
 As we will explain here, \scipionbox differs from these solutions in a larger flexibility both from the point of view of users and facility staff.
-[**this parragraph need to be improved]

### 2017-06-07 2.overall.tex
Updates from ShareLaTeX

--- 2017-06-07 overall.tex
@@
 In most cases, after importing data, the first step will be to correct for beam-induced movement.
-For this task, \scipionbox allows you to choose between five different algorithms from three different labs (an exhaustive list containing the approximately 150 algorithms available in \scipionbox is at appendix XX).
+For this task, \scipionbox allows you to choose between five different algorithms from three different labs (an exhaustive list containing the approximately 150 algorithms available at [**ROB \url{XXXX}]).
 This multiplicity of possibilities is available at each different processing step although, if found overwhelming, may be ignored by choosing one of the default prepacked workflows.
@@
 Another set of  specialized protocols are grouped under the name ``consensus''.
-For a given logical step (for example, particle picking) these protocols check if the datasets obtained from the same input data using different algorithm are identical.
+For a given logical step (for example, particle picking) these protocols check if the datasets obtained from the same input data using different algorithm are consistent.
 Continuing with the particle picking example, ``consensus picking'' will compare the particles selected for each one of the executed algorithms and report on the particles selected by most of them.
@@
 No problem, do it.
-\scipionbox allows workflows to be  modified on the fly.
+\scipionbox allows users to modify  workflows on the fly without modifiying a single code line.
 The modification will not interfere with the already executed steps and may take advantage from all already computed data.

### 2017-06-16 results.tex
after speaking with pablo
before start meassure results

@@
 \begin{itemize}
-   \item Check samba maximum speed between microscope PC and scipion box (IDEA: use DiskSpd o similar with no load in the computers, https://blogs.technet.microsoft.com/josebda/2014/10/13/diskspd-powershell-and-storage-performance-measuring-iops-throughput-and-latency-for-both-local-disks-and-smb-file-shares/)
+   \item Check samba maximum speed between microscope PC and scipion box. We will use iperf since there is version for windows and linux.
+   \begin{lstlisting}
+        1) check network speed
+        install  (may be iptraf also)
+        Server -> iperf -s -p 80
+        Client -> iperf -c server.cnb.csic.es -p 80
+
+        2) meassure samba
+        install Diskspd
+        https://unhandled.wordpress.com/2016/07/20/madness-testing-smb-direct-network-throughput-with-diskspd/
+        http://stealthpuppy.com/replicate-2015-folder-redirection-test/
+        http://darklight.pro/using-microsoft-diskspd-to-test-starwind-virtual-san-storage-performance/
+        (Compare local and remote storage)
+        diskspd.exe -c2G -d120 -w100 -t1 (-o8) -b8K -h -L \\mb-win81\diskspd\testfile.dat
+        -c -> file size
+        -d -> time at least 120 (600 better)
+        -w -> 100 write, 0 read
+        -t -> threadas per file (1 or 2)
+        -L -> capture latency information
+        -h -> disable hardware, software cache 
+        -b -> size of a read/write operation
+        -o 	Outstanding I/Os (queue depth) per worker thread. Setting this higher increases the intensity of the test, which pushes your storage harder.??
+    \end{lstlisting}
+
+   \begin{lstlisting}
+#smb.conf
+   read raw = Yes
+write raw = Yes
+socket options = TCP_NODELAY IPTOS_LOWDELAY SO_RCVBUF=131072 SO_SNDBUF=131072
+min receivefile size = 16384
+use sendfile = true
+aio read size = 16384
+aio write size = 16384
+    \end{lstlisting}
+\begin{verbatim}
+         OREALLY:
+     TCP_NODELAY
+Have the server send as many packets as necessary to keep delay low. This is
+used on telnet connections to give good response time, and is used—some-
+what counter-intuitively—to get good speed even when doing small requests
+or  when  acknowledgments  are  delayed  (as  seems  to  occur  with  Microsoft
+TCP/IP).  This  is  worth  a  30–50  percent  speedup  by  itself.  Incidentally,  in
+Samba 2.0.4,
+socket options = TCP_NODELAY
+became the default value for
+that option.
+IPTOS_LOWDELAY
+This is another option that trades off throughput for lower delay, but which
+affects  routers  and  other  systems,  not  the  server.  All  the  IPTOS  options  are
+new; they’re not supported by all operating systems and routers. If they are
+supported, set
+IPTOS_LOWDELAY
+ whenever you set
+TCP_NODELAY
+.
+SO_SNDBUF and SO_RCVBUF
+The send and receive buffers can often be the reset to a value higher than that
+of  the  operating  system.  This  yields  a  marginal  increase  of  speed  (until  it
+reaches a point of diminishing returns).
+
+read raw and write raw
+These are important performance configuration options; they enable Samba to use
+large reads and writes to the network, of up to 64KB in a single SMB request. They
+also  require  the  largest  SMB  packet  structures,
+
+\end{verbatim}
 
@@
 
+   \item Report on project size
+
+   \item what happens if I double size
+
+   \item save raw data instead of corrected by the gain
+
+   \item how to measure scipion overhead?????
+
+   \item Laura backup in amazon and ctffind4
+   \begin{lstlisting}
+   rsync -aHAXxv --numeric-ids --delete --progress -e "ssh -T -c arcfour -o Compression=no -x" [source_dir] [dest_host:/dest_dir]
+   \end{lstlisting}
+   \begin{itemize}
+    \item some preliminary tests with Laura
+    \item no matter what machine we choose the network speed with CNB is 250 Mbits/s
+    \item rsync report 25MB/seconds so there is not need to play here
+    \item disk is ssd
+    \item OS: Ubuntu 14.04, 8 CPUs 32GB, m4.2xlarge
+    rsync -av --progress -e "ssh -i ~/Documents/CNB/Cloud/AWS/rmarabini.pem" VMImages/ScipionCloud-v1.1-beta.ova ubuntu@54.154.136.191:
+
+#Comando iperf en server
+
+sudo iperf -s -p 80
+
+------------------------------------------------------------
+Server listening on TCP port 80
+TCP window size: 85.3 KByte (default)
+------------------------------------------------------------
+[  4] local 172.31.37.239 port 80 connected with 150.244.85.94 port 54796
+[ ID] Interval       Transfer     Bandwidth
+[  4]  0.0-10.0 sec   282 MBytes   236 Mbits/sec
+[  5] local 172.31.37.239 port 80 connected with 150.244.85.94 port 54832
+[  5]  0.0-10.1 sec   297 MBytes   247 Mbits/sec
+[  4] local 172.31.37.239 port 80 connected with 150.244.85.94 port 55290
+[  4]  0.0-40.1 sec  1.18 GBytes   253 Mbits/sec
+
+#Comando iperf en client (copiar durante 40 sec)
+
+iperf -c 54.154.136.191 -p 80 -t 40
+
    \end{itemize}
+
+\end{itemize}

### 2017-07-24 scipionbox.tex
Updates from ShareLaTeX
before start meassure results
previous to facility meeting

@@
 \usepackage{xspace}
+\usepackage{listings}
+
 %\externaldocument[Supp-]{supplementary_material}
@@
 
-\lyxaddress{{\small \vspace{-0.5cm}$^{(a)}$Escuela Polit\'{e}cnica Superior, Universidad Aut\'{o}noma de Madrid, 28049 Cantoblanco, Madrid, Spain.}}
-\lyxaddress{{\small$^{(b)}$ The National Resource for Automated Molecular Microscopy, The Scripps Research Institute, La Jolla, CA 92037, USA}}
-\lyxaddress{{\small$^{(c)}$ Baylor College of Medicine, Houston, Texas 77030, USA}}
-\lyxaddress{{\small$^{(d)}$ Lawrence Berkeley National Laboratory, Berkeley, CA 94720, USA}}
-\lyxaddress{{\small$^{(e)}$ Howard Hughes Medical Institute, Columbia University, NY 10032, USA}}
-\lyxaddress{{\small$^{(f)}$ Purdue University, Biological Sciences, IN 47907-2054, USA}}
-\lyxaddress{{\small$^{(g)}$ Biocomputing Unit, National Center for Biotechnology (CSIC), C/ Darwin, 3, Campus Universidad Aut\'{o}noma, 28049 Cantoblanco,
-Madrid, Spain.}}
-\lyxaddress{{\small$^{(h)}$ Biozentrum, University of Basel, CH - 4058 Basel, Switzerland}}
-\lyxaddress{{\small$^{(i)}$ MRC-LMB, Cambridge CB2 0QH, UK. 01223 267000, United Kingdom}}
-\lyxaddress{{\small$^{(j)}$IMPMC, Sorbonne Universit\'{e}s - CNRS UMR 7590, UPMC Univ Paris 6, MNHN, IRD UMR 206, 75005 Paris, France.}} 
-\lyxaddress{{\small$^{(k)}$Roosevelt University, Department of Biological, Chemical, and Physical Sciences, 1400 N. Roosevelt Blvd., Schaumburg, IL 60173, USA.}}
-\lyxaddress{{\small$^{(l)}$Laboratory of Structural Biology Research, National Institute of Arthritis, Musculoskeletal and Skin Diseases,
-National Institutes of Health, Bethesda, MD 20892, USA}}
-\lyxaddress{{\small$^{(m)}$Massachusetts Institute of Technology, USA}}
-%\lyxaddress{{\small$^{(14)}$Massachusetts Institute of Technology, USA}}
-% 
-% \lyxaddress{{\small \vspace{-0.5cm}$^{(3)}$Bioengineering Lab., Universidad CEU San Pablo, Campus Urb.
-% Montepríncipe s/n, 28668 Boadilla del Monte, Madrid, Spain.}}
-%
+%\lyxaddress{{\small \vspace{-0.5cm}$^{(a)}$Escuela Polit\'{e}cnica Superior, Universidad Aut\'{o}noma de Madrid, 28049 Cantoblanco, Madrid, Spain.}}
+%\lyxaddress{{\small$^{(g)}$ Biocomputing Unit, National Center for Biotechnology (CSIC), C/ Darwin, 3, Campus Universidad Aut\'{o}noma, 28049 Cantoblanco,
+%Madrid, Spain.}}
 

### 2017-07-24 1.introduction.tex

--- 2017-06-07 introduction.tex
@@
 The field is under a major transformation due to the arrival of better microscopes [**REF], new large area detectors [**REFS ddd] and automatization [**REFs to data colection software as EPU, leginon, serialEm,etc...]  which makes possible that a single microscope generates high quality data sets on the order of terabytes per day [**REF] while working for several days without interruption.
-Major challenges faced in the field are: (1) an efficient management of these huge datasets and their corresponding image processing workflows; (2) a raising interest from scientists outside the field who might lack the skills of experienced microscopist and; (3) the need for an effective transference of data and associated metadata in an increasingly distributed and collaborative environment.
+Major challenges faced in the field are: (1) an efficient management of these huge datasets and their corresponding image processing workflows; (2) a raising interest from scientists outside the field who might lack the skills of a experienced microscopist and; (3) the need for an effective transference of data and associated metadata in an increasingly distributed and collaborative environment.
 This situation places techniques that are automated and allow higher throughput in high demand.
@@
 Clearly, the future of projects that require this high-end microscopes is through centralized microscopy facilities, in which automatic acquired data needs to be monitored by users and facility staff.
-\scipionbox has been designed with this scenario in mind in which the early detection of issues may save the data acquisition shift which otherwise need to be repeated after a new application that may be allowed many months later.
+\scipionbox has been designed with this scenario in mind in which the early detection of issues may save the data acquisition shift which otherwise need to be repeated after a new application that may be granted many months later.
 

### 2017-07-24 2.overall.tex

@@
 There, \scipionbox was adapted to interface with this booking platform in order to produce better reports.
-Lately,  \scipionbox has been installed in a large Synchrotron  such as Diamond (Oxford, UK) where projects are handled by ISPyB (a customized lims -laboratory management information system- [**REF]) and data are saved on a distributed file system that heavily penalizes access to lists of files.
+Lately,  \scipionbox has been installed in a large Synchrotron  such as Diamond (Oxford, UK) where projects are handled by ISPyB (a customized laboratory management information system -LIMBS- [**REF]) and data are saved on a distributed file system that heavily penalizes access to lists of files.
 In this environment, \scipionbox needs to constantly interchange information with ISPyB, and it cannot longer check if there is a new file in a particular storage device but must wait to be informed by the system that a new file has been created.
@@
 Monitors were also designed in a modular way to be modified or extended. 
-Monitors may send alert emails, continuously generate html files or interact with lims so that users and staff may easily follow the data acquisition and processing in-house or remotely.
+Monitors may send alert emails, continuously generate HTML files or interact with LIMS so that users and staff may easily follow the data acquisition and processing in-house or remotely.
 At the project end, a review of the processing history is produced by default.
@@
 No problem, do it.
-\scipionbox allows users to modify  workflows on the fly without modifiying a single code line.
+\scipionbox allows users to modify  workflows on the fly without modifying a single code line.
 The modification will not interfere with the already executed steps and may take advantage from all already computed data.

### 2017-07-24 3.workflow.tex

--- 2017-06-07 workflow.tex
@@
 \section{Processing with ScipionBox}
+
+%Comment change in scale in the three examples: CNB, SNC and DIAMOND
+%more local, SNC national, Diamond international. Portal infrastructure has
+%different levels.
 
@@
 \subsection{CNB Facility}
+%Here there are Scipion general features and CNB specific. Move the general to the Processing with Scipion section.
 
@@
 The first one will perform 2D classification and the second one will create a initial 3D map based on this classes.
-\scipion currently implements these protocols but these implementations are not adequate for real time processing.
+\scipion currently implements such protocols but the implementation is not adequate for real time processing.
 
@@
 
-\subsection{XX Facility}
+\subsection{Swedish National CryoEM Facility}
 
-\subsection{YY Facility}
+\begin{verbatim}
+We need to cover here differences with CNB setup
+ CNB is a stand alone center. It does not work as a real National Facility
+ SNC already had an application portal and Scipion needs to interact
+ with it. 
+ - initial script
+ - same as CNB until session end
+ - when acquisistion done contaCT PORTAL AND UPLOAD REPORT
+\end{verbatim}
 
+\subsection{Diamond Light Source Facility??? or eBIC???}
+
+\begin{verbatim}
+
+We need to cover here differences with CNB setup
+
+ DEVELOPER RELATED
+ * Describe infrastructure in diamond: file system (distributed, user permissions),
+                                       no graphical output, queues, 
+ * Describe Scipion modification:      import (detect new files by socket message)
+                                       monitor, report to ISPyB 
+                                       scipion headless
+                                       execution as script
+
+USER RELATED
+ * how users interact with scipion?
+ * how data is visualized?
+\end{verbatim}
+
+%%%Diamond Light Source Facility  ends here%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %cat /proc/cpuinfo
@@
 
-%One of the most frequent requests by both beginners, and even experienced digital darkroom workers, is for detailed step-by-step instructions on how to perform various basic tasks in Photoshop.
-
-%One of the key goals of ScipionBox is to guarantee interoperability between algorithms implemented in different software packages.
-The possibility of choosing among several algorithms is provided at all levels of the workflow, in many cases adding the possibility of comparing the results of these methods to detect inconsistencies.
-%and do not need to be concerned with how the rest of the workflow is allocated.
-
-%ScipionBox design is agnostic to a particular workflow or software packet.
-
-%%Export options?
-
 \subsection{What will do in future}
@@
 %More complex reports
-[** ROB This is already commented so this section disappears ]
+[** this section may disappear]

### 2017-07-24 5.conclusions.tex

\section{Discussion}
[*+TODO]: Discuss the obtained numbers in the result section

\section{Conclusions}

The new generation of high-end cryo electron microscopes and direct electron detectors are behind the recent increase in cryo-EM  popularity. However, the run and maintenance costs of these new microscopes are high and can only be justified keeping the microscopes up and running 24 hours a day. In this way, microscopes tent to belong to centralize facilities where users are allocated microscope time (shifts) to perform a variety of research. Available shifts are a scarce commodity and therefore users (and staff) need to be able to monitor the quality of the acquisition process and get as soon as possible a first glimpse of the reconstructed specimen in order to guaranty an early detection of any issue.

 Despite workflow systems have been explored and used extensively in bioinformatics, there is little progress in that direction in the cryo-EM community. \scipionbox has been designed following and strict workflow paradigm. This approach is specially useful for tasks that need to go from a given start (movies) to a defined end (3D map) but there are many different paths (algorithms) to get there.

 In a workflow approach, users (including microscope staff) can reasonably be expected to edit the paths. They can potentially play with the pipeline in ways that aren't possible when having a programmer write code for them. Furthermore, no having the processing path hardcoded in the program means you don't have to recompile, so changes can be quickly tested   even at run time.

 \scipionbox workflow approach provides provenance (transparency) since for any created object (micrograph, CTFs, class, etc) it is stored which program creates it, using which parameters, and the exact path followed to create it.  In this way,  real tracking and reproducibility are provided.

\scipionbox is part of the \scipion package which is provided 
freely as open source software. Online documentation
describing \scipion download and installation is available
at \url{http://scipion.cnb.csic.es/m/download_form/} and \url{https://github.com/I2PC/scipion/wiki/How-to-Install}, respectively.\\\\

% to automatically process user data within a few minutes of data collection at a rate of a few terabytes per day.

%In order to extract useful results from the image data, images must be processed using software which is complex, often lacking intuitive user interfaces and documentation

### 2018-02-13 0.abstract.tex
W

\begin{abstract}

Three dimensional electron microscopy (3DEM) is becoming a very data-intensive field in which vast amounts of experimental data are acquired at high speed. To manage such large-scale image data  projects, we have previously described  a modular workflow system called \scipion. We present here the first main extension of \scipion that allows to process data in streaming. Streaming processing allows to overlap data acquisition with the first steps of digital image processing and therefore helps to: (1) detect problems at early stages, (2) save computing time and (3) provide users with a in deep evaluation of the data quality before the data acquisition is finished. This idea of processing the data while being acquired has been implemented in different laboratories mainly using custom-made scripts. The advantages of our \scipion solution are: (1) flexibility: users can customize their workflows from a  large collection of preloaded algorithms; (2) repeatability and traceability: workflows are stored and can be re-executed later, either with the same data or a different data set; (3) monitorization: there are special utilities that constantly check how the execution of the algorithms is going and inform about possible anomalies; (4) reports: we have developed several GUIs that are updated periodically and produce a graphical summary (e.g, CTF defocus values, system load, etc). This summary may be  generated in HTML format that can be easily published though a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided. At present, \scipion has been installed and is in production mode in three facilities. And it is being considered in several more.

\textbf{Keywords:} electron microscopy

\end{abstract}

### 2018-02-13 1.introduction.tex
W

@@
 3D electron microscopy (3DEM)  can provide rich information about structural characteristics of macromolecular complexes.
-The field is under a major transformation due to the arrival of better microscopes [**REF], new large area detectors [**REFS ddd] and automatization [**REFs to data colection software as EPU, leginon, serialEm,etc...]  which makes possible that a single microscope generates high quality data sets on the order of terabytes per day [**REF] while working for several days without interruption.
+The field is under a major transformation due to the arrival of better microscopes, new large area detectors and automatization \citep{kuhlbrandt2014a, Kuhlbrandt2014b}  which makes possible that a single microscope generates high quality data sets on the order of terabytes per day \citep{Saibil2015} while working for several days without interruption.
 Major challenges faced in the field are: (1) an efficient management of these huge datasets and their corresponding image processing workflows; (2) a raising interest from scientists outside the field who might lack the skills of a experienced microscopist and; (3) the need for an effective transference of data and associated metadata in an increasingly distributed and collaborative environment.
@@
 
-We will focus in this article in the critical stage of the image data acquisition in 3DEM.
-Our goal is  to develop software (\scipionbox) able to extend the functionality of the microscope data collecting programs.
-In this way, \scipionbox starts the processing of the movies as they are being acquired with the double aim of checking for possible data collection errors and execute a first user-tailored image processing workflow that provides users with an accurate measurement of the acquired data quality.
+%We will focus in this article in the critical stage of the image data acquisition in 3DEM.
+Our goal is  to further develop  (\scipion) such that it extends the functionality of the microscope data collecting programs.
+In this way, \scipion starts the processing of the movies as they are being acquired with the double aim of checking for possible data collection errors and execute a first user-tailored image processing workflow that provides users with an accurate measurement of the acquired data quality.
 
-Accounting for major government investment in increasingly costly electron microscopes requires a large user  base.
-Clearly, the future of projects that require this high-end microscopes is through centralized microscopy facilities, in which automatic acquired data needs to be monitored by users and facility staff.
-\scipionbox has been designed with this scenario in mind in which the early detection of issues may save the data acquisition shift which otherwise need to be repeated after a new application that may be granted many months later.
+%Accounting for major government investment in increasingly costly electron microscopes requires a large user  base.
 
-Many software packages have been developed to provide automated data collection, such as SerialEM (Mastronarde, 2005), Leginon (Suloway et al., 2005), UCSF-Tomography (Zheng et al., 2007), Tom Toolbox (Nickell et al.2005), EPU (Thermo Fisher Co.), Latitude in Digital Micrograph (Gatan Co.), or EMMenu
-(TVIPS, Germany), etc.
+Clearly, the future of projects that require high-end microscopes is through centralized microscopy facilities, in which automatic acquired data needs to be monitored by users and facility staff.
+\scipion has been redesigned with this scenario in mind in which the early detection of issues may save a data acquisition shift which otherwise needs to be repeated after a new application that may be granted many months later.
+In this way, \scipion starts the processing of the movies as they are being acquired with the double aim of checking for possible data collection errors and execute a first user-tailored image processing workflow that provides users with an accurate measurement of the acquired data quality.
+
+Many software packages have been developed to provide automated data collection, such as SerialEM \citep{Mastronarde2005}, Leginon \citep{Suloway2009}, UCSF-Tomography \citep{Zheng2007}, Tom Toolbox \citep{Nickell2005}, EPU \citep{EPU}, Latitude in Digital Micrograph \citep{Latitude}, or EMMenu \citep{emmenu}, etc.
 However, the next step,  which is to provide image processing while  data are being collected, is not so well served.
 In fact, in most places this is accomplished through home made scripts.
-Exceptions are UCSFImage4 [**REF] and Focus (Biayi et al 2017).
-As we will explain here, \scipionbox differs from these solutions in a larger flexibility both from the point of view of users and facility staff.
+Exceptions are UCSFImage4 \citep{Li2015} and Focus \citep{Biyani2017}.
+As we will explain here, \scipion differs from these solutions in a larger flexibility both from the point of view of users and facility staff.

### 2018-02-14 2.overall.tex
w

@@
-\section{Overall System Design: Worflows, Protocols, Monitors, Consensus and Beyond}
+\section{Overall System Design: Monitors, Consensus and Streaming}
 \label{overall}
-Just like snowflakes, no two \scipionbox installation are alike. So far, the system is running in 
-three centralized facilities, and each installation has been tailored to the particular needs of each facility.
-The original design satisfied the requirements of our home facility (National Center for Biotechnology, Madrid, Spain), in which \scipionbox automatically fetches newly recorded movies from a network mounted disk.
-After this first installation, \scipionbox was deployed on XXXXX (a centralized facility located at Stockholm [**provide full identification @delarosa]) where a rigorous booking system is followed.
-There, \scipionbox was adapted to interface with this booking platform in order to produce better reports.
-Lately,  \scipionbox has been installed in a large Synchrotron  such as Diamond (Oxford, UK) where projects are handled by ISPyB (a customized laboratory management information system -LIMBS- [**REF]) and data are saved on a distributed file system that heavily penalizes access to lists of files.
-In this environment, \scipionbox needs to constantly interchange information with ISPyB, and it cannot longer check if there is a new file in a particular storage device but must wait to be informed by the system that a new file has been created.
+
+In the following we describe the main modifications introduced in the original \scipion version in order to facilitate its deployment in centralized facilities.
+So far, \scipion is running in three of such facilities, and each installation has been tailored to the particular needs of each facility.
+The original design satisfied the requirements of our home facility (National Center for Biotechnology, Madrid, Spain), in which \scipion automatically fetches newly recorded movies from a network mounted disk.
+After this first installation, \scipion was deployed on XXXXX (a centralized facility located at Stockholm [**provide full identification @delarosa]) where a rigorous booking system is followed.
+There, \scipion was adapted to interface with this booking platform in order to produce better reports.
+Lately,  \scipion has been installed in a large synchrotron  such as Diamond (Oxford, UK) where projects are handled by ISPyB (a customized laboratory management information system -LIMBS- \citep{Delageniere2011}) and data are saved on a distributed file system that heavily penalizes access to lists of files.
+In this environment, \scipion needs to constantly interchange information with ISPyB, and it cannot longer check if there is a new file in a particular storage device but must wait to be informed by the system that a new file has been created.
 All these particularities, that will be described in detail later, could only be handled thanks to a careful design that prioritizes flexibility at many levels.
 
-\scipionbox uses the infrastructure provided by the Scipion project [Ref if not given already]
-that takes care of the creation and execution of workflows. These workflows may be imported from a set of predefined ones or created by the final users even if they have no programing skills.
+%\scipionbox uses the infrastructure provided by the Scipion project [Ref if not given already] that takes care of the creation and execution of workflows.
+These workflows may be imported from a set of predefined ones or created by the final users even if they have no programing skills.
 
-In most cases, after importing data, the first step will be to correct for beam-induced movement.
-For this task, \scipionbox allows you to choose between five different algorithms from three different labs (an exhaustive list containing the approximately 150 algorithms available at [**ROB \url{XXXX}]).
-This multiplicity of possibilities is available at each different processing step although, if found overwhelming, may be ignored by choosing one of the default prepacked workflows.
+%A typical \scipion workflow starts importing the microscope data, then the  beam-induced movement is corrected.
+For this task, \scipion allows you to choose between five different algorithms from three different labs (an exhaustive list containing the approximately 200 algorithms accessible through \scipion and ranked by usability is available at \url{http://calm-shelf-73264.herokuapp.com/report_protocols/protocolTable/}).
+This multiplicity of possibilities is available at each different processing step although, if found overwhelming, may be ignored by choosing one of the default prepackaged workflows.
 
-In addition to offering algorithms (in Scipion lingo ``protocols'') we provide 
-a set of ``monitors'', which are protocols that  check on the progress of other protocols.
-The monitors are designed to produce analysis plots, generate reports or raise alerts when some problems are detected.
-A monitor example is the CTF monitor, that checks the computed defocus values for
-each micrograph and can raise an alert if the values are above or below certain thresholds.
-Monitors were also designed in a modular way to be modified or extended. 
-Monitors may send alert emails, continuously generate HTML files or interact with LIMS so that users and staff may easily follow the data acquisition and processing in-house or remotely.
-At the project end, a review of the processing history is produced by default.
+Since \scipion original publication we have developed many \emph{metaprotocols}, that is, 
+protocols that either check the progress of other protocols or compare the results of equivalent protocols.
+
+The first type of metaprotocols are called \emph{monitors} and are used to produce live analysis plots, generate reports or raise alerts when some problems are detected.
+A monitor example is the \emph{CTF-monitor}, that checks the computed defocus values for
+each micrograph as they are generated.
+CTF-monitor may raise an alert if the defocus values are above or below certain thresholds and continuously generate HTML files so that users and staff may easily follow the data acquisition and processing in-house or remotely.
+At the project's end, a review of the processing history is produced.
+[** URL to report example?]
 %This feature was very important since we knew that difference facilities will have different needs and criteria to be fulfilled.
 
-Another set of  specialized protocols are grouped under the name ``consensus''.
+The second set of metaprotocols are grouped under the name \emph{consensus}.
 For a given logical step (for example, particle picking) these protocols check if the datasets obtained from the same input data using different algorithm are consistent.
-Continuing with the particle picking example, ``consensus picking'' will compare the particles selected for each one of the executed algorithms and report on the particles selected by most of them.
+Continuing with the particle picking example, \emph{consensus picking} will compare the particles selected for each one of the executed particle-picking algorithms and report on the particles selected by most of them.
 
-\scipionbox is a streaming system that runs on threads.
-This is a significant difference from a script, since it means that each module is run independently.
+The main change introduced in \scipion is \emph{stream processing}. Stream processing has been designed to analyze and react on real-time so that enables analysis of data as it is produced.
+In this way a new level of parallelization has been reached since the  data generated by the microscope continuously flows through a chain of protocols.
+For example, as movies are imported to the system they are aligned and their CTF is estimated.
+In \scipion stable version users may find  the first steps in the image processing adapted to stream.
+These steps include: movie alignment and CTF estimation.
+In \scipion developers version all particle picking algorithm was been adapted for Streaming as well as a few 2D classification algorithms.
+
+\scipion stream implementation is based on periodic updates of the protocol input and output data during the image processing.
+Since \scipion runs on threads, it is possible to devote one thread to update these information while the rest of the CPU power is devoted to run the different protocols.
+The use of threads marks an important difference between running a script or using \scipion, since it means that each module is run independently.
 For example, Do you want to add a new protocol or monitor in your workflow while it is running?
 No problem, do it.
-\scipionbox allows users to modify  workflows on the fly without modifying a single code line.
+\scipion allows users to modify  workflows on the fly without modifying a single code line.
 The modification will not interfere with the already executed steps and may take advantage from all already computed data.
 Did something go wrong and the microscope stopped?
-By default \scipionbox will stop after 2 hours with no microscope activity, but if you wish to continue your acquisition after a stop just press the continue button [*+ scipion 1.1.2], there is no need to restart from the beginning.
+By default \scipion will stop after 24 hours with no microscope activity, but if you wish to continue your acquisition after a stop just press the continue button, there is no need to restart from the beginning.
 
-Last but not least Scipion and therefore \scipionbox provides full data provenance.
-Data provenance is an important form of metadata that describes how a particular data set was generated, by detailing the steps, input data and parameters used in the computational process producing it.
+Last but not least \scipion provides full data provenance.
+Data provenance is an important form of metadata that describes how a particular data set was generated by detailing: the steps, input data and parameters used in the computation.
 Data provenance guaranties reproducibility and traceability.
-Since provenance is solved by Scipion and not Scipionbox we will not discuss here how it is been achieved.
-We just note that, before choosing which data has to be stored, it is necessary to define how these data have to be structured so that they can be later recovered and understood.
-This abstract modeling of the data also simplifies the software interoperability offered by Scipion.
+%Since provenance is solved by Scipion and not Scipionbox we will not discuss here how it is been achieved.
+We just note that, before choosing which data has to be stored, it is necessary to define how these data have to be structured so that they can be later recovered and understood .
+%This abstract modeling of the data also simplifies the software interoperability offered by Scipion.
 

### 2018-02-14 3.workflow.tex
w

@@
 
-\scipionbox is a very flexible framework that allows to create many different workflows and to choose among several algorithms at each workflow-step.
+\scipion is a very flexible framework that allows to create many different workflows and to choose among several algorithms at each workflow-step.
 In this work we do not recommend a particular workflow
 over others since we realize that different specimens, microscopes or facilities may have different requirements.
-Nevertheless, since different use cases may clarify \scipionbox capabilities, in the following we describe the pipeline offered by default to the users of three different facilities where \scipionbox has been installed.
+Nevertheless, since different use cases may clarify \scipionbox capabilities, in the following we describe the pipeline offered by default to the users of three different facilities where \scipion has been installed.
 
@@
 CNB setup is a quite straight forward one.
-Data sets are recorded using EPU which runs in a dedicated Windows box.
-Data are not locally saved but they are stored in a Linux server through a Samba share.
-The Linux server has two Ethernet connections, a 10 Gb point-to-point link with the Windows box and a 1 Gb link with the local network.
-It also has 32 CPUs at 2.40GHz, 62 Gb of RAM memory and 1 \textit{Quadro M4000} GPU.
-\scipionbox is installed in the Linux server.
+From a network point of view there are three different levels of accessibility: (1) a private network that cannot be accessed by regular users, (2) the CNB LAN (local area network)  that can be accessed by any user which works inside the hosting  institution and (3) finally, there is a single computer that may be accessed from outside the institution and therefore, it can see the WAN (wide area network).
+
+Data sets are recorded using EPU.
+The microscope control system has a 70 TB NAS (Network-attached storage) unit for saving movies as they are obtained (we will refer to this NAS as microscope-NAS).
+This storage unit is shared through a Samba share and connected to the private network using a 10Gb connexion.
+\scipion is executed on a Linux server which has two 1Gb NICs (network interface controllers) that connect it to the private and the LAN networks (we will refer to this machine as \scipionbox).
+This machine has 32 CPUs at 2.40GHz, 62 Gb of RAM memory and 2 \textit{Quadro M4000} GPUs.
+Finally, a second NAS with 4 1Gb NICs is linked to the private, LAN and WAN networks (we will refer to this second NAS as output-NAS).
+
+The data flow is a follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox will temporary read the movies from 
+the microscope-NAS but will not store them locally, (c) each hour, output-NAS will connect to \scipionbox and copy or update the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup the \scipion project and the movies to a USB disk or serve this information through the Internet and (e) in parallel \scipionbox transfers to  output-NAS different html reports that may be checked by users or staff with a HTML browser.
+[** create figure]
+
+Discuss transfer rates
+Discuss typical workflow
 

### 2018-02-14 2.overall.tex
w

@@
 In \scipion stable version users may find  the first steps in the image processing adapted to stream.
-These steps include: movie alignment and CTF estimation.
-In \scipion developers version all particle picking algorithm was been adapted for Streaming as well as a few 2D classification algorithms.
+These steps include: movie alignment, CTF estimation and particle picking.
+In \scipion developers version all several 2D classification algorithms algorithms are available.
 

### 2018-02-15 3.workflow.tex
split section 3 in three centers

--- 2017-06-07 workflow.tex
@@
 
-\scipionbox is a very flexible framework that allows to create many different workflows and to choose among several algorithms at each workflow-step.
+%Comment change in scale in the three examples: CNB, SNC and DIAMOND
+%more local, SNC national, Diamond international. Portal infrastructure has
+%different levels.
+
+\scipion is a very flexible framework that allows to create many different workflows and to choose among several algorithms at each workflow-step.
 In this work we do not recommend a particular workflow
 over others since we realize that different specimens, microscopes or facilities may have different requirements.
-Nevertheless, since different use cases may clarify \scipionbox capabilities, in the following we describe the pipeline offered by default to the users of three different facilities where \scipionbox has been installed.
+Nevertheless, since different use cases may clarify \scipionbox capabilities, in the following we describe the pipeline offered by default to the users of three different facilities where \scipion has been installed.
 
-\subsection{CNB Facility}
-
-CNB setup is a quite straight forward one.
-Data sets are recorded using EPU which runs in a dedicated Windows box.
-Data are not locally saved but they are stored in a Linux server through a Samba share.
-The Linux server has two Ethernet connections, a 10 Gb point-to-point link with the Windows box and a 1 Gb link with the local network.
-It also has 32 CPUs at 2.40GHz, 62 Gb of RAM memory and 1 \textit{Quadro M4000} GPU.
-\scipionbox is installed in the Linux server.
-
-The first decision a user must face is the backup policy.
-The recommended strategy is either to attach an external hard disk to the server or provide a remote storage location.
-\scipionbox will transfer both raw and processed data as soon as available.
-In this way, access to the Linux server seldom overlaps between two users and once a microscope shift ends, users do not need to wait for their data to be transferred.
-
-Once the image processing pipeline begins \scipionbox will start two independent protocols in parallel.
-The first protocol estimates the camera gain at each pixel for a few  movies ([**REF coss article][**ROB TODO: this protocol is not in streaming]).
-In the CNB setup the microscope camera is calibrated and internally performs the gain correction before writing the movies so the estimated gain image should be almost constant and equal to 1.
-The output of this protocol serves as a verification of the validity of the experimental gain image for the whole data set.
-The second protocol performs Scipion direct detector movie alignment.
-Currently \scipionbox offers  four  methods  for  DDD  movie  alignment: motioncorr (Li et al., 2013),
-correlation (an unpublished CPU version of the motioncorr method developed in Xmipp),
-unblur (Grant and Grigorieff, 2015) and optical flow (Abrishami et al., 2015).
-Within \scipionbox, movie frames can be aligned using any one of these approaches or any combination.
-Following the standard processing workflow, the resulting micrographs are screened to eliminate those with excessive astigmatism or drift.
-Scipion provides three CTF estimation methods:
-CTFFIND (Rohou and Grigorieff, 2015), gCTF (Zhang, 2016) and Xmipp CTF estimation (Sorzano et al., 2007; Vargas et al., 2013).
-At this point the user may compute the CTFs using several algorithms and compare them with the protocol \protocol{CTF consensus} that report up to which frequency two estimation of the same micrograph are equivalent.
-Further automatic filtering based on CTF is available through the protocol \protocol{CTF filtering} that selects micrographs based on the astigmatism, resolution and defocus.
-Another criteria for CTF selection is iciness, that is,  the ratio of the intensity in the resolution band between 3.5 and 3.9 \AA, and between 80 and 30 \AA, which gives a good estimate of the ice crystal content of the vitreous specimen.
-
-[**ROB TODO: I have not added the average protocol since I do not know exactly what it does, if somebody wants to contribute here go ahead]
-
-Up to this point the image processing pipeline is quite solid and may be executed in streaming without user intervention.
-Now, it is time to select (pick) particles.
-We have recently ported the eight different options available at \scipion  to streaming (see XXXX for details[**delarosa up to which point this is true]).
-In this way,  particles may be selected as soon as any new micrograph is available.
-We notice here that some of the ported algorithms require a training step in which either user intervention or information from a previous selection is need.
-For those users that execute several picking algorithms, the protocol \protocol{picking consensus} may be useful.
-\protocol{picking consensus}  calculates the agreement between different particle-picking algorithms and produces a list with the particle selected by most of them.
-[**ROB this algorithm is not in streaming yet, volunteers?]
-
-Two new protocols are under development and will join the described  pipeline in a near future.
-The first one will perform 2D classification and the second one will create a initial 3D map based on this classes.
-\scipion currently implements these protocols but these implementations are not adequate for real time processing.
-
-As commented in section \ref{overall}, to complement the streaming process we have developed a set of monitors and report possibilities.
-By default \protocol{system}, \protocol{CTF} and \protocol{report} monitors are executed.
-\protocol{System} monitor reports on  CPU and GPU utilization, memory consumption and disk I/O [**ROB TODO: last one to be done].
-\protocol{CTF} monitor plots the CTF defocus and astigmatism  and it will raise an alarm if these parameters are not within a given range.
-Finally the \protocol{report} protocol pushes all the created plots and summaries to a web server
-so this information can be accessed from outside.
-
-The proposed pipeline offers quite flexibility to decide which particular algorithms will be used by a given user.
-At CNB (and collaborating institutions) we collect protocol usage statistics and sort the different algorithms by popularity.
-This information is available at  \url{http://calm-shelf-73264.herokuapp.com/report_protocols/protocolTable/} and may help users to select   between the different algorithms.
-
-\subsection{XX Facility}
-
-\subsection{YY Facility}
-
-%cat /proc/cpuinfo
-%\begin{figure}[h]
-%\centering
-%     \includegraphics[width=50mm]{Images/to-be-done}
-%     \caption{TODO: DROP?
-I do not know if an image showing CNB setup is needed this image should show CNB setup, microscope connected to control-PC which saves data in \scipionbox using samba.
-Show Ethernet speed.
-Soy connection to external disk/host.
-%     Describe number GPUs, CPUs, SSD disks.
-%     }
-%     \label{cnbsetup}
-%\end{figure}
-
-%One of the most frequent requests by both beginners, and even experienced digital darkroom workers, is for detailed step-by-step instructions on how to perform various basic tasks in Photoshop.
-
-%One of the key goals of ScipionBox is to guarantee interoperability between algorithms implemented in different software packages.
-The possibility of choosing among several algorithms is provided at all levels of the workflow, in many cases adding the possibility of comparing the results of these methods to detect inconsistencies.
-%and do not need to be concerned with how the rest of the workflow is allocated.
-
-%ScipionBox design is agnostic to a particular workflow or software packet.
-
-%%Export options?
+\input{3.workflow_cnb.tex}
+\input{3.workflow_delarosa.tex}
+\input{3.workflow_diamond.tex}
 

### 2018-02-15 3.workflow_cnb.tex
split section 3 in three centers

--- 2017-06-07 workflow.tex
@@
-\section{Processing with ScipionBox}
-
-\scipionbox is a very flexible framework that allows to create many different workflows and to choose among several algorithms at each workflow-step.
-In this work we do not recommend a particular workflow
-over others since we realize that different specimens, microscopes or facilities may have different requirements.
-Nevertheless, since different use cases may clarify \scipionbox capabilities, in the following we describe the pipeline offered by default to the users of three different facilities where \scipionbox has been installed.
-
 \subsection{CNB Facility}
+%Here there are Scipion general features and CNB specific. Move the general to the Processing with Scipion section.
 
 CNB setup is a quite straight forward one.
-Data sets are recorded using EPU which runs in a dedicated Windows box.
-Data are not locally saved but they are stored in a Linux server through a Samba share.
-The Linux server has two Ethernet connections, a 10 Gb point-to-point link with the Windows box and a 1 Gb link with the local network.
-It also has 32 CPUs at 2.40GHz, 62 Gb of RAM memory and 1 \textit{Quadro M4000} GPU.
-\scipionbox is installed in the Linux server.
+From a network point of view there are three different levels of accessibility: (1) a private network that cannot be accessed by regular users, (2) the CNB LAN (local area network)  that can be accessed by any user which works inside the hosting  institution and (3) finally, there is a single computer that may be accessed from outside the institution and therefore, it can see the WAN (wide area network).
+
+Data sets are recorded using EPU.
+The microscope control system has a 70 TB NAS (Network-attached storage) unit for saving movies as they are obtained (we will refer to this NAS as microscope-NAS).
+This storage unit is shared through a Samba share and connected to the private network using a 10Gb connexion.
+\scipion is executed on a Linux server which has two 1Gb NICs (network interface controllers) that connect it to the private and the LAN networks (we will refer to this machine as \scipionbox).
+This machine has 32 CPUs at 2.40GHz, 62 Gb of RAM memory and 2 \textit{Quadro M4000} GPUs.
+Finally, a second NAS with 4 1Gb NICs is linked to the private, LAN and WAN networks (we will refer to this second NAS as output-NAS).
+
+The data flow is a follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox will temporary read the movies from 
+the microscope-NAS but will not store them locally, (c) each hour, output-NAS will connect to \scipionbox and copy or update the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup the \scipion project and the movies to a USB disk or serve this information through the Internet and (e) in parallel \scipionbox transfers to  output-NAS different html reports that may be checked by users or staff with a HTML browser.
+[** create figure]
+
+Discuss transfer rates
+Discuss typical workflow
 
@@
 This information is available at  \url{http://calm-shelf-73264.herokuapp.com/report_protocols/protocolTable/} and may help users to select   between the different algorithms.
-
-\subsection{XX Facility}
-
-\subsection{YY Facility}
-
-%cat /proc/cpuinfo
-%\begin{figure}[h]
-%\centering
-%     \includegraphics[width=50mm]{Images/to-be-done}
-%     \caption{TODO: DROP?
-I do not know if an image showing CNB setup is needed this image should show CNB setup, microscope connected to control-PC which saves data in \scipionbox using samba.
-Show Ethernet speed.
-Soy connection to external disk/host.
-%     Describe number GPUs, CPUs, SSD disks.
-%     }
-%     \label{cnbsetup}
-%\end{figure}
-
-%One of the most frequent requests by both beginners, and even experienced digital darkroom workers, is for detailed step-by-step instructions on how to perform various basic tasks in Photoshop.
-
-%One of the key goals of ScipionBox is to guarantee interoperability between algorithms implemented in different software packages.
-The possibility of choosing among several algorithms is provided at all levels of the workflow, in many cases adding the possibility of comparing the results of these methods to detect inconsistencies.
-%and do not need to be concerned with how the rest of the workflow is allocated.
-
-%ScipionBox design is agnostic to a particular workflow or software packet.
-
-%%Export options?
-
-\subsection{What will do in future}
-%Particle picking
-%Extract particle
-%2D classify (either classes or identify outliers)
-%Initial model
-%More complex reports
-[** ROB This is already commented so this section disappears ]

### 2018-02-15 3.workflow_diamond.tex
split section 3 in three centers

\subsection{Diamond Light Source Facility??? or eBIC???}

\begin{verbatim}

We need to cover here differences with CNB setup

 DEVELOPER RELATED
 * Describe infrastructure in diamond: file system (distributed, user permissions),
                                       no graphical output, queues, 
 * Describe Scipion modification:      import (detect new files by socket message)
                                       monitor, report to ISPyB 
                                       scipion headless
                                       execution as script

USER RELATED
 * how users interact with scipion?
 * how data is visualized?
\end{verbatim}

### 2018-02-19 2.overall.tex
re-done up to 3.workflow_cnb

@@
-\section{Overall System Design: Monitors, Consensus and Streaming}
+\section{Main Modifications introduced in  \scipion: Streaming, Monitors and Consensus}
 \label{overall}
@@
 In the following we describe the main modifications introduced in the original \scipion version in order to facilitate its deployment in centralized facilities.
-So far, \scipion is running in three of such facilities, and each installation has been tailored to the particular needs of each facility.
-The original design satisfied the requirements of our home facility (National Center for Biotechnology, Madrid, Spain), in which \scipion automatically fetches newly recorded movies from a network mounted disk.
-After this first installation, \scipion was deployed on XXXXX (a centralized facility located at Stockholm [**provide full identification @delarosa]) where a rigorous booking system is followed.
-There, \scipion was adapted to interface with this booking platform in order to produce better reports.
-Lately,  \scipion has been installed in a large synchrotron  such as Diamond (Oxford, UK) where projects are handled by ISPyB (a customized laboratory management information system -LIMBS- \citep{Delageniere2011}) and data are saved on a distributed file system that heavily penalizes access to lists of files.
-In this environment, \scipion needs to constantly interchange information with ISPyB, and it cannot longer check if there is a new file in a particular storage device but must wait to be informed by the system that a new file has been created.
-All these particularities, that will be described in detail later, could only be handled thanks to a careful design that prioritizes flexibility at many levels.
+
+The main change introduced in \scipion is \emph{stream processing}. Stream processing has been designed to analyze and react on real-time so that enables analysis of data as it is produced.
+In this way a new level of parallelization has been reached since the  data generated by the microscope continuously flows through a chain of protocols.
+For example, as movies are imported to the system they are aligned and their CTF is estimated.
+In \scipion stable version users may find  the first steps in the image processing adapted to stream.
+These steps include: movie alignment, CTF estimation and particle picking.
+In \scipion developers version  several 2D classification algorithms are available.
+
+\scipion stream implementation is based on periodic updates of the protocol input and output data during the image processing.
+Since \scipion runs on threads, it is possible to devote one thread to update these information while the rest of the CPU power is devoted to run the different protocols.
+The use of threads marks an important difference between running a script or using \scipion, since  each module is run independently creating a very flexible processing framework.
+For example, do you want to add a new protocol in your workflow while it is running?
+No problem, do it.
+\scipion allows users to modify  workflows on the fly without modifying a single code line.
+The modification will not interfere with the already executed steps and may take advantage from all already computed data.
+Did something go wrong and the microscope stopped?
+By default \scipion will stop after 24 hours with no microscope activity, but if you wish to continue your acquisition after a stop just press the continue button, there is no need to restart from the beginning.
 
@@
 CTF-monitor may raise an alert if the defocus values are above or below certain thresholds and continuously generate HTML files so that users and staff may easily follow the data acquisition and processing in-house or remotely.
-At the project's end, a review of the processing history is produced.
+An example of this reports may be seen at \url{http://nolan.cnb.csic.es/scipionbox/} [**ROB NOTE: select a nice one an complete URL]%At the project's end, a review of the processing history is produced.
 [** URL to report example?]
@@
 For a given logical step (for example, particle picking) these protocols check if the datasets obtained from the same input data using different algorithm are consistent.
-Continuing with the particle picking example, \emph{consensus picking} will compare the particles selected for each one of the executed particle-picking algorithms and report on the particles selected by most of them.
-
-The main change introduced in \scipion is \emph{stream processing}. Stream processing has been designed to analyze and react on real-time so that enables analysis of data as it is produced.
-In this way a new level of parallelization has been reached since the  data generated by the microscope continuously flows through a chain of protocols.
-For example, as movies are imported to the system they are aligned and their CTF is estimated.
-In \scipion stable version users may find  the first steps in the image processing adapted to stream.
-These steps include: movie alignment, CTF estimation and particle picking.
-In \scipion developers version all several 2D classification algorithms algorithms are available.
-
-\scipion stream implementation is based on periodic updates of the protocol input and output data during the image processing.
-Since \scipion runs on threads, it is possible to devote one thread to update these information while the rest of the CPU power is devoted to run the different protocols.
-The use of threads marks an important difference between running a script or using \scipion, since it means that each module is run independently.
-For example, Do you want to add a new protocol or monitor in your workflow while it is running?
-No problem, do it.
-\scipion allows users to modify  workflows on the fly without modifying a single code line.
-The modification will not interfere with the already executed steps and may take advantage from all already computed data.
-Did something go wrong and the microscope stopped?
-By default \scipion will stop after 24 hours with no microscope activity, but if you wish to continue your acquisition after a stop just press the continue button, there is no need to restart from the beginning.
+Continuing with the particle picking example, \emph{consensus picking} will compare the particles selected for each one of the executed particle-picking algorithms and produce as output a particle set containing those particles selected by most of the algorithms.
 

### 2018-02-19 3.workflow.tex
re-done up to 3.workflow_cnb

@@
-\section{Processing with ScipionBox}
+\section{\scipion Setup in the Different Facilities}
 
@@
 
-\scipion is a very flexible framework that allows to create many different workflows and to choose among several algorithms at each workflow-step.
-In this work we do not recommend a particular workflow
-over others since we realize that different specimens, microscopes or facilities may have different requirements.
+So far, \scipion is being used routinely in four Cryo-EM facilities: \emph{Science for life Laboratory} (SciLifeLab, \url{http://www.scilifelab.se}),  \emph{European Synchrotron Radiation Facility} (ESRF, \url{http://www.esrf.eu/}), \emph{Diamond Light Source Synchrotron} (Diamond, \url{http://www.diamond.ac.uk}) and \emph{National Center for Biotechnology} (CNB, \url{http://www.cnb.csic.es}).
+\scipion original design satisfied the requirements of our home facility (CNB), in which \scipion automatically fetches newly recorded movies from a network mounted disk.
+After this first installation, \scipion was deployed on SciLifeLab (a centralized facility located at Stockholm) where a rigorous booking system is followed.
+There, \scipion was adapted to interface with this booking platform in order to produce better reports.
+Lately,  \scipion has been installed in a large synchrotron  such as ESRF and Diamond  where projects are handled by ISPyB (a customized laboratory management information system -LIMBS- \citep{Delageniere2011}) and data are saved on a distributed file system that heavily penalizes access to lists of files.
+In this environment, \scipion needs to constantly interchange information with ISPyB, and it cannot longer check if there is a new file in a particular storage device but must wait to be informed by the system that a new file has been created.
+All these particularities, that will be described in detail in the following, could only be handled thanks to a careful design that prioritizes flexibility at many levels.
+
+%\scipion is a very flexible framework that allows to create many different workflows and to choose among several algorithms at each workflow-step.
+In this work we do not recommend a particular workflow  over others since we realize that different specimens, microscopes or facilities may have different requirements.
 Nevertheless, since different use cases may clarify \scipionbox capabilities, in the following we describe the pipeline offered by default to the users of three different facilities where \scipion has been installed.

### 2018-02-19 2.overall.tex
move the workflow section on CNB to a more general place

@@
 The first type of metaprotocols are called \emph{monitors} and are used to produce live analysis plots, generate reports or raise alerts when some problems are detected.
-A monitor example is the \emph{CTF-monitor}, that checks the computed defocus values for
+A monitor example is the \protocol{CTF-monitor}, that checks the computed defocus values for
 each micrograph as they are generated.
@@
 %Export options?
+
+\section{Image Processing Workflows}
+
+\scipion is a very flexible framework that allows to create many different workflows and to choose among several algorithms at each workflow-step.
+In this work we do not recommend a particular workflow  over others since we realize that different specimens, microscopes or facilities may have different requirements.
+Nevertheless, since a use case may clarify \scipion capabilities, in the following we describe the pipeline offered by default
+at our home facility.
+
+The workflow starts two independent protocols in parallel.
+The first protocol estimates the camera gain at each pixel for a few  movies \citep{sorzano2018}.
+% In the CNB setup the microscope camera is calibrated and internally performs the gain correction before writing the movies so the estimated gain image should be almost constant and equal to 1.
+The output of this protocol serves as verification of the validity of the experimental gain image for the whole data set.
+The second protocol performs Scipion direct detector movie alignment.
+Currently \scipion offers  four  methods  for  DDD  movie  alignment: motioncorr2 (Li et al., 2013), correlation (an unpublished CPU version of the motioncorr2 method developed in Xmipp), unblur (Grant and Grigorieff, 2015) and optical flow (Abrishami et al., 2015).
+Within \scipion, movie frames can be aligned using any one of these approaches or any combination.
+Following the standard processing workflow, the resulting micrographs are screened to eliminate those with excessive astigmatism or drift.
+Scipion provides three CTF estimation methods:
+CTFFIND (Rohou and Grigorieff, 2015), gCTF (Zhang, 2016) and Xmipp CTF estimation (Sorzano et al., 2007; Vargas et al., 2013).
+At this point the user may compute the CTFs using several algorithms and compare them with the protocol \protocol{CTF consensus} that report up to which frequency two estimation of the same micrograph are equivalent.
+Further automatic filtering based on CTF is available through the protocol \protocol{CTF filtering} that selects micrographs based on the astigmatism, resolution and defocus.
+Another criteria for CTF selection is iciness, that is,  the ratio of the intensity in the resolution band between 3.5 and 3.9 \AA, and between 6 and 30 \AA, which gives a good estimate of the ice crystal content of the vitreous specimen.
+[**ROB: do we have iceness implemented?]
+
+Up to this point the image processing pipeline is quite solid and may be executed in streaming without user intervention.
+The next step is to select (pick) particles.
+We have recently ported all the  different options available at \scipion  to streaming.
+In this way,  particles may be selected as soon as any new micrograph is available.
+We note here that some of the ported algorithms require a training step in which either user intervention or information from a previous selection is need.
+For those users that execute several picking algorithms, the protocol \protocol{picking consensus} may be useful.
+\protocol{picking consensus}  calculates the agreement between different particle-picking algorithms and produces a list with the particle selected by most of them.
+
+Two new protocols are under development and will join the described  pipeline in a near future.
+The first one will perform 2D classification and the second one will create a initial 3D map based on this classes.
+A first version of these protocols is available in the \scipion developement version.
+
+As commented in section \ref{overall}, to complement the streaming process we have developed a set of monitors and report possibilities.
+By default \protocol{system}, \protocol{CTF} and \protocol{report} monitors are executed.
+\protocol{System monitor} reports on  CPU and GPU utilization, memory consumption and disk I/O.
+\protocol{CTF monitor} plots the CTF defocus and astigmatism  and it will raise an alarm if these parameters are not within a given range.
+Finally the \protocol{report} protocol pushes all the created plots and summaries to a web server
+so this information can be accessed from outside.
+
+The proposed pipeline offers quite flexibility to decide which particular algorithms will be used by a given user.
+At CNB (and collaborating institutions) we collect protocol usage statistics and sort the different algorithms by popularity.
+This information is available at  \url{http://scipion.i2pc.es/report_protocols/protocolTable/} and may help users to select   between the different algorithms.

### 2018-02-19 3.workflow_cnb.tex
re-done up to 3.workflow_cnb
move the workflow section on CNB to a more general place

@@
-\subsection{CNB Facility}
-%Here there are Scipion general features and CNB specific. Move the general to the Processing with Scipion section.
+\subsection{CNB}
 
-CNB setup is a quite straight forward one.
+We are going to structure this section in three pats.
+In the first one we discuss the network setup that allow us to move the data from the microscope to the user laboratory.
+Then we present the standard image processing workflow and finally we comment on the additional software (different from \scipion) used in process data.
+
+\subsubsection{Networking}
+
+At the beginning CNB networking setup was a quite straight forward one, movies where stored in the  same machine that ran \scipion and processed data was saved to external USB hard disks by this same machine.
+The situation changed when the new Falcon-III detector was installed.
+After that, data was produced at sustained speeds as high as 90 MB/sec (when acquiring four movies by hole in lineal mode) and the main problem became to move smoothly the data through the entire pipeline from the microscope to the final user laboratory.
+In the following we describe such a pipeline.
+
 From a network point of view there are three different levels of accessibility: (1) a private network that cannot be accessed by regular users, (2) the CNB LAN (local area network)  that can be accessed by any user which works inside the hosting  institution and (3) finally, there is a single computer that may be accessed from outside the institution and therefore, it can see the WAN (wide area network).
@@
 This storage unit is shared through a Samba share and connected to the private network using a 10Gb connexion.
-\scipion is executed on a Linux server which has two 1Gb NICs (network interface controllers) that connect it to the private and the LAN networks (we will refer to this machine as \scipionbox).
-This machine has 32 CPUs at 2.40GHz, 62 Gb of RAM memory and 2 \textit{Quadro M4000} GPUs.
+\scipion is executed on a Linux server which has two 1Gb NICs (network interface controllers) that connect it to the private and the LAN networks respectively (we will refer to this machine as \scipionbox).
+\scipionbox has 32 CPUs at 2.40GHz, 62 Gb of RAM memory and 2 \textit{Quadro M4000} GPUs.
 Finally, a second NAS with 4 1Gb NICs is linked to the private, LAN and WAN networks (we will refer to this second NAS as output-NAS).
@@
 The data flow is a follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox will temporary read the movies from 
-the microscope-NAS but will not store them locally, (c) each hour, output-NAS will connect to \scipionbox and copy or update the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup the \scipion project and the movies to a USB disk or serve this information through the Internet and (e) in parallel \scipionbox transfers to  output-NAS different html reports that may be checked by users or staff with a HTML browser.
-[** create figure]
+the microscope-NAS but will not store them locally, (c) each hour, output-NAS will connect to \scipionbox and copy or update the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup the \scipion project and the movies to a USB disk or serve this information through the Internet and (e) in parallel \scipionbox transfers to  output-NAS different html reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox}).
 
-Discuss transfer rates
-Discuss typical workflow
+Although we try to discourage the use of external USB hard disks versus network.
+At present, this is the preferred option for more of our users.
+USB 3.0 drives can theoretically handle data flows of 300 MB/s, but our experience is that, in this busy environment, many of them  have trouble exceeding 50 MB/s.
+In the ideal situation both raw and processed data are transfered as soon as available.
+In this way, access to the servers seldom overlaps between two users and once a microscope shift ends, users do not need to wait for their data to be transferred.
+fggdfgdsgdsfgdf.
+Comment on ssh cifers
 
-The first decision a user must face is the backup policy.
-The recommended strategy is either to attach an external hard disk to the server or provide a remote storage location.
-\scipionbox will transfer both raw and processed data as soon as available.
-In this way, access to the Linux server seldom overlaps between two users and once a microscope shift ends, users do not need to wait for their data to be transferred.
+comment on iperf and smbdata
 
-Once the image processing pipeline begins \scipionbox will start two independent protocols in parallel.
-The first protocol estimates the camera gain at each pixel for a few  movies ([**REF coss article][**ROB TODO: this protocol is not in streaming]).
-In the CNB setup the microscope camera is calibrated and internally performs the gain correction before writing the movies so the estimated gain image should be almost constant and equal to 1.
-The output of this protocol serves as a verification of the validity of the experimental gain image for the whole data set.
-The second protocol performs Scipion direct detector movie alignment.
-Currently \scipionbox offers  four  methods  for  DDD  movie  alignment: motioncorr (Li et al., 2013),
-correlation (an unpublished CPU version of the motioncorr method developed in Xmipp),
-unblur (Grant and Grigorieff, 2015) and optical flow (Abrishami et al., 2015).
-Within \scipionbox, movie frames can be aligned using any one of these approaches or any combination.
-Following the standard processing workflow, the resulting micrographs are screened to eliminate those with excessive astigmatism or drift.
-Scipion provides three CTF estimation methods:
-CTFFIND (Rohou and Grigorieff, 2015), gCTF (Zhang, 2016) and Xmipp CTF estimation (Sorzano et al., 2007; Vargas et al., 2013).
-At this point the user may compute the CTFs using several algorithms and compare them with the protocol \protocol{CTF consensus} that report up to which frequency two estimation of the same micrograph are equivalent.
-Further automatic filtering based on CTF is available through the protocol \protocol{CTF filtering} that selects micrographs based on the astigmatism, resolution and defocus.
-Another criteria for CTF selection is iciness, that is,  the ratio of the intensity in the resolution band between 3.5 and 3.9 \AA, and between 80 and 30 \AA, which gives a good estimate of the ice crystal content of the vitreous specimen.
-
-[**ROB TODO: I have not added the average protocol since I do not know exactly what it does, if somebody wants to contribute here go ahead]
-
-Up to this point the image processing pipeline is quite solid and may be executed in streaming without user intervention.
-Now, it is time to select (pick) particles.
-We have recently ported the eight different options available at \scipion  to streaming (see XXXX for details[**delarosa up to which point this is true]).
-In this way,  particles may be selected as soon as any new micrograph is available.
-We notice here that some of the ported algorithms require a training step in which either user intervention or information from a previous selection is need.
-For those users that execute several picking algorithms, the protocol \protocol{picking consensus} may be useful.
-\protocol{picking consensus}  calculates the agreement between different particle-picking algorithms and produces a list with the particle selected by most of them.
-[**ROB this algorithm is not in streaming yet, volunteers?]
-
-Two new protocols are under development and will join the described  pipeline in a near future.
-The first one will perform 2D classification and the second one will create a initial 3D map based on this classes.
-\scipion currently implements such protocols but the implementation is not adequate for real time processing.
-
-As commented in section \ref{overall}, to complement the streaming process we have developed a set of monitors and report possibilities.
-By default \protocol{system}, \protocol{CTF} and \protocol{report} monitors are executed.
-\protocol{System} monitor reports on  CPU and GPU utilization, memory consumption and disk I/O [**ROB TODO: last one to be done].
-\protocol{CTF} monitor plots the CTF defocus and astigmatism  and it will raise an alarm if these parameters are not within a given range.
-Finally the \protocol{report} protocol pushes all the created plots and summaries to a web server
-so this information can be accessed from outside.
-
-The proposed pipeline offers quite flexibility to decide which particular algorithms will be used by a given user.
-At CNB (and collaborating institutions) we collect protocol usage statistics and sort the different algorithms by popularity.
-This information is available at  \url{http://calm-shelf-73264.herokuapp.com/report_protocols/protocolTable/} and may help users to select   between the different algorithms.
+\subsubsection{EMadmin}

### 2018-02-19 2.overall.tex
going home

@@
 Another criteria for CTF selection is iciness, that is,  the ratio of the intensity in the resolution band between 3.5 and 3.9 \AA, and between 6 and 30 \AA, which gives a good estimate of the ice crystal content of the vitreous specimen.
-[**ROB: do we have iceness implemented?]
 
@@
 The next step is to select (pick) particles.
-We have recently ported all the  different options available at \scipion  to streaming.
+We have recently ported the eight different options available at \scipion  to streaming.
 In this way,  particles may be selected as soon as any new micrograph is available.
@@
 
-Two new protocols are under development and will join the described  pipeline in a near future.
+Two new streaming protocols are under development and will join the described  pipeline in a near future.
 The first one will perform 2D classification and the second one will create a initial 3D map based on this classes.
-A first version of these protocols is available in the \scipion developement version.
+A first version of these protocols is available in the \scipion developers version.
 
-As commented in section \ref{overall}, to complement the streaming process we have developed a set of monitors and report possibilities.
-By default \protocol{system}, \protocol{CTF} and \protocol{report} monitors are executed.
-\protocol{System monitor} reports on  CPU and GPU utilization, memory consumption and disk I/O.
+As commented in section \ref{overall}, to monitor, chart and create alerts  we have developed a set of metaprotocols.
+By default \protocol{system}, \protocol{CTF} and \protocol{report monitors} are executed.
+\protocol{System monitor} reports on  CPU and GPU utilization, memory consumption and, disk I/O.
 \protocol{CTF monitor} plots the CTF defocus and astigmatism  and it will raise an alarm if these parameters are not within a given range.
 Finally the \protocol{report} protocol pushes all the created plots and summaries to a web server
-so this information can be accessed from outside.
+so this information can be accessed by users who are not physically in the microscope room.
 

### 2018-02-20 6.acknowledgments.tex
Updates from ShareLaTeX
ROB: done with the editing prior to 28 Feb meeting

--- 2017-06-07 acknowledgments.tex
@@
 
+----------------
+
     AFTER THIS LINE I JUST SAVED SOME TEXT THAT MAY BE USEFUL IN THE FUTURE DISREGARD THIS MATERIAL
+
     --------------------
 
-\section{Getting Started}
+\section{Project stability and funding}
 
-\section{Project stability and funding. getting started}
-
-Scipion and ScipionBox development is been done in the context of the Instruct Image Processing Cen-
+Scipion development is been done in the context of the Instruct Image Processing Cen-
 ter (I2PC, http://i2pc.cnb.csic.es). Instruct is the European initiative for Strategic

### 2018-02-20 0.abstract.tex
w

@@
 Three dimensional electron microscopy (3DEM) is becoming a very data-intensive field in which vast amounts of experimental data are acquired at high speed.
-To manage such large-scale image data  projects, we have previously described  a modular workflow system called \scipion.
+To manage such large-scale image data  projects, we have previously described  a modular workflow system called \scipion \citep{conesa2018}.
 We present here the first main extension of \scipion that allows to process data in streaming.
@@
 This summary may be  generated in HTML format that can be easily published though a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided.
-At present, \scipion has been installed and is in production mode in three facilities.
-And it is being considered in several more.
+At present, \scipion has been installed and is in production mode in four Cryo-EM facilities.
+And it is being considered in many more.
 

### 2018-02-20 1.introduction.tex
w

@@
 However, the next step,  which is to provide image processing while  data are being collected, is not so well served.
-In fact, in most places this is accomplished through home made scripts.
+In fact, in most places this is accomplished through home made scripts (i.e.
+\citep{Pichkur2018}).
 Exceptions are UCSFImage4 \citep{Li2015} and Focus \citep{Biyani2017}.
-As we will explain here, \scipion differs from these solutions in a larger flexibility both from the point of view of users and facility staff.
+As we will show in this article, \scipion differs from these solutions in a larger flexibility both from the point of view of users and facility staff.

### 2018-02-20 3.workflow.tex
going home

@@
 
-%Comment change in scale in the three examples: CNB, SNC and DIAMOND
-%more local, SNC national, Diamond international. Portal infrastructure has
-%different levels.
+The data flow created by an electron microscope requires a careful design of the IT infrastructures.  Different facilities has adopted different solutions and therefore, 
+how \scipion  access data and interchange information is very facility dependent. In this section we discuss  the different solutions applied in the different facilities in which \scipion is running.
 

### 2018-02-20 3.workflow_cnb.tex

@@
 
-We are going to structure this section in three pats.
+This section in divided in two parts.
 In the first one we discuss the network setup that allow us to move the data from the microscope to the user laboratory.
-Then we present the standard image processing workflow and finally we comment on the additional software (different from \scipion) used in process data.
+Then  we comment on an additional software (called  \emadmin) that connects \scipion with the microscope acquisition
+software (EPU) and records the facility activity.
 
-\subsubsection{Networking}
+\subsubsection{Network Setup}
 
@@
 The data flow is a follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox will temporary read the movies from 
-the microscope-NAS but will not store them locally, (c) each hour, output-NAS will connect to \scipionbox and copy or update the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup the \scipion project and the movies to a USB disk or serve this information through the Internet and (e) in parallel \scipionbox transfers to  output-NAS different html reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox}).
+the microscope-NAS but will not store them locally, (c) each hour, output-NAS will connect to \scipionbox and copy or update the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup the \scipion project and the movies to a USB disk or serve this information through the Internet and, (e) in parallel, \scipionbox transfers to  output-NAS different html reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox}).
 
-Although we try to discourage the use of external USB hard disks versus network.
-At present, this is the preferred option for more of our users.
+Although we try to discourage the use of external USB hard disks versus network as way of retrieving the raw and processing data, this is the preferred option for more of our users.
 USB 3.0 drives can theoretically handle data flows of 300 MB/s, but our experience is that, in this busy environment, many of them  have trouble exceeding 50 MB/s.
-In the ideal situation both raw and processed data are transfered as soon as available.
-In this way, access to the servers seldom overlaps between two users and once a microscope shift ends, users do not need to wait for their data to be transferred.
-fggdfgdsgdsfgdf.
-Comment on ssh cifers
+In most cases, 50 MB/s is enough to transfer the data before the following microscope shift starts.
+Nevertheless, real-world is tough and it happens that two, or even three,  users try to download their data simultaneously.
+Fortunately, output-NAS works as front-end and isolates the data acquiring and processing infrastructure from the user downloading.
+Network users access to output-NAS using a chrooted common account.
+This account only allows rsync connections and make impossible to retrieve data if the name of the project to be downloaded is unknown.
+In order to speed up transfer weak ciphers (as \emph{arcfour}) has been activated in output-NAS.
 
-comment on iperf and smbdata
+\subsubsection{\emadmin}
 
-\subsubsection{EMadmin}
+\emadmin was born as a script that created a normalized tree of directories  in which to store the movies acqured by the microscope and help users to choose between a few predetermined image processing workflows.
+At present it was grown in a client-server application that in addition to the old script capacities it has incorporate the following functions:
+
+\begin{enumerate}
+ \item Transparent addition of new workflows. \scipion is able to export any executed workflow as a json file. \emadmin can import this files as use them to create new predetermined workflows.
+ \item Automatic creation of pdf user's report containing: microscope acquisition parameters and histograms summarizing data resolution and  defocus. 
+ \item The above mentioned information is stored in a database for further analysis. By default the system plots the  average resolution and astigmatism of the different acquisitions versus time
+ \item Remote management. \emadmin is a client server application, the client maybe executed in a machine different from the one that executes \scipion
+ \item Automatic backup of the \scipion project
+ \item \scipion is able to create live html reports on the image acquisition process. \emadmin keeps track of all of them and facilitates the access to this information
+\end{enumerate}
+
+\emadmin is free software available at github repository (\url{https://github.com/rmarabini/webservices/tree/master/EMadmin}).
+You are welcomed to try it but the software is provided as it is and no support or maintenance service is offered.

### 2018-02-20 5.conclusions.tex
move the workflow section on CNB to a more general place
going home

@@
 \section{Discussion}
-[*+TODO]: Discuss the obtained numbers in the result section
+[**TODO]: Discuss the obtained numbers in the result section
+[**REmove this section?]
 
@@
 The new generation of high-end cryo electron microscopes and direct electron detectors are behind the recent increase in cryo-EM  popularity.
-However, the run and maintenance costs of these new microscopes are high and can only be justified keeping the microscopes up and running 24 hours a day.
-In this way, microscopes tent to belong to centralize facilities where users are allocated microscope time (shifts) to perform a variety of research.
-Available shifts are a scarce commodity and therefore users (and staff) need to be able to monitor the quality of the acquisition process and get as soon as possible a first glimpse of the reconstructed specimen in order to guaranty an early detection of any issue.
+However, the run and maintenance costs of these new microscopes are high and can only be justified by keeping the microscopes up and running 24 hours a day.
+Therefore, microscopes tend to belong to centralized facilities where users are allocated microscope time (shifts) to perform a variety of tasks.
+Available shifts are a scarce commodity and therefore users (and staff) need to be able to monitor the quality of the acquisition process and obtain as soon as possible a first glimpse of the reconstructed specimen in order to provide an early detection of any issue.
 
- Despite workflow systems have been explored and used extensively in bioinformatics, there is little progress in that direction in the cryo-EM community.
-\scipionbox has been designed following and strict workflow paradigm.
-This approach is specially useful for tasks that need to go from a given start (movies) to a defined end (3D map) but there are many different paths (algorithms) to get there.
+ Workflow systems have been explored and used extensively in bioinformatics but there is relatively little progress in that direction in the cryo-EM community.
+\scipion has been designed following a strict workflow paradigm.
+This approach is specially useful for tasks that need to go from a given start (movies) to a defined end (3D map) but there are many different paths (algorithms) and steps to get there.
 
@@
 They can potentially play with the pipeline in ways that aren't possible when having a programmer write code for them.
-Furthermore, no having the processing path hardcoded in the program means you don't have to recompile, so changes can be quickly tested   even at run time.
-
- \scipionbox workflow approach provides provenance (transparency) since for any created object (micrograph, CTFs, class, etc) it is stored which program creates it, using which parameters, and the exact path followed to create it.
+Furthermore, not having the processing path hardcoded in the program means you don't have to recompile, so changes can be quickly tested   even at run time.
+The \scipion workflow approach provides provenance, that is,  for any created object (micrograph, CTFs, class, etc) \scipion stores metadata such as which programs, which parameters, and which exact path was followed to create it.
 In this way,  real tracking and reproducibility are provided.
 
-\scipionbox is part of the \scipion package which is provided 
+ \scipion stream capabilities can handle large and continuously produced volumes of data.
+The benefits of streaming are clear: (1) it helps to process data as soon as it is generated,  providing almost instant results on data quality; (2) it saves user processing time at their home institutions since part of the processing is done while data is being acquired; and (3)
+ it further reduces the overall experimental time by allowing processing the data while it is transferred to the final user laboratory.
+\scipion has been deployed in cloud and there is a  a public image, with all the required preinstalled cryoEM software,  available at Amazon WebServices EC2 and EGI federated clouds.
+These images make use of \scipion streaming capabilities, that is, they are able to process data as soon as it has been transfered to an Amazon storage unit, and can be used to handle  production peaks.
+
+\scipion is provided 
 freely as open source software. Online documentation
 describing \scipion download and installation is available
-at \url{http://scipion.cnb.csic.es/m/download_form/} and \url{https://github.com/I2PC/scipion/wiki/How-to-Install}, respectively.\\\\
+at \url{http://scipion.cnb.csic.es/m/download_form/} and \url{https://github.com/I2PC/scipion/wiki/How-to-Install}, respectively.
+Cloud installation instruction are available at \url{https://github.com/I2PC/scipion/wiki/Scipion-in-the-Cloud}.
+Other public software mentioned in this article as \emadmin, XXXX and YYYY are available at \url{https://github.com/rmarabini/webservices/tree/master/EMadmin}, www eee respectively.
+If you would like more information about how to install \scipion or you have suggestions on how to improve it contact us at scipion@cnb.csic.es.
 

### 2018-02-22 3.workflow_scilifelab.tex
Intro for Facility

\subsection{Swedish National CryoEM Facility}

\begin{verbatim}
We need to cover here differences with CNB setup
 CNB is a stand alone center. It does not work as a real National Facility
 SNC already had an application portal and Scipion needs to interact
 with it. 
 - initial script
 - same as CNB until session end
 - when acquisistion done contaCT PORTAL AND UPLOAD REPORT
\end{verbatim}

The Swedish National Cryo-EM Facility offers access to state-of-the-art equipment and expertise 
in single particle cryo-EM and cryo electron tomography (cryo-ET). The Facility has two nodes: at SciLifeLab 
in Stockholm and at Umeå University. SciLifeLab in Stockholm offers single-particle service with a 
Talos Arctica for sample optimisation and a Titan Krios for high-resolution data collection. 
The Umeå node is expected to become operational during 2018, and will offer cryo-ET with a 
Titan Krios and a Scios DualBeam SEM.

The facilities can be accessed by Swedish researchers through a peer-reviewed routine. Applications can be 
submitted through an Application Portal (reference???) and will be evaluated once every three months based on their 
scientific merit and technical feasibility by a national Project Evaluation Committee. On the other side, 
some time of the facility is reserved for internal research groups at Stockholm University. Internally, the allocation
of the microscope (and other instruments) time is managed by a Booking System (reference???).

### 2018-02-27 3.workflow_scilifelab.tex
A bit more on SciLifeLab intro

@@
 of the microscope (and other instruments) time is managed by a Booking System (reference???).
+
+At the beginning of 2016, the Cryo-EM facility at SciLifeLab became one of the early adopters of Scipion 
+for on the fly data processing. Before, there were some home-made scripts to perform motion correction (with motioncor)
+and CTF estimation (with Gctf). The microscope operators recognized the importance of  
+a more general framework with the possibility to be extended with new programs. Another strong point was the 
+the availability of graphical tools for data analysis and quick feedback about the data collection. Initially, the 
+same setup script from the CNB was used, but it has some limitations for this more complex setup with more 
+requirements. In the following sections we will briefly describe the computational infrastructure at the
+SciLifeLab node and the implementation of a Session Wizard to automate the initial setup while fetching 
+information from other sources external to Scipion.
+
+\subsubsection{Infrastructure}
+
+\subsubsection{Session Wizard}

### 2018-02-28 3.workflow_scilifelab.tex
Added description of the Session Wizard

@@
 \subsection{Swedish National CryoEM Facility}
-
-\begin{verbatim}
-We need to cover here differences with CNB setup
- CNB is a stand alone center. It does not work as a real National Facility
- SNC already had an application portal and Scipion needs to interact
- with it. 
- - initial script
- - same as CNB until session end
- - when acquisistion done contaCT PORTAL AND UPLOAD REPORT
-\end{verbatim}
 
@@
 the availability of graphical tools for data analysis and quick feedback about the data collection. Initially, the 
-same setup script from the CNB was used, but it has some limitations for this more complex setup with more 
+same setup script from the CNB was used, but it had some limitations for this more complex setup with more 
 requirements. In the following sections we will briefly describe the computational infrastructure at the
@@
 \subsubsection{Session Wizard}
+
+The initial attempt to ease the setup of Scipion projects for stream processing was based  
+on the original script developed at the CNB. But it was difficult to reuse the existing code due to a different folder
+structure and, more important, a different logic about users and microscopes management. To address the new set
+of requirements, a new ``wizard'' is being developed at SciLifeLab and can be found in the following open source repository: 
+\url{https://github.com/delarosatrevin/scipion-session}
+
+For this new wizard, a basic data model was designed representing the current processing and data 
+organization at SciLifeLab. The model is composed by the following entities:
+
+\begin{itemize}
+\setlength\itemsep{0em}
+ \item \textit{User}: represents all persons that can book microscope time or are registered in the Application Portal 
+ related to national projects.
+ \item \textit{Reservation}:  is a resource (e.g, Krios, Talos, Vitrobot, etc) booking for a given period of time.
+ \item \textit{Order}:  is a given national project with an unique identifier in the Application Portal
+ \item \textit{Project}:  can be a either a national project (related to a given Order) or an internal project.
+ \item \textit{Session}:  stores information about the usage of the micrograph for a given day.
+\end{itemize}
+
+When an user starts a data collection, the wizard is also executed using the command \textit{session-titan} 
+(or \textit{session-talos}). The wizard will fetch the reservations for this day using a webservices API provided by 
+the Booking System and will check wich user is operating the microscope. From the reservation information, 
+the wizard will also detect if it is an internal project or a national facility project. If the latter,
+more information will be retrieved from the Applicacion Portal (such as principal investigator, project code, 
+contact person or invoice address). After gathering all this information, the wizard setup the Scipion project
+for streaming processing. For example, the movies-import protocol will already point to the location where
+the files will be written for this session. Moreover, taking into account which microscope-camera is being 
+used, the files pattern can be inferred.
+
+The current implementation of the wizard stores each session in a simple database file. This database, together
+with extra information from the Booking System and the Appplication Portal, is used to generate reports for 
+a given period with the associated invoices. The generation of a session report is under development, that 
+will provide a summary and useful information for users to access the acquired data. The existing data model 
+represents the specify usage at the SciLifeLab facility and there are not inmediate plans to make it more 
+general and easy to reuse. Nonetheless, the model is quite simple and should to be difficult to modify it to fit 
+other requirements.

### 2018-02-28 2.overall.tex
Changes to Overall

@@
 
-In the following we describe the main modifications introduced in the original \scipion version in order to facilitate its deployment in centralized facilities.
+In the following we describe the main modifications made to \scipion in order to facilitate its deployment and functioning in centralized facilities.
 
@@
 For example, as movies are imported to the system they are aligned and their CTF is estimated.
-In \scipion stable version users may find  the first steps in the image processing adapted to stream.
-These steps include: movie alignment, CTF estimation and particle picking.
-In \scipion developers version  several 2D classification algorithms are available.
+In \scipion version 1.1, users may find  the first steps in the image processing adapted to streaming, including movie alignment and CTF estimation.
+We have also implemented particle picking and extraction, that are going to be released soon in the new version 1.2.
+In developers version, several 2D classification algorithms are being adapted for this new way of processing.
 
 \scipion stream implementation is based on periodic updates of the protocol input and output data during the image processing. 
-Since \scipion runs on threads, it is possible to devote one thread to update these information while the rest of the CPU power is devoted to run the different protocols.
-The use of threads marks an important difference between running a script or using \scipion, since  each module is run independently creating a very flexible processing framework.
-For example, do you want to add a new protocol in your workflow while it is running?
-No problem, do it.
+[**JMRT The following description is not accurate...Since \scipion runs on threads, it is possible to devote one thread to update these information while the rest of the CPU power is devoted to run the different protocols.
+The use of threads marks an important difference between running a script or using \scipion, since  each module is run independently creating a very flexible processing framework].
+[**JMRT The following sentence seems too colloquial compared with the other style...For example, do you want to add a new protocol in your workflow while it is running?
+No problem, do it].
 \scipion allows users to modify  workflows on the fly without modifying a single code line.
@@
 
-Since \scipion original publication we have developed many \emph{metaprotocols}, that is, 
+Since \scipion original publication we have developed many \emph{metaprotocols} [**JMRT not sure if the 
+introduction of the metaprotocols definition helps the reader...anyway, there are metaprotocols in the first
+version, so this is not completely new], that is, 
 protocols that either check the progress of other protocols or compare the results of equivalent protocols.
@@
 
-The second set of metaprotocols are grouped under the name \emph{consensus}.
-For a given logical step (for example, particle picking) these protocols check if the datasets obtained from the same input data using different algorithm are consistent.
+The second set of metaprotocols are grouped under the name \emph{consensus} [**JMRT Consesus protocol existed in the first release].
+For a given logical step (for example, particle picking) these protocols check if the datasets obtained from the same input data using a different algorithm are consistent.
 Continuing with the particle picking example, \emph{consensus picking} will compare the particles selected for each one of the executed particle-picking algorithms and produce as output a particle set containing those particles selected by most of the algorithms.
 
-Last but not least \scipion provides full data provenance.
+Despite the modifications introduced to \scipion data and process model for stream processing, the framework still provides full data provenance.
 Data provenance is an important form of metadata that describes how a particular data set was generated by detailing: the steps, input data and parameters used in the computation.
@@
 
-\scipion is a very flexible framework that allows to create many different workflows and to choose among several algorithms at each workflow-step.
+\scipion is a flexible framework that allows to create different workflows by choosing among many algorithms at each step.
 In this work we do not recommend a particular workflow  over others since we realize that different specimens, microscopes or facilities may have different requirements.
 Nevertheless, since a use case may clarify \scipion capabilities, in the following we describe the pipeline offered by default
-at our home facility.
+at the CNB.
 
@@
 The output of this protocol serves as verification of the validity of the experimental gain image for the whole data set.
-The second protocol performs Scipion direct detector movie alignment.
-Currently \scipion offers  four  methods  for  DDD  movie  alignment: motioncorr2 (Li et al., 2013), correlation (an unpublished CPU version of the motioncorr2 method developed in Xmipp), unblur (Grant and Grigorieff, 2015) and optical flow (Abrishami et al., 2015).
+The second protocol performs direct detector movie alignment.
+Currently \scipion offers  four  methods for movie alignment: motioncorr2 (Li et al., 2013), correlation (an unpublished CPU version of the motioncorr2 method developed in Xmipp [**JMRT this one is only global-alignment, so I will not say that is CPU version of motioncorr2), unblur (Grant and Grigorieff, 2015) and optical flow (Abrishami et al., 2015).
 Within \scipion, movie frames can be aligned using any one of these approaches or any combination.
-Following the standard processing workflow, the resulting micrographs are screened to eliminate those with excessive astigmatism or drift.
+Following the standard [**JMRT standard?
+recommend or not recommend?
+maybe common?] processing workflow, the resulting micrographs are screened to eliminate those with excessive astigmatism or drift [**JMRT This screen is not done after CTF?].
 Scipion provides three CTF estimation methods:
-CTFFIND (Rohou and Grigorieff, 2015), gCTF (Zhang, 2016) and Xmipp CTF estimation (Sorzano et al., 2007; Vargas et al., 2013).
+CTFFIND4 (Rohou and Grigorieff, 2015), gCTF (Zhang, 2016) and Xmipp CTF estimation (Sorzano et al., 2007; Vargas et al., 2013).
 At this point the user may compute the CTFs using several algorithms and compare them with the protocol \protocol{CTF consensus} that report up to which frequency two estimation of the same micrograph are equivalent.
@@
 The next step is to select (pick) particles.
-We have recently ported the eight different options available at \scipion  to streaming.
+We have recently adapted eight particle picking methods in \scipion  to streaming.
 In this way,  particles may be selected as soon as any new micrograph is available.
-We note here that some of the ported algorithms require a training step in which either user intervention or information from a previous selection is need.
+Some picking algorithms require a training step in which either user intervention or information from a previous selection is needed.
+The streaming functionality is separated from the picking logic, so if a new algorithm is added in the future it will work in streaming without extra effort from the develoepr.
 For those users that execute several picking algorithms, the protocol \protocol{picking consensus} may be useful.
@@
 
-The proposed pipeline offers quite flexibility to decide which particular algorithms will be used by a given user.
-At CNB (and collaborating institutions) we collect protocol usage statistics and sort the different algorithms by popularity.
+The proposed pipeline is very flexible to decide which particular algorithms will be used by a given user.
+At the CNB (and collaborating institutions) we collect protocol usage statistics and sort the different algorithms by popularity.
 This information is available at  \url{http://scipion.i2pc.es/report_protocols/protocolTable/} and may help users to select between the different algorithms.

### 2018-02-28 0.abstract.tex
Renamed workflow file and fixed Scipion reference

--- 2018-02-13 0.abstract.tex
@@
 
-Three dimensional electron microscopy (3DEM) is becoming a very data-intensive field in which vast amounts of experimental data are acquired at high speed.
-To manage such large-scale image data  projects, we have previously described  a modular workflow system called \scipion.
-We present here the first main extension of \scipion that allows to process data in streaming.
-Streaming processing allows to overlap data acquisition with the first steps of digital image processing and therefore helps to: (1) detect problems at early stages, (2) save computing time and (3) provide users with a in deep evaluation of the data quality before the data acquisition is finished.
-This idea of processing the data while being acquired has been implemented in different laboratories mainly using custom-made scripts.
-The advantages of our \scipion solution are: (1) flexibility: users can customize their workflows from a  large collection of preloaded algorithms; (2) repeatability and traceability: workflows are stored and can be re-executed later, either with the same data or a different data set; (3) monitorization: there are special utilities that constantly check how the execution of the algorithms is going and inform about possible anomalies; (4) reports: we have developed several GUIs that are updated periodically and produce a graphical summary (e.g, CTF defocus values, system load, etc).
-This summary may be  generated in HTML format that can be easily published though a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided.
-At present, \scipion has been installed and is in production mode in three facilities.
-And it is being considered in several more.
+Three dimensional electron microscopy (3DEM) is becoming a very data-intensive field in which vast amounts of experimental images are acquired at high speed.
+To manage such large-scale projects, we have previously developed  a modular workflow system called \scipion \citep{delarosatrevin2016}.
+We present here a major extension of \scipion that allows stream processing of EM images while the data is being acquired.
+This approach helps to detect problems at early stages, saves computing time and provides users with an in deep evaluation of the data quality before the acquisition is finished.
+This idea has been implemented in several laboratories by using custom-made scripts.
+The advantages of our proposed solution are: (1) flexibility: users can customize their workflows from a  large collection of preloaded algorithms; (2) repeatability and traceability: workflows are stored and can be re-executed later, either with the same data or a different data set; (3) monitorization: there are utilities that constantly check how the execution of the algorithms is going and inform about possible anomalies; (4) reports: we have developed several graphical interfaces that are updated periodically producing a summary (e.g, CTF defocus values, system load, etc).
+This summary may be generated in HTML format and easily published though a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided.
+At present, \scipion has been installed and is in production mode in several Cryo-EM facilities and it is being considered in many more.
 

### 2018-02-28 3.workflow.tex
Renamed workflow file and fixed Scipion reference

@@
 The data flow created by an electron microscope requires a careful design of the IT infrastructures.  Different facilities has adopted different solutions and therefore, 
-how \scipion  access data and interchange information is very facility dependent. In this section we discuss  the different solutions applied in the different facilities in which \scipion is running.
+[**JMRT I don't really like how this sentence sounds.
+It suggest that Scipion installations are very different among facilities and this is not True, even if the initial wizard is or how folders are structured] how \scipion  access data and interchange information is very facility dependent.
+In this section we discuss  the different solutions applied in the different facilities in which \scipion is running.
 
 So far, \scipion is being used routinely in four Cryo-EM facilities: \emph{Science for life Laboratory} (SciLifeLab, \url{http://www.scilifelab.se}),  \emph{European Synchrotron Radiation Facility} (ESRF, \url{http://www.esrf.eu/}), \emph{Diamond Light Source Synchrotron} (Diamond, \url{http://www.diamond.ac.uk}) and \emph{National Center for Biotechnology} (CNB, \url{http://www.cnb.csic.es}).
-\scipion original design satisfied the requirements of our home facility (CNB), in which \scipion automatically fetches newly recorded movies from a network mounted disk.
+\scipion original design satisfied the requirements of the CNB, in which \scipion automatically fetches newly recorded movies from a network mounted disk.
 After this first installation, \scipion was deployed on SciLifeLab (a centralized facility located at Stockholm) where a rigorous booking system is followed.
@@
 In this environment, \scipion needs to constantly interchange information with ISPyB, and it cannot longer check if there is a new file in a particular storage device but must wait to be informed by the system that a new file has been created.
-All these particularities, that will be described in detail in the following, could only be handled thanks to a careful design that prioritizes flexibility at many levels.
+All these particularities, that will be described in detail in the following, could only be handled thanks to a careful design that prioritizes flexibility at several levels.
 
@@
 \input{3.workflow_cnb.tex}
-\input{3.workflow_delarosa.tex}
+\input{3.workflow_scilifelab.tex}
 \input{3.workflow_diamond.tex}

### 2018-02-28 0.abstract.tex
modified line (see main body)
Different EM facilities have 
implemented their stream processing solution using custom-made scripts, but this script approach is not flexible and present a slow response time to new feature requests

@@
 This approach helps to detect problems at early stages, saves computing time and provides users with an in deep evaluation of the data quality before the acquisition is finished.
-This idea has been implemented in several laboratories by using custom-made scripts.
+Different EM facilities have
+implemented their stream processing solution using custom-made scripts, but this script approach is not flexible and present a slow response time to new feature requests.
 The advantages of our proposed solution are: (1) flexibility: users can customize their workflows from a  large collection of preloaded algorithms; (2) repeatability and traceability: workflows are stored and can be re-executed later, either with the same data or a different data set; (3) monitorization: there are utilities that constantly check how the execution of the algorithms is going and inform about possible anomalies; (4) reports: we have developed several graphical interfaces that are updated periodically producing a summary (e.g, CTF defocus values, system load, etc).
-This summary may be generated in HTML format and easily published though a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided.
+This summary may be generated in HTML format and  published though a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided.
 At present, \scipion has been installed and is in production mode in several Cryo-EM facilities and it is being considered in many more.

### 2018-02-28 2.overall.tex
change functioning by execution
Update 2.overall.tex

@@
 \label{overall}
+[**ROB this section goes not flow smoothly, needs a second pass]
 
-In the following we describe the main modifications made to \scipion in order to facilitate its deployment and functioning in centralized facilities.
+In the following we describe the main modifications made to \scipion in order to facilitate its deployment and execution in centralized facilities.
 
 The main change introduced in \scipion is \emph{stream processing}. Stream processing has been designed to analyze and react on real-time so that enables analysis of data as it is produced.
-In this way a new level of parallelization has been reached since the  data generated by the microscope continuously flows through a chain of protocols.
+In this way, a new level of parallelization has been reached since the  data generated by the microscope continuously flows through a chain of protocols.
 For example, as movies are imported to the system they are aligned and their CTF is estimated.
@@
 We have also implemented particle picking and extraction, that are going to be released soon in the new version 1.2.
-In developers version, several 2D classification algorithms are being adapted for this new way of processing.
+In the developers version, several 2D classification algorithms are being adapted for this new way of processing.
 
@@
 [**JMRT The following description is not accurate...Since \scipion runs on threads, it is possible to devote one thread to update these information while the rest of the CPU power is devoted to run the different protocols.
-The use of threads marks an important difference between running a script or using \scipion, since  each module is run independently creating a very flexible processing framework].
-[**JMRT The following sentence seems too colloquial compared with the other style...For example, do you want to add a new protocol in your workflow while it is running?
-No problem, do it].
-\scipion allows users to modify  workflows on the fly without modifying a single code line.
+The use of threads marks an important difference between running a script or using \scipion, since  each module is run independently creating a very flexible processing framework][**ROB: Any suggestion to make it more accurate?].
+\scipion allows users to modify the workflow on the fly without modifying a single code line.
 The modification will not interfere with the already executed steps and may take advantage from all already computed data.
-Did something go wrong and the microscope stopped?
-By default \scipion will stop after 24 hours with no microscope activity, but if you wish to continue your acquisition after a stop just press the continue button, there is no need to restart from the beginning.
+By default \scipion will automatically stop after 24 hours with no microscope activity, but you may  continue the acquisition after a stop by pressing the continue button, there is no need to restart from the beginning.
 
@@
 introduction of the metaprotocols definition helps the reader...anyway, there are metaprotocols in the first
-version, so this is not completely new], that is, 
+version, so this is not completely new][**ROB there is no claim that the concept of metaprotocol is new, The only claim is that many protocols of this kind have been introduced in this version], that is,
 protocols that either check the progress of other protocols or compare the results of equivalent protocols.
@@
 The first type of metaprotocols are called \emph{monitors} and are used to produce live analysis plots, generate reports or raise alerts when some problems are detected.
-A monitor example is the \protocol{CTF-monitor}, that checks the computed defocus values for
-each micrograph as they are generated.
+A monitor example is the \protocol{CTF-monitor}, that checks the computed defocus values for each micrograph as they are generated.
 CTF-monitor may raise an alert if the defocus values are above or below certain thresholds and continuously generate HTML files so that users and staff may easily follow the data acquisition and processing in-house or remotely.
@@
 
-The second set of metaprotocols are grouped under the name \emph{consensus} [**JMRT Consesus protocol existed in the first release].
+The second set of metaprotocols are grouped under the name \emph{consensus} [**JMRT Consesus protocol existed in the first release, **ROB Ok this is true I need to think about how to rewrite this].
 For a given logical step (for example, particle picking) these protocols check if the datasets obtained from the same input data using a different algorithm are consistent.
@@
 
-Despite the modifications introduced to \scipion data and process model for stream processing, the framework still provides full data provenance.
+Despite the modifications introduced to \scipion data and process model, the framework still provides full data provenance.
 Data provenance is an important form of metadata that describes how a particular data set was generated by detailing: the steps, input data and parameters used in the computation.
@@
 In this work we do not recommend a particular workflow  over others since we realize that different specimens, microscopes or facilities may have different requirements.
-Nevertheless, since a use case may clarify \scipion capabilities, in the following we describe the pipeline offered by default
-at the CNB.
+Nevertheless, since a use case may clarify \scipion capabilities, in the following we describe the pipeline offered by default at the CNB.
 
@@
 The second protocol performs direct detector movie alignment.
-Currently \scipion offers  four  methods for movie alignment: motioncorr2 (Li et al., 2013), correlation (an unpublished CPU version of the motioncorr2 method developed in Xmipp [**JMRT this one is only global-alignment, so I will not say that is CPU version of motioncorr2), unblur (Grant and Grigorieff, 2015) and optical flow (Abrishami et al., 2015).
-Within \scipion, movie frames can be aligned using any one of these approaches or any combination.
-Following the standard [**JMRT standard?
-recommend or not recommend?
-maybe common?] processing workflow, the resulting micrographs are screened to eliminate those with excessive astigmatism or drift [**JMRT This screen is not done after CTF?].
+Currently \scipion offers  four  methods for movie alignment: motioncorr2 (Li et al., 2013), correlation (an unpublished CPU version that implements the global aligment method used by motioncorr2), unblur (Grant and Grigorieff, 2015) and optical flow (Abrishami et al., 2015).
+Then the CTF of the resulting micrographs are calculated.
 Scipion provides three CTF estimation methods:
 CTFFIND4 (Rohou and Grigorieff, 2015), gCTF (Zhang, 2016) and Xmipp CTF estimation (Sorzano et al., 2007; Vargas et al., 2013).
-At this point the user may compute the CTFs using several algorithms and compare them with the protocol \protocol{CTF consensus} that report up to which frequency two estimation of the same micrograph are equivalent.
+At this point the user may compute the CTFs using several algorithms and compare them with the protocol \protocol{CTF consensus} that reports up to which frequency two estimation of the same micrograph are equivalent.
 Further automatic filtering based on CTF is available through the protocol \protocol{CTF filtering} that selects micrographs based on the astigmatism, resolution and defocus.
@@
 Some picking algorithms require a training step in which either user intervention or information from a previous selection is needed.
-The streaming functionality is separated from the picking logic, so if a new algorithm is added in the future it will work in streaming without extra effort from the develoepr.
+The streaming functionality is separated from the picking logic, so if in the future a new algorithm is added, it will work in streaming without extra effort from the algorithm developer.
 For those users that execute several picking algorithms, the protocol \protocol{picking consensus} may be useful.
@@
 \protocol{CTF monitor} plots the CTF defocus and astigmatism  and it will raise an alarm if these parameters are not within a given range.
-Finally the \protocol{report} protocol pushes all the created plots and summaries to a web server
-so this information can be accessed by users who are not physically in the microscope room.
+Finally the \protocol{report} protocol pushes all the created plots and summaries to a web server so this information can be accessed by users who are not physically in the microscope room.
 
-The proposed pipeline is very flexible to decide which particular algorithms will be used by a given user.
+The proposed pipeline is very flexible and can be easily modified by each user.
 At the CNB (and collaborating institutions) we collect protocol usage statistics and sort the different algorithms by popularity.

### 2018-02-28 3.workflow.tex
Update 3.workflow.tex

@@
-\section{\scipion Setup in the Different Facilities}
+\section{\scipion Setup at the Different Facilities}
 
-The data flow created by an electron microscope requires a careful design of the IT infrastructures.  Different facilities has adopted different solutions and therefore, 
-[**JMRT I don't really like how this sentence sounds.
-It suggest that Scipion installations are very different among facilities and this is not True, even if the initial wizard is or how folders are structured] how \scipion  access data and interchange information is very facility dependent.
+The data flow created by an electron microscope requires a careful design of the IT infrastructures.
+Different facilities has adopted different solutions and therefore, how \scipion  access data and interchange information may require minor adjustments.
 In this section we discuss  the different solutions applied in the different facilities in which \scipion is running.
@@
 There, \scipion was adapted to interface with this booking platform in order to produce better reports.
-Lately,  \scipion has been installed in a large synchrotron  such as ESRF and Diamond  where projects are handled by ISPyB (a customized laboratory management information system -LIMBS- \citep{Delageniere2011}) and data are saved on a distributed file system that heavily penalizes access to lists of files.
+Lately,  \scipion has been installed in  large synchrotrons  such as ESRF and Diamond  where projects are handled by ISPyB (a customized laboratory management information system -LIMBS- \citep{Delageniere2011}) and data are saved on a distributed file system that heavily penalizes access to lists of files.
 In this environment, \scipion needs to constantly interchange information with ISPyB, and it cannot longer check if there is a new file in a particular storage device but must wait to be informed by the system that a new file has been created.

### 2018-02-28 3.workflow_scilifelab.tex
Update 3.workflow_scilifelab.tex

@@
 The facilities can be accessed by Swedish researchers through a peer-reviewed routine. Applications can be 
-submitted through an Application Portal (reference???) and will be evaluated once every three months based on their 
+submitted through an Application Portal [**reference???] and will be evaluated once every three months based on their 
 scientific merit and technical feasibility by a national Project Evaluation Committee. On the other side, 
 some time of the facility is reserved for internal research groups at Stockholm University. Internally, the allocation
-of the microscope (and other instruments) time is managed by a Booking System (reference???).
+of the microscope (and other instruments) time is managed by a Booking System [** name of the booking system and reference???].
 
-At the beginning of 2016, the Cryo-EM facility at SciLifeLab became one of the early adopters of Scipion 
+At the beginning of 2016, the Cryo-EM facility at SciLifeLab became one of the early adopters of \scipion 
 for on the fly data processing. Before, there were some home-made scripts to perform motion correction (with motioncor)
@@
 \subsubsection{Infrastructure}
+[** to be done]
 
@@
 
-The initial attempt to ease the setup of Scipion projects for stream processing was based  
+The initial attempt to ease the setup of \scipion projects for stream processing was based  
 on the original script developed at the CNB. But it was difficult to reuse the existing code due to a different folder

### 2018-03-09 0.abstract.tex

@@
 Different EM facilities have
-implemented their stream processing solution using custom-made scripts, but this script approach is not flexible and present a slow response time to new feature requests.
+implemented their stream processing solution using custom-made scripts, but this script based approach is not flexible and present a slow response time to new feature requests.
 The advantages of our proposed solution are: (1) flexibility: users can customize their workflows from a  large collection of preloaded algorithms; (2) repeatability and traceability: workflows are stored and can be re-executed later, either with the same data or a different data set; (3) monitorization: there are utilities that constantly check how the execution of the algorithms is going and inform about possible anomalies; (4) reports: we have developed several graphical interfaces that are updated periodically producing a summary (e.g, CTF defocus values, system load, etc).

### 2018-03-09 1.introduction.tex

@@
 3D electron microscopy (3DEM)  can provide rich information about structural characteristics of macromolecular complexes.
-The field is under a major transformation due to the arrival of better microscopes, new large area detectors and automatization \citep{kuhlbrandt2014a, Kuhlbrandt2014b}  which makes possible that a single microscope generates high quality data sets on the order of terabytes per day \citep{Saibil2015} while working for several days without interruption.
-Major challenges faced in the field are: (1) an efficient management of these huge datasets and their corresponding image processing workflows; (2) a raising interest from scientists outside the field who might lack the skills of a experienced microscopist and; (3) the need for an effective transference of data and associated metadata in an increasingly distributed and collaborative environment.
-This situation places techniques that are automated and allow higher throughput in high demand.
+The field is under a major transformation due to the arrival of better microscopes, new large area detectors and automatization \citep{kuhlbrandt2014a, Kuhlbrandt2014b}.
+These improvements make possible that a single microscope generates high quality data sets on the order of terabytes per day \citep{Saibil2015} while working for several days without interruption.
+Major challenges faced in the field are: (1) an efficient management of these huge datasets and their corresponding image processing workflows; (2) a raising interest from scientists outside the field who might lack the skills of a experienced microscopist and; (3) the need for an effective data and metadata transfer in an increasingly distributed and collaborative environment.
+This situation strongly demands the application of techniques that are highly automated to allow high throughput.
 
@@
 Clearly, the future of projects that require high-end microscopes is through centralized microscopy facilities, in which automatic acquired data needs to be monitored by users and facility staff.
-\scipion has been redesigned with this scenario in mind in which the early detection of issues may save a data acquisition shift which otherwise needs to be repeated after a new application that may be granted many months later.
+[**JMRT VERY LARGE AND TWISTED SENTENCE \scipion has been redesigned with this scenario in mind in which the early detection of issues may save a data acquisition shift which otherwise needs to be repeated after a new application that may be granted many months later].
 In this way, \scipion starts the processing of the movies as they are being acquired with the double aim of checking for possible data collection errors and execute a first user-tailored image processing workflow that provides users with an accurate measurement of the acquired data quality.
@@
 Many software packages have been developed to provide automated data collection, such as SerialEM \citep{Mastronarde2005}, Leginon \citep{Suloway2009}, UCSF-Tomography \citep{Zheng2007}, Tom Toolbox \citep{Nickell2005}, EPU \citep{EPU}, Latitude in Digital Micrograph \citep{Latitude}, or EMMenu \citep{emmenu}, etc.
-However, the next step,  which is to provide image processing while  data are being collected, is not so well served.
+However, the next step,  which is to provide image processing while  data are being collected, is not so well addressed.
 In fact, in most places this is accomplished through home made scripts (i.e.

### 2018-03-19 2.overall.tex

@@
 \label{overall}
-[**ROB this section goes not flow smoothly, needs a second pass]
+[**ROB this section does not flow smoothly, needs a second pass]
 
@@
 \scipion stream implementation is based on periodic updates of the protocol input and output data during the image processing. 
-[**JMRT The following description is not accurate...Since \scipion runs on threads, it is possible to devote one thread to update these information while the rest of the CPU power is devoted to run the different protocols.
-The use of threads marks an important difference between running a script or using \scipion, since  each module is run independently creating a very flexible processing framework][**ROB: Any suggestion to make it more accurate?].
-\scipion allows users to modify the workflow on the fly without modifying a single code line.
-The modification will not interfere with the already executed steps and may take advantage from all already computed data.
+Each protocol run in Scipion is an independent process that can additionally create more processing threads.
+Each run that works in streaming mode has one of it threads monitoring input and updating output, that could overlap with the computing.
+The separation between the execution engine and the processing logic of each algorithm marks an important difference between running a script or using \scipion.The latter is a more general and extensible approach that allows users to modify the workflow on the fly without modifying a single code line.
+Moreover, it also allows to quickly extend the current functionality with new methods.
+
+%** JMRT: The following line is a more specify explanation, while the previous paragraph was more general
+%The modification will not interfere with the already executed steps and may take advantage from all already computed data.
 By default \scipion will automatically stop after 24 hours with no microscope activity, but you may  continue the acquisition after a stop by pressing the continue button, there is no need to restart from the beginning.
@@
 
-Since \scipion original publication we have developed many \emph{metaprotocols} [**JMRT not sure if the 
-introduction of the metaprotocols definition helps the reader...anyway, there are metaprotocols in the first
-version, so this is not completely new][**ROB there is no claim that the concept of metaprotocol is new, The only claim is that many protocols of this kind have been introduced in this version], that is,
+Since \scipion original publication we have developed many \emph{metaprotocols}, that is, 
 protocols that either check the progress of other protocols or compare the results of equivalent protocols.

### 2018-03-19 3.workflow_scilifelab.tex

@@
 \subsubsection{Infrastructure}
-[** to be done]
+At the heart of the SciLifeLab computational infrastructure is a storage and pre-processing server (the
+staging server). This machine has roughly 200 TB of storage (4 ZFS RAIDZ2 pools
+with 11 HDDs each), 2 NVIDIA GeForce GTX 1070 GPUs, 2 Intel Xeon E5-2630v4 CPUs
+(10 cores, 2.2 GHz, HT), 384 GB RAM, and a dual-port 10 GbE network card. The
+storage pool is exported via NFS and Samba, so the microscope computers can
+write data there directly and other machines can access it.
+
+Unfortunately the K2 computer writes data to a local SSD RAID, the Falcon-III to
+a dedicated storage server (the offload server), in order to avoid interruption
+of data collection due to network issues. So when users start data collection,
+they start a script that continously moves the produced data to the staging
+server.
+
+By having all data on the staging server, in an organized folder structure, we
+can run pre-processing directly on the server. If processing is started soon
+after data collection the incoming file data should still be in RAM and does not
+have to be read from disk, which is important with performance in mind. Users
+can copy their data to external USB drives from two workstations in the
+microscope room, that have access to the storage pool via NFS.
+There is also a dedicated download server with access to the NFS export,
+reachable from the Internet. Here users can log in to a restricted shell that
+allows downloads via rsync (over SSH) or SFTP.
+
+All these computers are conntected to a 10 GbE network switch. This is a private
+network without external access. The only machine accessible from the Internet
+is the data download server, which has a dual-port 10 GbE network adapter (one
+used for the private network, one for the external network). Both the K2
+computer and the Falcon-III offload server had one unused 10GbE interface that
+could be used to connect them to our private network, without having to
+interfere with the default setup.
 

### 2018-03-21 1.introduction.tex
w

@@
 Major challenges faced in the field are: (1) an efficient management of these huge datasets and their corresponding image processing workflows; (2) a raising interest from scientists outside the field who might lack the skills of a experienced microscopist and; (3) the need for an effective data and metadata transfer in an increasingly distributed and collaborative environment.
-This situation strongly demands the application of techniques that are highly automated to allow high throughput.
+%This situation strongly demands the application of techniques that are highly automated to allow high throughput.
 
@@
 Clearly, the future of projects that require high-end microscopes is through centralized microscopy facilities, in which automatic acquired data needs to be monitored by users and facility staff.
-[**JMRT VERY LARGE AND TWISTED SENTENCE \scipion has been redesigned with this scenario in mind in which the early detection of issues may save a data acquisition shift which otherwise needs to be repeated after a new application that may be granted many months later].
+\scipion has been redesigned with this scenario in mind.
+% in which the early detection of issues may save a data acquisition shift which otherwise needs to be repeated after a new application that may be granted many months later.
 In this way, \scipion starts the processing of the movies as they are being acquired with the double aim of checking for possible data collection errors and execute a first user-tailored image processing workflow that provides users with an accurate measurement of the acquired data quality.
 
-Many software packages have been developed to provide automated data collection, such as SerialEM \citep{Mastronarde2005}, Leginon \citep{Suloway2009}, UCSF-Tomography \citep{Zheng2007}, Tom Toolbox \citep{Nickell2005}, EPU \citep{EPU}, Latitude in Digital Micrograph \citep{Latitude}, or EMMenu \citep{emmenu}, etc.
+Many software packages have been developed to provide automated data collection, such as SerialEM \citep{Mastronarde2005}, Leginon \citep{Suloway2009}, UCSF-Tomography \citep{Zheng2007}, Tom Toolbox \citep{Nickell2005}, EPU \citep{EPU}, Latitude in Digital Micrograph \citep{Latitude}, EMMenu \citep{emmenu}, etc.
 However, the next step,  which is to provide image processing while  data are being collected, is not so well addressed.
-In fact, in most places this is accomplished through home made scripts (i.e.
-\citep{Pichkur2018}).
+In fact, in most places this is accomplished through home made scripts \citep[i.e.][]{Pichkur2018}.
+%(i.e.
+\citet{Pichkur2018}).
 Exceptions are UCSFImage4 \citep{Li2015} and Focus \citep{Biyani2017}.

### 2018-03-21 2.overall.tex
w

@@
-\section{Main Modifications introduced in  \scipion: Streaming, Monitors and Consensus}
+\section{Main Modifications introduced in  \scipion: Streaming and Metaprotocols}
 \label{overall}
-[**ROB this section does not flow smoothly, needs a second pass]
 
-In the following we describe the main modifications made to \scipion in order to facilitate its deployment and execution in centralized facilities.
+In the following we describe the most important modifications made to \scipion in order to facilitate its deployment and execution in centralized facilities.
 
-The main change introduced in \scipion is \emph{stream processing}. Stream processing has been designed to analyze and react on real-time so that enables analysis of data as it is produced.
+\subsection{Stream Processing}
+The main change introduced in \scipion is \emph{stream processing}, that is, computing on data directly as it is produced or received.
 In this way, a new level of parallelization has been reached since the  data generated by the microscope continuously flows through a chain of protocols.
@@
 
-\scipion stream implementation is based on periodic updates of the protocol input and output data during the image processing. 
-Each protocol run in Scipion is an independent process that can additionally create more processing threads.
-Each run that works in streaming mode has one of it threads monitoring input and updating output, that could overlap with the computing.
-The separation between the execution engine and the processing logic of each algorithm marks an important difference between running a script or using \scipion.The latter is a more general and extensible approach that allows users to modify the workflow on the fly without modifying a single code line.
-Moreover, it also allows to quickly extend the current functionality with new methods.
+%\scipion is a multilayered application, the first layer handles the interaction with the user (\emph{presentation layer}), the second saves and retrieves the data used by the application  (\emph{data layer}) and finally, the third layer (\emph{business layer}) process the data following the user instructions.
+\scipion stream implementation make extensive use of threads. %A thread is the smallest unit of processing that can be scheduled by an operating system. 
+Using threads, a programmer can effectively create two or more tasks that run at the same time and share the protocol resources (i.e. memory). 
+%is based on periodic updates of the protocol input and output data during the image processing. 
+%Each protocol execution in \scipion is an independent process that can create processing threads. 
+Each protocol that works in streaming mode has one of its threads monitoring input and updating output  while other threads take care of the data processing.
+This clear separation between data access and data processing marks an important difference between running a script or using \scipion.
+Since the later facilitates a form of computing in which several computations are executed  concurrently, instead of sequentially.
+Moreover, \scipion workflow engine, allows users to modify the workflow on the fly without modifying a single code line.
+%In practice, the latter follows  a more general and extensible approach that allows users to modify the workflow on the fly without modifying a single code line.
+%Moreover, it also allows to quickly develop new workflows .
+
+%Since allow to a form of computing in which several computations are executed during overlapping time periods—concurrently—instead of sequentially
 
@@
 
+\subsection{Metaprotocols}
 Since \scipion original publication we have developed many \emph{metaprotocols}, that is, 
@@
 CTF-monitor may raise an alert if the defocus values are above or below certain thresholds and continuously generate HTML files so that users and staff may easily follow the data acquisition and processing in-house or remotely.
-An example of this reports may be seen at \url{http://nolan.cnb.csic.es/scipionbox/} [**ROB NOTE: select a nice one an complete URL]%At the project's end, a review of the processing history is produced.
+An example of this reports may be seen at \url{http://nolan.cnb.csic.es/scipionbox/} [**ROB NOTE: select a nice one an complete URL, I need to ask to javi chichon]%At the project's end, a review of the processing history is produced.
 [** URL to report example?]
@@
 
-The second set of metaprotocols are grouped under the name \emph{consensus} [**JMRT Consesus protocol existed in the first release, **ROB Ok this is true I need to think about how to rewrite this].
+The second set of metaprotocols are grouped under the name \emph{consensus}.
 For a given logical step (for example, particle picking) these protocols check if the datasets obtained from the same input data using a different algorithm are consistent.
@@
 
-Despite the modifications introduced to \scipion data and process model, the framework still provides full data provenance.
-Data provenance is an important form of metadata that describes how a particular data set was generated by detailing: the steps, input data and parameters used in the computation.
+We would like to finish this section assuring that despite the modifications introduced to \scipion data and process model, the framework provides full data provenance.
+That is, for any object created by a  protocol, \scipion always annotate how this object has been generated by detailing: the steps, input data and parameters used in the computation.
 Data provenance guaranties reproducibility and traceability.

### 2018-03-21 3.workflow.tex
w

@@
 
-So far, \scipion is being used routinely in four Cryo-EM facilities: \emph{Science for life Laboratory} (SciLifeLab, \url{http://www.scilifelab.se}),  \emph{European Synchrotron Radiation Facility} (ESRF, \url{http://www.esrf.eu/}), \emph{Diamond Light Source Synchrotron} (Diamond, \url{http://www.diamond.ac.uk}) and \emph{National Center for Biotechnology} (CNB, \url{http://www.cnb.csic.es}).
-\scipion original design satisfied the requirements of the CNB, in which \scipion automatically fetches newly recorded movies from a network mounted disk.
+So far, \scipion is being used routinely in four Cryo-EM facilities: \emph{Science for life Laboratory} (SciLifeLab, \url{http://www.scilifelab.se}),  \emph{European Synchrotron Radiation Facility} (ESRF, \url{http://www.esrf.eu/}), \emph{eBIC facility at Diamond Light Source} (Diamond, \url{http://www.diamond.ac.uk}) and \emph{National Center for Biotechnology} (CNB, \url{http://www.cnb.csic.es}).[**update] \scipion original design satisfied the requirements of the CNB, in which \scipion automatically fetches newly recorded movies from a network mounted disk.
 After this first installation, \scipion was deployed on SciLifeLab (a centralized facility located at Stockholm) where a rigorous booking system is followed.

### 2018-03-21 3.workflow_cnb.tex
w

@@
 \subsection{CNB}
+The National Center for Biotechnology (CNB) forms part of the Spanish National Research Council(CSIC), the largest public research institution in Spain. 
+The CNB Cryo-Electron Microscopy Facility is a joint effort of the Centro Nacional de Biotecnología (CNB-CSIC) and the Centro de Investigaciones Biológicas (CIB-CSIC).
+The mission of the facility is to provide researchers access to the latest cryo-EM technology for high resolution imaging.
 
@@
 In the first one we discuss the network setup that allow us to move the data from the microscope to the user laboratory.
-Then  we comment on an additional software (called  \emadmin) that connects \scipion with the microscope acquisition
-software (EPU) and records the facility activity.
+Then  we comment on an additional software (called  \emadmin) that connects \scipion with the microscope acquisition software (EPU) and records the facility activity.
 

### 2018-03-21 3.workflow_diamond.tex
w

@@
-\subsection{Diamond Light Source Facility??? or eBIC???}
+\subsection{The eBIC facility at Diamond Light Source}
 
-\begin{verbatim}
+Diamond Light Source is the UK National Synchrotron User Facility.
+As such, it provides access to large-scale experimental facilities to the UK and worldwide scientific communities to conduct experiments that are not possible in their home laboratories.
+As a consequence of this national role, Diamond has significant experience at managing many visiting scientists and user experiments, and the associated data and post-processing.
+%In the area of Macromolecular Crystallography, this has become a highly automated process which allows users to visualise electron density maps within minutes of the experiment being completed (REF.
+Aller P et al Methods Mol Biol 2015)
+%The recent ``Revolution Resolution'' (REF.
+Kulbrandt W, Science 2014) in cryoEM has transformed the impact and demand of cryoEM in Structural Biology.
+However, the instrumentation and running costs of the latest generation of EM necessary for high-end cryoEM is beyond scope of all but the best-funded home laboratories.
+Due to the huge demand for high-end microscope access,
+The electron Bio-Imaging Centre (eBIC) was set up at Diamond~\citep{diamond2017}.
+The centre houses four Thermofisher Titan Krios, a Talos Arctica and a Scios dual-beam.
+All of the transmission electron microscopes are equipped with counting-mode direct electron detectors and Volta phase plates.
 
-We need to cover here differences with CNB setup
+Diamond has thousands of users every year, both academic and commercial, and it is vital that their data is kept secure and confidential.
+A user management system has been developed which allocates each user with a unique federal identity (IDs), and associates those user IDs with the experiments in which they partake in.
+Visits are structured such that all raw and processed data collected during an experiment are kept in one directory, and only the associated visitors have read and write permissions in that directory.
+Users may access their data through ISPyB \citep{ispb2011},  a Laboratory Information Management System (LIMS) combining sample tracking and experiment reporting  %The link between the user and their visits enables data management such as archiving data to a long-term tape storage system, or the Laboratory Information Management System ISPyB (REF.
+Delageniere et al Bioinformatics 2011).
+In contrast to the CNB setup, Diamond has integrated the microscopes in much the same fashion to that of the beamlines at the facility, as is shown in figure X
 
- DEVELOPER RELATED
- * Describe infrastructure in diamond: file system (distributed, user permissions),
-                                       no graphical output, queues, 
- * Describe Scipion modification:      import (detect new files by socket message)
-                                       monitor, report to ISPyB 
-                                       scipion headless
-                                       execution as script
+Data is collected onto a local machine %in a similar way to the CNB setup, but 
+and then immediately moved via 10Gbps Ethernet to the associated visit directory in the centralised storage, a multi Petabyte GPFS High Performance Parallel file system.
+Once the raw data is present on the GPFS system, high speed interconnects to the central cluster enable rapid data processing.
+%Services such as these are shared amongst the whole facility allowing for flexibility in their deployment and generally allowing better service.
 
-USER RELATED
- * how users interact with scipion?
- * how data is visualized?
-\end{verbatim}
+Once the data has reached the central file system, it is available to be processed by Scipion.
+This process is facilitated by the use of a Scipion-workflow-template which enable novice users to trigger semi-customised on the fly processing.
+This template consists of a small UI (figure XX) which allows the user to put in the specific values pertinent to their experiment, namely the session ID (or visit number), which microscope they are using, and several other key data collection parameters.
+Once set this makes use of a global template for Scipion as set up by the eBIC team, and substitutes the appropriate information into it, and then saves it to disk.
+
+Following this Diamond makes use of a central ActiveMQ [** what is this?
+a queue system] service which is triggered upon the template file being available in the visit directory, and runs a headless (i.e.
+without UI) version of Scipion which takes the template and executes in streaming mode.
+THIS SHOULD PROBABLY INCLUDE SOMETHING ABOUT THE DETECTION OF NEW FILES BY SOCKET MESSAGE, BUT I DON’T KNOW HOW THIS WORKS.
+Me neither, I thought it polled the filesystem?
+
+To monitor the progress of the automatic processes, the Scipion pipeline at Diamond also includes custom nodes which allow it to interface with ISPyB.
+This provides the users with a web interface showing general metadata about their experiment as well as new real-time updates on steps of the process and relevant output such as motion corrected images, drift values and CTF images (Figure XXX)

### 2018-03-21 scipionbox.tex
W
w

@@
 \def\scipion{\textit{Scipion}\xspace}
+\def\emadmin{\textit{EMAdmin}\xspace}
 \newcommand{\protocol}[1]{\texttt{#1}}
@@
 \begin{document}
-\title{ScipionBox: automatic processing for }
-\author{XXXXX XXXXX$^{(a)}$, 
-        YYYYY YYYYY$^{(b)}$, \\
-        zzzzzz$^{(m)}$,
-        yyyyyyy$^{(b)}$
+\title{Using Scipion for Stream Image Processing at EM Facilities}
+\author{XXXX$^{(a)}$,\\
+        YYYY$^{(b)}$,\\
         }
@@
 
-%\lyxaddress{{\small \vspace{-0.5cm}$^{(a)}$Escuela Polit\'{e}cnica Superior, Universidad Aut\'{o}noma de Madrid, 28049 Cantoblanco, Madrid, Spain.}}
-%\lyxaddress{{\small$^{(g)}$ Biocomputing Unit, National Center for Biotechnology (CSIC), C/ Darwin, 3, Campus Universidad Aut\'{o}noma, 28049 Cantoblanco,
-%Madrid, Spain.}}
+\lyxaddress{{\small \vspace{-0.5cm}$^{(a)}$Escuela Polit\'{e}cnica Superior, Universidad Aut\'{o}noma de Madrid, 28049 Cantoblanco, Madrid, Spain.}}
+\lyxaddress{{\small$^{(b)}$ Biocomputing Unit, National Center for Biotechnology (CSIC), C/ Darwin, 3, Campus Universidad Aut\'{o}noma, 28049 Cantoblanco, Madrid, Spain.}}
 
@@
 \newpage
-\input{abstract.tex}
-\input{introduction.tex}
-\input{overall.tex}
-\input{workflow.tex}
-\input{results.tex}
-\input{conclusions.tex}
-\input{acknowledgments.tex}
+\input{0.abstract.tex}
+\input{1.introduction.tex}
+\input{2.overall.tex}
+\input{3.workflow.tex}
+\input{4.results.tex}
+\input{5.conclusions.tex}
+\input{6.acknowledgments.tex}
 \appendix

### 2018-03-21 2.overall.tex
casa

@@
 Each protocol that works in streaming mode has one of its threads monitoring input and updating output  while other threads take care of the data processing.
-This clear separation between data access and data processing marks an important difference between running a script or using \scipion.
-Since the later facilitates a form of computing in which several computations are executed  concurrently, instead of sequentially.
+This clear separation between data access and data processing marks an important difference between running a script or using \scipion since it facilitates a form of computing in which several computations are executed  concurrently, instead of sequentially.
 Moreover, \scipion workflow engine, allows users to modify the workflow on the fly without modifying a single code line.
@@
 A monitor example is the \protocol{CTF-monitor}, that checks the computed defocus values for each micrograph as they are generated.
-CTF-monitor may raise an alert if the defocus values are above or below certain thresholds and continuously generate HTML files so that users and staff may easily follow the data acquisition and processing in-house or remotely.
-An example of this reports may be seen at \url{http://nolan.cnb.csic.es/scipionbox/} [**ROB NOTE: select a nice one an complete URL, I need to ask to javi chichon]%At the project's end, a review of the processing history is produced.
+\protocol{CTF-monitor} may raise an alert if the defocus values are above or below certain thresholds and continuously generate HTML files so that users and staff may easily follow the data acquisition and processing either in-house or remotely.
+An example of these reports may be seen at \url{http://nolan.cnb.csic.es/scipionbox/} [**ROB NOTE: select a nice one an complete URL, I need to ask to javi chichon]%At the project's end, a review of the processing history is produced.
 [** URL to report example?]
@@
 For a given logical step (for example, particle picking) these protocols check if the datasets obtained from the same input data using a different algorithm are consistent.
-Continuing with the particle picking example, \emph{consensus picking} will compare the particles selected for each one of the executed particle-picking algorithms and produce as output a particle set containing those particles selected by most of the algorithms.
+Continuing with the particle picking example, \protocol{consensus picking} will compare the particles selected for each one of the executed particle-picking algorithms and produce as output a particle set containing those particles selected by most of the algorithms.
 
 We would like to finish this section assuring that despite the modifications introduced to \scipion data and process model, the framework provides full data provenance.
-That is, for any object created by a  protocol, \scipion always annotate how this object has been generated by detailing: the steps, input data and parameters used in the computation.
+That is, \scipion provides detailed information about the origin of the data XXXXX
+
+ easily share traceable information and research results is central to their use.
+
+in such a way  ensuring reproducibility in scientific research
+
+that data will be traceable and reproducible
+
+keeps a historical record of how the data was generated reproducibility and traceability.
+
+%for any object created by a  protocol, \scipion always annotate how this object has been generated by detailing: the steps, input data and parameters used in the computation.
 Data provenance guaranties reproducibility and traceability.

### 2018-03-21 1.introduction.tex
casa

@@
 As we will show in this article, \scipion differs from these solutions in a larger flexibility both from the point of view of users and facility staff.
+
+The organization of this document is as follows.
+First, we present the modifications introduced in \scipion to adapt it to large centralized microscopy facilities.
+Then we present a typical workflow to demostrate \scipion capabilities.
+Finally,  we describe in detail how \scipion is being used in three representative facilities.

### 2018-03-21 2.overall.tex
casa

@@
 For example, as movies are imported to the system they are aligned and their CTF is estimated.
-In \scipion version 1.1, users may find  the first steps in the image processing adapted to streaming, including movie alignment and CTF estimation.
+In \scipion version 1.1, users may find  the first steps in the image processing workflow adapted to streaming, including movie alignment and CTF estimation.
 We have also implemented particle picking and extraction, that are going to be released soon in the new version 1.2.
@@
 %Each protocol execution in \scipion is an independent process that can create processing threads. 
-Each protocol that works in streaming mode has one of its threads monitoring input and updating output  while other threads take care of the data processing.
-This clear separation between data access and data processing marks an important difference between running a script or using \scipion since it facilitates a form of computing in which several computations are executed  concurrently, instead of sequentially.
+Each protocol that works in streaming mode, has one of its threads monitoring input and updating output  data while other threads take care of the data processing.
+This clear separation between data access and data processing marks an important difference between running a script or using \scipion since it facilitates a form of computing in which several jobs are executed  concurrently, instead of sequentially.
 Moreover, \scipion workflow engine, allows users to modify the workflow on the fly without modifying a single code line.
@@
 We would like to finish this section assuring that despite the modifications introduced to \scipion data and process model, the framework provides full data provenance.
-That is, \scipion provides detailed information about the origin of the data XXXXX
-
- easily share traceable information and research results is central to their use.
-
-in such a way  ensuring reproducibility in scientific research
-
-that data will be traceable and reproducible
-
-keeps a historical record of how the data was generated reproducibility and traceability.
+That is, \scipion provides detailed information about the origin of the data  and the process to generate it.
 
@@
 In this work we do not recommend a particular workflow  over others since we realize that different specimens, microscopes or facilities may have different requirements.
-Nevertheless, since a use case may clarify \scipion capabilities, in the following we describe the pipeline offered by default at the CNB.
+Nevertheless, since a use case may clarify \scipion capabilities, in the following we describe the pipeline offered by default at the CNB cryoEM Service.
 
@@
 By default \protocol{system}, \protocol{CTF} and \protocol{report monitors} are executed.
-\protocol{System monitor} reports on  CPU and GPU utilization, memory consumption and, disk I/O.
+\protocol{System monitor} reports on  CPU and GPU usage, memory consumption and, disk I/O.
 \protocol{CTF monitor} plots the CTF defocus and astigmatism  and it will raise an alarm if these parameters are not within a given range.

### 2018-03-21 3.workflow.tex
casa

--- 2018-02-28 3.workflow.tex
@@
 
-So far, \scipion is being used routinely in four Cryo-EM facilities: \emph{Science for life Laboratory} (SciLifeLab, \url{http://www.scilifelab.se}),  \emph{European Synchrotron Radiation Facility} (ESRF, \url{http://www.esrf.eu/}), \emph{Diamond Light Source Synchrotron} (Diamond, \url{http://www.diamond.ac.uk}) and \emph{National Center for Biotechnology} (CNB, \url{http://www.cnb.csic.es}).
+So far, \scipion is being used routinely in seven Cryo-EM facilities:
+\begin{itemize}
+ \itemsep0em
+
+ \item \emph{Ume\aa\ Core Facility for Electron Microscopy} (UCEM, \url{http://www.kbc.umu.se/english/ucem})[** Jose Miguel can you verify this?] 
+ \item \emph{Science for life Laboratory} (SciLifeLab, \url{http://www.scilifelab.se})
+ \item \emph{ESRF Cryo-Electron Microscope} (\url{http://www.esrf.eu/home/UsersAndScience/Experiments/MX/About_our_beamlines/CM01.html})
+ \item \emph{Electron Bio-Imaging Centre at Diamond Light Source} (eBIC, \url{http://www.diamond.ac.uk/Science/Integrated-facilities/eBIC.html}) and
+ \item \emph{CNB CryoEM Service} (\url{http://www.cnb.csic.es/index.php/en/research/core-facilities/microscopia-crioelectronica})
+ \item \emph{Molecular Microscopy Consortium} (\url{https://www.niehs.nih.gov/research/atniehs/facilities/mmc/index.cfm})[**Pablo can you confirm this?]
+ \item \emph{National Cancer Institute - NIH} (\url{https://cmm.nci.nih.gov/})[**Pablo can you confirm this?]
+
+\end{itemize}
+
 \scipion original design satisfied the requirements of the CNB, in which \scipion automatically fetches newly recorded movies from a network mounted disk.
@@
 
+SAY WE ARE GOING TO DESCRIBE THREE REPRESENTAIVE CENTER
 %\scipion is a very flexible framework that allows to create many different workflows and to choose among several algorithms at each workflow-step.

### 2018-03-21 3.workflow.tex
casa

@@
  \item \emph{ESRF Cryo-Electron Microscope} (\url{http://www.esrf.eu/home/UsersAndScience/Experiments/MX/About_our_beamlines/CM01.html})
- \item \emph{Electron Bio-Imaging Centre at Diamond Light Source} (eBIC, \url{http://www.diamond.ac.uk/Science/Integrated-facilities/eBIC.html}) and
+ \item \emph{Electron Bio-Imaging Centre at Diamond Light Source} (eBIC, \url{http://www.diamond.ac.uk/Science/Integrated-facilities/eBIC.html})
  \item \emph{CNB CryoEM Service} (\url{http://www.cnb.csic.es/index.php/en/research/core-facilities/microscopia-crioelectronica})
@@
 
-SAY WE ARE GOING TO DESCRIBE THREE REPRESENTAIVE CENTER
 %\scipion is a very flexible framework that allows to create many different workflows and to choose among several algorithms at each workflow-step.

### 2018-03-22 2.overall.tex
before class

@@
 In this work we do not recommend a particular workflow  over others since we realize that different specimens, microscopes or facilities may have different requirements.
-Nevertheless, since a use case may clarify \scipion capabilities, in the following we describe the pipeline offered by default at the CNB cryoEM Service.
+Nevertheless, since a use case may clarify \scipion capabilities, in the following we describe the pipeline offered by default at the \emph{Centro Nacional de Biotecnología} (CNB) cryoEM Service.
 

### 2018-03-22 3.workflow.tex
before class

@@
 
- \item \emph{Ume\aa\ Core Facility for Electron Microscopy} (UCEM, \url{http://www.kbc.umu.se/english/ucem})[** Jose Miguel can you verify this?] 
- \item \emph{Science for life Laboratory} (SciLifeLab, \url{http://www.scilifelab.se})
+ %\item \emph{Ume\aa\ Core Facility for Electron Microscopy} (UCEM, \url{http://www.kbc.umu.se/english/ucem})[** Jose Miguel can you verify this?] 
+ \item \emph{Science for life Laboratory} (SciLifeLab, \url{http://www.scilifelab.se}). Both at Stockholm and Ume\aa\ facilities)
  \item \emph{ESRF Cryo-Electron Microscope} (\url{http://www.esrf.eu/home/UsersAndScience/Experiments/MX/About_our_beamlines/CM01.html})
- \item \emph{Electron Bio-Imaging Centre at Diamond Light Source} (eBIC, \url{http://www.diamond.ac.uk/Science/Integrated-facilities/eBIC.html})
+ \item \emph{eBIC at Diamond Light Source} ( \url{http://www.diamond.ac.uk/Science/Integrated-facilities/eBIC.html})
  \item \emph{CNB CryoEM Service} (\url{http://www.cnb.csic.es/index.php/en/research/core-facilities/microscopia-crioelectronica})
- \item \emph{Molecular Microscopy Consortium} (\url{https://www.niehs.nih.gov/research/atniehs/facilities/mmc/index.cfm})[**Pablo can you confirm this?]
+ \item \emph{Molecular Microscopy Consortium - NIH} (\url{https://www.niehs.nih.gov/research/atniehs/facilities/mmc/index.cfm})[**Pablo can you confirm this?]
  \item \emph{National Cancer Institute - NIH} (\url{https://cmm.nci.nih.gov/})[**Pablo can you confirm this?]
@@
 \end{itemize}
+and it is being considered in many more.
 
@@
 In this environment, \scipion needs to constantly interchange information with ISPyB, and it cannot longer check if there is a new file in a particular storage device but must wait to be informed by the system that a new file has been created.
-All these particularities, that will be described in detail in the following, could only be handled thanks to a careful design that prioritizes flexibility at several levels.
+
+In the following, through three use cases, we describe in detail how \scipion has been adapted to the different environments.
+
+%All these particularities, that will be described in detail in the following, could only be handled thanks to a careful design that prioritizes flexibility at several levels.
 
@@
 
-\subsection{What will do in future}
+\subsection{\scipion near future}
 %Particle picking
@@
 %More complex reports
-[** this section may disappear]
+plug-in style, independend

### 2018-03-22 3.workflow_cnb.tex
before class

--- 2018-02-20 3.workflow_cnb.tex
@@
-\subsection{CNB}
+\subsection{Case 1: Scipion at CNB}
+The National Center for Biotechnology (CNB) forms part of the Spanish National Research Council (CSIC), the largest public research institution in Spain. 
+The CNB Cryo-Electron Microscopy Facility is a joint effort of the Centro Nacional de Biotecnolog\'{i}a (CNB-CSIC) and the Centro de Investigaciones Biol\'{o}gicas (CIB-CSIC).
+The mission of the facility is to provide researchers access to the latest cryo-EM technology for high resolution imaging.
 
@@
 In the first one we discuss the network setup that allow us to move the data from the microscope to the user laboratory.
-Then  we comment on an additional software (called  \emadmin) that connects \scipion with the microscope acquisition
-software (EPU) and records the facility activity.
+Then  we comment on an additional software (called  \emadmin) that connects \scipion with the microscope acquisition software (\epu) and records the facility activity.
 
-\subsubsection{Network Setup}
+\subsubsection{Network Setup and IT Infrastructure}
 
@@
 
-Data sets are recorded using EPU.
+Data sets are recorded using \epu.
 The microscope control system has a 70 TB NAS (Network-attached storage) unit for saving movies as they are obtained (we will refer to this NAS as microscope-NAS).
@@
 The data flow is a follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox will temporary read the movies from 
-the microscope-NAS but will not store them locally, (c) each hour, output-NAS will connect to \scipionbox and copy or update the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup the \scipion project and the movies to a USB disk or serve this information through the Internet and, (e) in parallel, \scipionbox transfers to  output-NAS different html reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox}).
+the microscope-NAS but will not store them locally, (c) each hour, output-NAS will connect to \scipionbox and copy -or update- the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup \scipion project and the movies to a USB disk or serve this information through the Internet and, (e) in parallel, \scipionbox transfers to  output-NAS different html reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox}).
 
-Although we try to discourage the use of external USB hard disks versus network as way of retrieving the raw and processing data, this is the preferred option for more of our users.
-USB 3.0 drives can theoretically handle data flows of 300 MB/s, but our experience is that, in this busy environment, many of them  have trouble exceeding 50 MB/s.
+Although CNB staff try to discourage the use of external USB hard disks versus network as way of retrieving the raw and processing data, this is the preferred option for most users.
+USB 3.0 drives can theoretically handle data flows of 300 MB/s, but CNB experience is that, in this busy environment, many of them  have trouble exceeding 50 MB/s.
 In most cases, 50 MB/s is enough to transfer the data before the following microscope shift starts.
@@
 
-\emadmin was born as a script that created a normalized tree of directories  in which to store the movies acqured by the microscope and help users to choose between a few predetermined image processing workflows.
+\emadmin was born as a script that created a normalized tree of directories  in which to store the movies acquired by the microscope, helped users to choose between a few predetermined image processing workflows and finally launched \scipion.
 At present it was grown in a client-server application that in addition to the old script capacities it has incorporate the following functions:
@@
 \emadmin is free software available at github repository (\url{https://github.com/rmarabini/webservices/tree/master/EMadmin}).
-You are welcomed to try it but the software is provided as it is and no support or maintenance service is offered.
+You are welcomed to try it but be aware that \emadmin is designed to be used at CNB and therefore, no support is provided for it.

### 2018-03-22 3.workflow_scilifelab.tex
before class

@@
-\subsection{Swedish National CryoEM Facility}
+\subsection{Case 2: Scipion at SciLifeLab}
 
-The Swedish National Cryo-EM Facility offers access to state-of-the-art equipment and expertise 
-in single particle cryo-EM and cryo electron tomography (cryo-ET). The Facility has two nodes: at SciLifeLab 
-in Stockholm and at Umeå University. SciLifeLab in Stockholm offers single-particle service with a 
+The Swedish National Cryo-EM Facility offers access to state-of-the-art equipment and expertise in single particle cryo-EM and cryo electron tomography (cryo-ET).
+The Facility has two nodes: at SciLifeLab in Stockholm and at Ume\aa\ University.
+SciLifeLab in Stockholm offers single-particle service with a
 Talos Arctica for sample optimisation and a Titan Krios for high-resolution data collection.
-The Umeå node is expected to become operational during 2018, and will offer cryo-ET with a 
-Titan Krios and a Scios DualBeam SEM.
+The Ume\aa\ node is expected to become operational during 2018, and will offer cryo-ET with a Titan Krios and a Scios DualBeam SEM.
 
-The facilities can be accessed by Swedish researchers through a peer-reviewed routine. Applications can be 
-submitted through an Application Portal [**reference???] and will be evaluated once every three months based on their 
-scientific merit and technical feasibility by a national Project Evaluation Committee. On the other side, 
-some time of the facility is reserved for internal research groups at Stockholm University. Internally, the allocation
-of the microscope (and other instruments) time is managed by a Booking System [** name of the booking system and reference???].
+The facilities can be accessed by Swedish researchers through 
+a peer-reviewed process.
+Applications can be submitted through an Application Portal (\url{https://cryoem.scilifelab.se/}).
+% and will be evaluated once every three months based on their scientific merit and technical feasibility by a national Project Evaluation Committee.
+%On the other side, some time of the facility is reserved for internal research groups at Stockholm University.
+Internally, the allocation of the microscope (and other instruments) time is managed by a Booking System [**ROB-JM name of the booking system and reference???].
 
-At the beginning of 2016, the Cryo-EM facility at SciLifeLab became one of the early adopters of \scipion 
-for on the fly data processing. Before, there were some home-made scripts to perform motion correction (with motioncor)
-and CTF estimation (with Gctf). The microscope operators recognized the importance of  
-a more general framework with the possibility to be extended with new programs. Another strong point was the 
-the availability of graphical tools for data analysis and quick feedback about the data collection. Initially, the 
-same setup script from the CNB was used, but it had some limitations for this more complex setup with more 
-requirements. In the following sections we will briefly describe the computational infrastructure at the
-SciLifeLab node and the implementation of a Session Wizard to automate the initial setup while fetching 
-information from other sources external to Scipion.
+At the beginning of 2016, the Cryo-EM facility at SciLifeLab became one of the early adopters of \scipion for on the fly data processing.
+Before, there were some home-made scripts to perform motion correction (with motioncor) and CTF estimation (with Gctf).
+The microscope operators recognized the importance of
+a more general framework with the possibility to be extended with new programs.
+Another strong point was the the availability of graphical tools for data analysis and quick feedback about the data collection.
+Initially, the same setup script used at CNB was adopted, but it did not fully satisfy the facility requirements and a new one has been developed.
+In the following sections we will briefly describe the computational infrastructure at the SciLifeLab node [** The one in  Stockholm?] and the implementation of a Session Wizard to automate the initial setup while fetching information from other sources external to \scipion.
 
-\subsubsection{Infrastructure}
+\subsubsection{Network Setup and IT Infrastructure}
 At the heart of the SciLifeLab computational infrastructure is a storage and pre-processing server (the
@@
 
-Unfortunately the K2 computer writes data to a local SSD RAID, the Falcon-III to
-a dedicated storage server (the offload server), in order to avoid interruption
-of data collection due to network issues. So when users start data collection,
-they start a script that continously moves the produced data to the staging
-server.
+Unfortunately, the K2 [** Jose Miguel, what is K2?, the computer described in the previous paragraph?] computer writes data to a local SSD RAID while the Falcon-III does it to a dedicated storage server (the offload server).
+%, in order to avoid interruption of data collection due to network issues.
+So when users start the data collection, they launch a script that continuously moves the produced data to the staging server.
 
-By having all data on the staging server, in an organized folder structure, we
-can run pre-processing directly on the server. If processing is started soon
-after data collection the incoming file data should still be in RAM and does not
-have to be read from disk, which is important with performance in mind. Users
+[** ROB. I do not understand this sentence, what preprocessing:  By having all data on the staging server, in an organized folder structure, we
+can run pre-processing directly on the server.] If data processing is started soon
+after data collection the incoming data file should be in RAM-cache and  no
+disk access in needed, which is important with performance in mind. Users
 can copy their data to external USB drives from two workstations in the
@@
 There is also a dedicated download server with access to the NFS export,
-reachable from the Internet. Here users can log in to a restricted shell that
+reachable though the Internet. Here users can log in to a restricted shell that
 allows downloads via rsync (over SSH) or SFTP.
 
-All these computers are conntected to a 10 GbE network switch. This is a private
+All these computers are connected to a 10 GbE network switch. This is a private
 network without external access. The only machine accessible from the Internet
@@
 \subsubsection{Session Wizard}
-
-The initial attempt to ease the setup of \scipion projects for stream processing was based  
-on the original script developed at the CNB. But it was difficult to reuse the existing code due to a different folder
-structure and, more important, a different logic about users and microscopes management. To address the new set
-of requirements, a new ``wizard'' is being developed at SciLifeLab and can be found in the following open source repository: 
-\url{https://github.com/delarosatrevin/scipion-session}
-
-For this new wizard, a basic data model was designed representing the current processing and data 
-organization at SciLifeLab. The model is composed by the following entities:
+In order to interface \scipion with the microscope infrastructure an upgraded version of the original script developed at the CNB was used.
+Unfortunately, it was difficult to reuse the existing code due to a different folder structure and, more important, a different logic about users and microscopes management.
+To address the new set of requirements, a new wizard is under development at SciLifeLab.
+For this new wizard, a basic data model was designed representing the current processing and data organization at SciLifeLab.
+The model is composed by the following entities:
 
@@
 
-When an user starts a data collection, the wizard is also executed using the command \textit{session-titan} 
-(or \textit{session-talos}). The wizard will fetch the reservations for this day using a webservices API provided by 
-the Booking System and will check wich user is operating the microscope. From the reservation information, 
-the wizard will also detect if it is an internal project or a national facility project. If the latter,
-more information will be retrieved from the Applicacion Portal (such as principal investigator, project code, 
-contact person or invoice address). After gathering all this information, the wizard setup the Scipion project
-for streaming processing. For example, the movies-import protocol will already point to the location where
-the files will be written for this session. Moreover, taking into account which microscope-camera is being 
-used, the files pattern can be inferred.
+When an user starts a data collection, the wizard is also executed using the command \textit{session-titan} (or \textit{session-talos}).
+The wizard will fetch the reservations for this day using a webservices API provided by the Booking System and will check which user is operating the microscope.
+From the reservation information, the wizard will also detect if it is an internal project or a national facility project.
+If the latter, more information will be retrieved from the Application Portal (such as principal investigator, project code, contact person or invoice address).
+After gathering all this information, the wizard setup the \scipion project for streaming processing.
+%For example, the movies-import protocol will already point to the location where the files will be written for this session.
+Moreover, taking into account which microscope-camera is being  used, the files pattern can be inferred.
 
-The current implementation of the wizard stores each session in a simple database file. This database, together
-with extra information from the Booking System and the Appplication Portal, is used to generate reports for 
-a given period with the associated invoices. The generation of a session report is under development, that 
-will provide a summary and useful information for users to access the acquired data. The existing data model 
-represents the specify usage at the SciLifeLab facility and there are not inmediate plans to make it more 
-general and easy to reuse. Nonetheless, the model is quite simple and should to be difficult to modify it to fit 
-other requirements.
+The current implementation of the wizard stores each session in a simple database file.
+This database, together with extra information from the Booking System and the Appplication Portal, is used to generate reports for a given period with the associated invoices.
+The generation of a session report is under development, that
+will provide a summary and useful information for users to access the acquired data.
+
+The described session wizard can be found at  \url{https://github.com/delarosatrevin/scipion-session}.
+It has been design to satisfy SciLifeLab requirements and, therefore, although readers are welcomed to try it, no support is offered.
+
+%existing data model represents the specify usage at the SciLifeLab facility and there are not immediate plans to make it more general and easy to reuse.
+Nonetheless, the model is quite simple and should to be difficult to modify it to fit other requirements.

### 2018-03-22 5.conclusions.tex
before class

@@
-\section{Discussion}
-[**TODO]: Discuss the obtained numbers in the result section
-[**REmove this section?]
+%\section{Discussion}
+%[**TODO]: Discuss the obtained numbers in the result section
+%[**REmove this section?]
 
@@
 Furthermore, not having the processing path hardcoded in the program means you don't have to recompile, so changes can be quickly tested   even at run time.
-The \scipion workflow approach provides provenance, that is,  for any created object (micrograph, CTFs, class, etc) \scipion stores metadata such as which programs, which parameters, and which exact path was followed to create it.
+The \scipion workflow approach provides provenance, that is,  for any created object (micrograph, CTFs, class, etc) \scipion stores metadata such as which programs, which parameters, and which exact path was followed to create any output data.
 In this way,  real tracking and reproducibility are provided.
@@
 Cloud installation instruction are available at \url{https://github.com/I2PC/scipion/wiki/Scipion-in-the-Cloud}.
-Other public software mentioned in this article as \emadmin, XXXX and YYYY are available at \url{https://github.com/rmarabini/webservices/tree/master/EMadmin}, www eee respectively.
+Other public software mentioned in this article as \emadmin, \emadmin and \emph{wizard session} are available at \url{https://github.com/rmarabini/webservices/tree/master/EMadmin} and \url{\url{https://github.com/delarosatrevin/scipion-session}} respectively.
 If you would like more information about how to install \scipion or you have suggestions on how to improve it contact us at scipion@cnb.csic.es.

### 2018-03-22 6.acknowledgments.tex
go home

@@
 
-UPDATE THIS acknowledgments
-
-Javier Vargas is the recipient of a ‘‘Comfuturo” grant funded by
-the Fundación General CSIC, C.O.S. Sorzano is the recipient of a
-Ramón y Cajal fellowship from the Spanish Ministry of Economy
-and Competitiveness (MINECO). The authors acknowledge financial
-support from the MINECO through Grants AIC-A-2011-0638,
-BFU2013-41249-P and BIO2013-44647-R, the Comunidad de
-Madrid through Grant CAM (S2010/BMD-2305), the European
-Union (EU) and Horizon 2020 through Grants EINFRA-2014-2:
-EGI Engage (Proposal: 654142) and EINFRA-2015-1: West-Life
-(Proposal: 675858).
-This work was partially funded by Instruct, a component of the
-European Strategy Forum on Research Infrastructures (ESFRI), and
-supported by national member subscriptions.
+The authors would like to acknowledge economical support from:
+The Spanish Ministry of Economy and Competitiveness through Grants BIO2013-44647-R, BIO2016-76400-R(AEI/FEDER, UE) and AEI/FEDER BFU 2016 74868P, the Comunidad Aut\'{o}noma de Madrid through Grant: S2017/BMD-3817, European Union (EU) and Horizon 2020 through grant CORBEL (INFRADEV-1-2014-1, Proposal: 654248).
+This work used the EGI Infrastructure and is co-funded by the EGI-Engage project (Horizon 2020) under Grant number 654142.
+European Union (EU) and Horizon 2020 through grant West-Life (EINFRA-2015-1, Proposal: 675858) European Union (EU) and Horizon 2020 through grant Elixir - EXCELERATE (INFRADEV-3-2015, Proposal: 676559) European Union (EU) and Horizon 2020 through grant iNEXT (INFRAIA-1-2014-2015, Proposal: 653706).
+The authors acknowledge the support and the use of resources of Instruct, a Landmark ESFRI project.
 

### 2018-03-22 2.overall.tex
from home

@@
 The second protocol performs direct detector movie alignment.
-Currently \scipion offers  four  methods for movie alignment: motioncorr2 (Li et al., 2013), correlation (an unpublished CPU version that implements the global aligment method used by motioncorr2), unblur (Grant and Grigorieff, 2015) and optical flow (Abrishami et al., 2015).
+Currently \scipion offers  four  methods for movie alignment: motioncor2 (Li et al., 2013), correlation (an unpublished CPU version that implements the global alignment method used by motioncor2), unblur (Grant and Grigorieff, 2015) and optical flow (Abrishami et al., 2015).
 Then the CTF of the resulting micrographs are calculated.

### 2018-03-22 3.workflow_scilifelab.tex
from home

@@
 a peer-reviewed process.
-Applications can be submitted through an Application Portal (\url{https://cryoem.scilifelab.se/}).
+Applications can be submitted through an application portal (\url{https://cryoem.scilifelab.se/}).
 % and will be evaluated once every three months based on their scientific merit and technical feasibility by a national Project Evaluation Committee.
 %On the other side, some time of the facility is reserved for internal research groups at Stockholm University.
-Internally, the allocation of the microscope (and other instruments) time is managed by a Booking System [**ROB-JM name of the booking system and reference???].
+Internally, a online booking system centralize the reservations of the microscopes and other instruments [**ROB-JM name of the booking system and reference???].
 
 At the beginning of 2016, the Cryo-EM facility at SciLifeLab became one of the early adopters of \scipion for on the fly data processing.
-Before, there were some home-made scripts to perform motion correction (with motioncor) and CTF estimation (with Gctf).
-The microscope operators recognized the importance of
-a more general framework with the possibility to be extended with new programs.
-Another strong point was the the availability of graphical tools for data analysis and quick feedback about the data collection.
+Before, there were some home-made scripts to perform motion correction (with motioncorr) and CTF estimation (with Gctf) but the microscope operators recognized the importance of a more general framework.
+In particular, they were interested in the possibility of modify easily the processing workflow as well as provide users and staff with  graphical tools for data analysis and quick feedback regarding the data collection.
 Initially, the same setup script used at CNB was adopted, but it did not fully satisfy the facility requirements and a new one has been developed.
@@
 
-Unfortunately, the K2 [** Jose Miguel, what is K2?, the computer described in the previous paragraph?] computer writes data to a local SSD RAID while the Falcon-III does it to a dedicated storage server (the offload server).
+Unfortunately, the K2 [** Jose Miguel, what is K2?, do you have 2 cameras?
+in the same microscope or in different ones, this is confusing, comment when ] computer writes data to a local SSD RAID while the Falcon-III does it to a dedicated storage server.
 %, in order to avoid interruption of data collection due to network issues.
@@
 
-[** ROB. I do not understand this sentence, what preprocessing:  By having all data on the staging server, in an organized folder structure, we
-can run pre-processing directly on the server.] If data processing is started soon
-after data collection the incoming data file should be in RAM-cache and  no
-disk access in needed, which is important with performance in mind. Users
-can copy their data to external USB drives from two workstations in the
-microscope room, that have access to the storage pool via NFS.
-There is also a dedicated download server with access to the NFS export,
-reachable though the Internet. Here users can log in to a restricted shell that
-allows downloads via rsync (over SSH) or SFTP.
+%ata on the staging server, in an organized folder structure, we
+%can run pre-processing directly on the server.] If data processing is started soon
+%after data collection the incoming data file should be in RAM-cache and  no
+%disk access in needed, which is important with performance in mind. 
+After processing, users can copy their data to external USB drives from two workstations in the
+microscope room that have access to the storage pool via NFS.
+There is also a dedicated download server which have access to the NFS export. This download server 
+can be reached though the Internet and allow downloads via rsync (over SSH) or SFTP.
 
@@
 is the data download server, which has a dual-port 10 GbE network adapter (one
-used for the private network, one for the external network). Both the K2
-computer and the Falcon-III offload server had one unused 10GbE interface that
-could be used to connect them to our private network, without having to
-interfere with the default setup.
+used for the private network, one for the external network).
+%Both the K2 computer and the Falcon-III dedicated storage server had one unused 10GbE interface that could be used to connect them to our private network, without having to interfere with the default setup.
 
@@
 When an user starts a data collection, the wizard is also executed using the command \textit{session-titan} (or \textit{session-talos}).
-The wizard will fetch the reservations for this day using a webservices API provided by the Booking System and will check which user is operating the microscope.
-From the reservation information, the wizard will also detect if it is an internal project or a national facility project.
+The wizard will fetch the reservations for this day using the webservice API provided by the Booking System [** chanr booking system by the official name] and will check which user is operating the microscope.
+From the reservation information, the wizard will also detect if it is an internal or a national facility project.
 If the latter, more information will be retrieved from the Application Portal (such as principal investigator, project code, contact person or invoice address).
-After gathering all this information, the wizard setup the \scipion project for streaming processing.
+After gathering all this information, the wizard will store this information, create the processing environment and launch \scipion.
 %For example, the movies-import protocol will already point to the location where the files will be written for this session.
@@
 The current implementation of the wizard stores each session in a simple database file.
-This database, together with extra information from the Booking System and the Appplication Portal, is used to generate reports for a given period with the associated invoices.
-The generation of a session report is under development, that
-will provide a summary and useful information for users to access the acquired data.
+This database, together with extra information from the Booking System and the Application Portal, is used to generate reports for a given period and the associated invoices.
+%The generation of a session report is under development, that will provide a summary and useful information for users to access the acquired data.
+%[** this last sentence is not very informative]
 
 The described session wizard can be found at  \url{https://github.com/delarosatrevin/scipion-session}.
-It has been design to satisfy SciLifeLab requirements and, therefore, although readers are welcomed to try it, no support is offered.
+It has been design to satisfy SciLifeLab requirements and, although readers are welcomed to try it, no support is offered.
 

### 2018-03-23 3.workflow_diamond.tex
before class
go home

@@
-\subsection{The eBIC facility at Diamond Light Source}
+\subsection{Case 3: Scipion at eBIC facility at Diamond Light Source}
 
@@
 In contrast to the CNB setup, Diamond has integrated the microscopes in much the same fashion to that of the beamlines at the facility, as is shown in figure X
+\subsubsection{Network Setup and IT Infrastructure}
 
@@
 This provides the users with a web interface showing general metadata about their experiment as well as new real-time updates on steps of the process and relevant output such as motion corrected images, drift values and CTF images (Figure XXX)
+
+\subsubsection{Additional Software Developments: }
+
+\subsection{Additional Sotware Developement: ISyB monitor}
+
+0) headless <-- no GUI
+1) Problem: access file system: very spensive using glob
+2) Modify import movies so it can start  sever that waits until "something" tells it that there is a new file. [quien reporta activeQ como]
+3) import though webservice << rama de diamond. pull request
+[** quien crea el sistema de directorios, create project]
+4) Wizard that creates normalized enviroment? Scipion workflow customiser. -> template scipion
+5) IspyB monitor -> connect to IspyB.(thumbnail-mic and ctf, defocus, resolution)
+[headless version]
+urls: where is this in scipion (given that there is no SNRF and Diamond)
+
+where is scipion monitor customizer
+
+>>> que desencadena el proceso?

### 2018-03-23 3.workflow_scilifelab.tex
go home

@@
 %On the other side, some time of the facility is reserved for internal research groups at Stockholm University.
-Internally, a online booking system centralize the reservations of the microscopes and other instruments [**ROB-JM name of the booking system and reference???].
+Internally, a online booking system called \emph{Booked Scheduler} centralizes the reservations of the microscopes and other instruments.
 
@@
 Initially, the same setup script used at CNB was adopted, but it did not fully satisfy the facility requirements and a new one has been developed.
-In the following sections we will briefly describe the computational infrastructure at the SciLifeLab node [** The one in  Stockholm?] and the implementation of a Session Wizard to automate the initial setup while fetching information from other sources external to \scipion.
+In the following sections we will briefly describe the computational infrastructure at the SciLifeLab node at Stockholm and the implementation of a Session Wizard to automate the initial setup while fetching information from other sources external to \scipion.
+[** thalos falcon III, titan k2 and falcon III]
 
@@
 
-\subsubsection{Session Wizard}
+\subsubsection{Additional Software Developments: \emph{Session Wizard}}
 In order to interface \scipion with the microscope infrastructure an upgraded version of the original script developed at the CNB was used.
@@
 
+[** quitar modelo de datos]
 \begin{itemize}

### 2018-03-23 6.acknowledgments.tex
go home

@@
 The authors would like to acknowledge economical support from:
+
 The Spanish Ministry of Economy and Competitiveness through Grants BIO2013-44647-R, BIO2016-76400-R(AEI/FEDER, UE) and AEI/FEDER BFU 2016 74868P, the Comunidad Aut\'{o}noma de Madrid through Grant: S2017/BMD-3817, European Union (EU) and Horizon 2020 through grant CORBEL (INFRADEV-1-2014-1, Proposal: 654248).
+The ``Knut \& Alice Wallenberg Foundation'', and ``A Pilot Facility development grant from Science for Life Laboratory''.
+
 This work used the EGI Infrastructure and is co-funded by the EGI-Engage project (Horizon 2020) under Grant number 654142.
@@
 The authors acknowledge the support and the use of resources of Instruct, a Landmark ESFRI project.
+This work has used the Cryo-EM Swedish National Facility funded by the Knut and Alice Wallenberg, Family Erling Persson and Kempe Foundations, SciLifeLab, Stockholm University and Ume\aa\ University.
+
+We acknowledge Diamond for access and support of the Cryo-EM facilities at the UK national electron bio-imaging centre (eBIC), funded by the Wellcome Trust, MRC and BBSRC.
 

### 2018-03-23 scipionbox.tex
before class
go home

@@
 \def\emadmin{\textit{EMAdmin}\xspace}
+\def\epu{\textit{EPU}\xspace}
 \newcommand{\protocol}[1]{\texttt{#1}}
@@
 \title{Using Scipion for Stream Image Processing at EM Facilities}
-\author{XXXX$^{(a)}$,\\
-        YYYY$^{(b)}$,\\
+\author{Scipion Boys$^{(a)}$,\\
+Jos\'e Miguel de la Rosa-Trev\'{i}n$^{(c)}$,\\
+Marta Carroni$^{(c)}$,\\
+Stefan Fleichman$^{(c)}$,\\
+Erik Lindahl$^{(c,d)}$\\
+Ashton AW$^{(e)}$,\\
+Basham M$^{(e)}$,\\
+Clare DK$^{(e)}$,\\
+Savage K$^{(e)}$,\\
+Siebert CA$^{(e)}$,\\
         }
@@
 \lyxaddress{{\small$^{(b)}$ Biocomputing Unit, National Center for Biotechnology (CSIC), C/ Darwin, 3, Campus Universidad Aut\'{o}noma, 28049 Cantoblanco, Madrid, Spain.}}
-
+\lyxaddress{{\small$^{(c)}$ Department of Biochemistry and Biophysics, Science for Life Laboratory, Stockholm University, Stockholm, Sweden}} 
+\lyxaddress{{\small$^{(d)}$ Swedish e-Science Research Center, KTH Royal Institute of Technology, Stockholm, Sweden}}
+\lyxaddress{{\small$^{(e)}$ Diamond Light Source, Harwell Science and Innovation Campus, Didcot OX11 0DE, England}}
 %This document provides a preliminary analysis of the data collected by the CTF Challenge (http://i2pc.cnb.csic.es/3dembenchmark/LoadHome.htm).

### 2018-03-29 scipionbox-abstract-only.tex
Added abstract only file for ACA conference

--- 2018-03-23 scipionbox.tex
@@
 \title{Using Scipion for Stream Image Processing at EM Facilities}
-\author{Scipion Boys$^{(a)}$,\\
-Jos\'e Miguel de la Rosa-Trev\'{i}n$^{(c)}$,\\
-Marta Carroni$^{(c)}$,\\
-Stefan Fleichman$^{(c)}$,\\
-Erik Lindahl$^{(c,d)}$\\
-Ashton AW$^{(e)}$,\\
-Basham M$^{(e)}$,\\
-Clare DK$^{(e)}$,\\
-Savage K$^{(e)}$,\\
-Siebert CA$^{(e)}$,\\
+
+\author{
+J.M. de la Rosa-Trev\'{i}n$^{(c)}$, \\
+J. G\'omez-Blanco$^{(b)}$, P. Conesa, Y. Rancel, L. del Cano, \\
+R. Arranz, F.J. Chich\'on, J. Mart\'in-Benito, \\
+M. Carroni$^{(c)}$, S. Fleichman$^{(c)}$, E. Lindahl$^{(c,d)}$,\\ 
+M. Basham M$^{(e)}$, D.K. Clare $^{(e)}$, K. Savage$^{(e)}$, \\
+C.A. Siebert CA$^{(e)}$, A.W, Ashton$^{(e)}$, \\
+R. Marabini $^{(a,b)}$, C.O.S. Sorzano$^{(b)}$, J.M. Carazo $^{(b)}$ 
 }
@@
 \lyxaddress{{\small$^{(e)}$ Diamond Light Source, Harwell Science and Innovation Campus, Didcot OX11 0DE, England}}
-%This document provides a preliminary analysis of the data collected by the CTF Challenge (http://i2pc.cnb.csic.es/3dembenchmark/LoadHome.htm).
-Figures/tables/plots/etc are located at the document's end.
-\setcounter{figure}{2}%I do not know why but figure counter starts with -1
-\textbf{Corresponding Author:} %Roberto Marabini, Ph= 34 91 5854510, email: roberto@cnb.csic.es\\
-%Escuela Polit\'{e}cnica Superior\\
-%Universidad Aut\'{o}noma de Madrid\\
-%28049 Cantoblanco\\
-%Madrid\\
-%Spain
 
@@
 \input{0.abstract.tex}
-\input{1.introduction.tex}
-\input{2.overall.tex}
-\input{3.workflow.tex}
-\input{4.results.tex}
-\input{5.conclusions.tex}
-\input{6.acknowledgments.tex}
-\appendix
-
-\bibliographystyle{elsart-harv}
-\bibliography{large}
-
-\clearpage
-%\input{listOfFigures.tex}
-
-\newpage
-%tables
-%\input{numberImagesDataSet.tex}
-%%%%%\input{markersLegend.tex}
-%\input{numberImagesUpload.tex}
-%\newpage
 

### 2018-04-02 3.workflow.tex
almost there switty

@@
  %\item \emph{Ume\aa\ Core Facility for Electron Microscopy} (UCEM, \url{http://www.kbc.umu.se/english/ucem})[** Jose Miguel can you verify this?] 
- \item \emph{Science for life Laboratory} (SciLifeLab, \url{http://www.scilifelab.se}). Both at Stockholm and Ume\aa\ facilities)
+ \item \emph{The Swedish National Cryo-EM Facility}.
+The Facility has two nodes: at \scilifelab in Stockholm and at Ume\aa\ University (\url{http://www.scilifelab.se}, \url{http://www.kbc.umu.se/english/ucem/cryo-em/})
  \item \emph{ESRF Cryo-Electron Microscope} (\url{http://www.esrf.eu/home/UsersAndScience/Experiments/MX/About_our_beamlines/CM01.html})
@@
 \input{3.workflow_diamond.tex}
-
-\subsection{\scipion near future}
-%Particle picking
-%Extract particle
-%2D classify (either classes or identify outliers)
-%Initial model
-%More complex reports
-plug-in style, independend

### 2018-04-02 3.workflow_diamond.tex
almost there switty

--- 2018-03-21 3.workflow_diamond.tex
@@
-\subsection{The eBIC facility at Diamond Light Source}
+\subsection{Case 3: Scipion at eBIC facility (Diamond Light Source)}
 
@@
 Due to the huge demand for high-end microscope access,
-The electron Bio-Imaging Centre (eBIC) was set up at Diamond~\citep{diamond2017}.
+Recently, the electron Bio-Imaging Centre (eBIC) has been set up at Diamond~\citep{diamond2017}.
 The centre houses four Thermofisher Titan Krios, a Talos Arctica and a Scios dual-beam.
 All of the transmission electron microscopes are equipped with counting-mode direct electron detectors and Volta phase plates.
+
+\subsubsection{Network Setup and IT Infrastructure}
 
@@
 Data is collected onto a local machine %in a similar way to the CNB setup, but 
-and then immediately moved via 10Gbps Ethernet to the associated visit directory in the centralised storage, a multi Petabyte GPFS High Performance Parallel file system.
+and then immediately moved via 10Gbps Ethernet to the associated visit directory in the centralized storage, a multi Petabyte GPFS high performance parallel file system.
 Once the raw data is present on the GPFS system, high speed interconnects to the central cluster enable rapid data processing.
@@
 
-Once the data has reached the central file system, it is available to be processed by Scipion.
-This process is facilitated by the use of a Scipion-workflow-template which enable novice users to trigger semi-customised on the fly processing.
-This template consists of a small UI (figure XX) which allows the user to put in the specific values pertinent to their experiment, namely the session ID (or visit number), which microscope they are using, and several other key data collection parameters.
-Once set this makes use of a global template for Scipion as set up by the eBIC team, and substitutes the appropriate information into it, and then saves it to disk.
+Diamond IT infrastructure is complex and requires to integrate a large collection of heterogeneous systems.
+In this environment a major problem is to achieve interaction and data exchange between diverse applications.
+One of the solutions to this problem has been to implement \emph{indirect communication} through a message queuing system.
+Simply said; a message queuing system is a software where queues can be defined, applications may connect to the queue and send or retrieve  messages.
+A message can include any kind of information, for example, about a process/task that should start on another application (that could be on another server).
+%The message queuing software stores the messages until a receiving application connects and takes a message off the queue.
+The receiving application then processes the message in an appropriate manner.
 
-Following this Diamond makes use of a central ActiveMQ [** what is this?
-a queue system] service which is triggered upon the template file being available in the visit directory, and runs a headless (i.e.
-without UI) version of Scipion which takes the template and executes in streaming mode.
-THIS SHOULD PROBABLY INCLUDE SOMETHING ABOUT THE DETECTION OF NEW FILES BY SOCKET MESSAGE, BUT I DON’T KNOW HOW THIS WORKS.
-Me neither, I thought it polled the filesystem?
+\subsection{Additional Software Developments: ISyB monitor, Scipion Headless}
 
-To monitor the progress of the automatic processes, the Scipion pipeline at Diamond also includes custom nodes which allow it to interface with ISPyB.
-This provides the users with a web interface showing general metadata about their experiment as well as new real-time updates on steps of the process and relevant output such as motion corrected images, drift values and CTF images (Figure XXX)
+From the image processing point of view, in Diamond, a typical cryoEM session starts by launching a custom made script that performs two actions.
+First, it  allows users to specify values pertinent to their experiment, namely the session ID (or visit number), which microscope they are using, and several other key data collection parameters.
+Once set, a workflow file is created and then saved to disk.
+The file creation triggers the message system \emph{ActiveMQ} and a ``file creation'' message is created.
+The second action performed by the custom made script is to start a \scipion instance in the microscope control computer.
+The instance runs in headless mode. \scipion headless mode has been developed specifically for eBIC and allows \scipion to be executed without user interface in large clusters with no graphical board.
+
+%(a) script: (1) create workflow, (2) launchs headless scipion
+%(b) scipion starts webservice [knows nothing about the queue]
+%(c) second script (name please) (running all the time?) ask activeMQ for messages
+%    speaks with import movies [decuple Scipion from ActiveMQ], more generic solution
+
+After \scipion is launched the \protocol{import movie} protocol starts.
+\protocol{Import movie} needs to look for a list of files on the file system with names matching a pattern but, in Diamond distributed file system, this is a very expensive task.
+Therefore, \protocol{import movie} has been modified so it opens a webservice that waits until an auxiliary script connects to \emph{ActiveMQ}, retrieves all the  messages reporting creation of new files than satisfy the search pattern and reports the actual filenames.
+It would have been possible to suppress the auxiliary script and make \protocol{import movie} access directly to \emph{ActiveMQ} but we decided against this solution because we did not want to incorporate in \scipion code valid for a single facility.
+In this way, all code that is related with Diamond's IT infrastructure is encapsulated in the auxiliary script.
+
+To monitor the progress of the automatic image processing CNB and \scilifelab cryo-EM services make use of \scipion default monitor system.
+In Diamond, the electronic notebook ISPyB is the standard tool to record and show metadata about the different experiments.
+In this way, a new monitor has been developed which, through a webservice, allows \scipion to directly interface with ISPyB.
+This setup provides the users with a web interface showing general metadata about their experiment as well as  real-time updates on steps of the process and relevant output such as motion corrected images, drift values, CTF images, etc.
+
+Although most of developments described in this subsection will be integrated in the main \scipion code repository at present they are available in a developer's branch  at \url{https://github.com/I2PC/scipion/tree/release-1.1-headless-devel}.

### 2018-04-02 3.workflow_scilifelab.tex
almost there switty

@@
-\subsection{Case 2: Scipion at SciLifeLab}
+\subsection{Case 2: Scipion at The Swedish National Cryo-EM Facility}
 
 The Swedish National Cryo-EM Facility offers access to state-of-the-art equipment and expertise in single particle cryo-EM and cryo electron tomography (cryo-ET).
-The Facility has two nodes: at SciLifeLab in Stockholm and at Ume\aa\ University.
-SciLifeLab in Stockholm offers single-particle service with a
+The Facility has two nodes: at \scilifelab in Stockholm and at Ume\aa\ University.
+\scilifelab in Stockholm offers single-particle service with a
 Talos Arctica for sample optimisation and a Titan Krios for high-resolution data collection.
@@
 
-At the beginning of 2016, the Cryo-EM facility at SciLifeLab became one of the early adopters of \scipion for on the fly data processing.
+At the beginning of 2016, the Cryo-EM facility at \scilifelab became one of the early adopters of \scipion for on the fly data processing.
 Before, there were some home-made scripts to perform motion correction (with motioncorr) and CTF estimation (with Gctf) but the microscope operators recognized the importance of a more general framework.
@@
 Initially, the same setup script used at CNB was adopted, but it did not fully satisfy the facility requirements and a new one has been developed.
-In the following sections we will briefly describe the computational infrastructure at the SciLifeLab node at Stockholm and the implementation of a Session Wizard to automate the initial setup while fetching information from other sources external to \scipion.
-[** thalos falcon III, titan k2 and falcon III]
+In the following sections we will briefly describe the computational infrastructure at the \scilifelab node and the implementation of a Session Wizard to automate the initial setup while fetching information from other sources external to \scipion.
+%[** thalos falcon III, titan k2 and falcon III]
 
 \subsubsection{Network Setup and IT Infrastructure}
-At the heart of the SciLifeLab computational infrastructure is a storage and pre-processing server (the
-staging server). This machine has roughly 200 TB of storage (4 ZFS RAIDZ2 pools
-with 11 HDDs each), 2 NVIDIA GeForce GTX 1070 GPUs, 2 Intel Xeon E5-2630v4 CPUs
-(10 cores, 2.2 GHz, HT), 384 GB RAM, and a dual-port 10 GbE network card. The
-storage pool is exported via NFS and Samba, so the microscope computers can
-write data there directly and other machines can access it.
-
-Unfortunately, the K2 [** Jose Miguel, what is K2?, do you have 2 cameras?
+At the heart of the \scilifelab computational infrastructure there is a storage and pre-processing server (the staging server).
+This machine has roughly 200 TB of storage (4 ZFS RAIDZ2 pools with 11 HDDs each), 2 NVIDIA GeForce GTX 1070 GPUs, 2 Intel Xeon E5-2630v4 CPUs (10 cores, 2.2 GHz, HT), 384 GB RAM, and a dual-port 10 GbE network card.
+The storage pool is exported via NFS and Samba, so the microscope computers can write data there directly and other machines can access it.
+%Unfortunately, the K2 [** Jose Miguel, what is K2?, do you have 2 cameras?
 in the same microscope or in different ones, this is confusing, comment when ] computer writes data to a local SSD RAID while the Falcon-III does it to a dedicated storage server.
 %, in order to avoid interruption of data collection due to network issues.
-So when users start the data collection, they launch a script that continuously moves the produced data to the staging server.
+
+In order to avoid interruption of data collection due to network issues each microscope has their own dedicated storage, that in the case of the Titan microscope is double since it may be connected to two different cameras, a K2 and a Falcon III.
+In this way, after starting the data collection, users launch a script that continuously moves the produced data from the adequate storage to the staging server.
 
@@
 %disk access in needed, which is important with performance in mind. 
-After processing, users can copy their data to external USB drives from two workstations in the
-microscope room that have access to the storage pool via NFS.
-There is also a dedicated download server which have access to the NFS export. This download server 
-can be reached though the Internet and allow downloads via rsync (over SSH) or SFTP.
+After processing, users may  copy their data to external USB drives from two workstations in the microscope room that have access to the storage pool via NFS.
+There is also a dedicated download server which have access to the NFS export. This download server can be reached though the Internet and allow downloads via rsync (over SSH) or SFTP.
 
@@
 Unfortunately, it was difficult to reuse the existing code due to a different folder structure and, more important, a different logic about users and microscopes management.
-To address the new set of requirements, a new wizard is under development at SciLifeLab.
-For this new wizard, a basic data model was designed representing the current processing and data organization at SciLifeLab.
-The model is composed by the following entities:
+To address the new set of requirements, a new wizard is under development at \scilifelab.
+For this new wizard, a basic data model was designed representing the current processing and data organization at \scilifelab.
+%The model is composed by the following entities:
 
-[** quitar modelo de datos]
-\begin{itemize}
-\setlength\itemsep{0em}
- \item \textit{User}: represents all persons that can book microscope time or are registered in the Application Portal 
- related to national projects.
- \item \textit{Reservation}:  is a resource (e.g, Krios, Talos, Vitrobot, etc) booking for a given period of time.
- \item \textit{Order}:  is a given national project with an unique identifier in the Application Portal
- \item \textit{Project}:  can be a either a national project (related to a given Order) or an internal project.
- \item \textit{Session}:  stores information about the usage of the micrograph for a given day.
-\end{itemize}
+%[** quitar modelo de datos]
+%\begin{itemize}
+%\setlength\itemsep{0em}
+% \item \textit{User}: represents all persons that can book microscope time or are registered in the Application Portal 
+% related to national projects.
+% \item \textit{Reservation}:  is a resource (e.g, Krios, Talos, Vitrobot, etc) booking for a given period of time.
+% \item \textit{Order}:  is a given national project with an unique identifier in the Application Portal
+% \item \textit{Project}:  can be a either a national project (related to a given Order) or an internal project.
+% \item \textit{Session}:  stores information about the usage of the micrograph for a given day.
+%\end{itemize}
 
 When an user starts a data collection, the wizard is also executed using the command \textit{session-titan} (or \textit{session-talos}).
-The wizard will fetch the reservations for this day using the webservice API provided by the Booking System [** chanr booking system by the official name] and will check which user is operating the microscope.
+The wizard will fetch the reservations for this day using the webservice API provided by \emph{Booked Scheduler} and will check which user is operating the microscope.
 From the reservation information, the wizard will also detect if it is an internal or a national facility project.
@@
 The described session wizard can be found at  \url{https://github.com/delarosatrevin/scipion-session}.
-It has been design to satisfy SciLifeLab requirements and, although readers are welcomed to try it, no support is offered.
+It has been design to satisfy \scilifelab requirements and, although readers are welcomed to try it, no support is offered.
 
-%existing data model represents the specify usage at the SciLifeLab facility and there are not immediate plans to make it more general and easy to reuse.
+%existing data model represents the specify usage at the \scilifelab facility and there are not immediate plans to make it more general and easy to reuse.
 Nonetheless, the model is quite simple and should to be difficult to modify it to fit other requirements.

### 2018-04-02 4.future.tex
almost there switty

\subsection{Future Developments}

Currently, \scipion includes the appropriate code (python wrappers) to talk to the integrated EM packages (Xmipp, Eman, RELION, ...) at the time of the release. This implies that an update of any of the integrated EM packages immediately after the \scipion release, won't be available until next \scipion release (we are aiming for one release a year). To decouple EM-package releases from \scipion releases, we are working on making the wrappers for EM packages  independent from \scipion. Plans for ensuring this independence include: wrapper installation through pip (\url{https://pip.pypa.io/en/stable/}) and the reimplementation of the wrappers as plug-ins that will be automatically detected and added to the application menu.

Additionally, in collaboration with the Xmipp team, we are pushing the streaming workflow with the aim to obtain a first initial volume. We are also adding structure modeling capabilities, by integration of some functionality from programs such as Coot \citep{emsley2010:coot}, Refmac \citep{Murshudov1997:refmac}, etc.

As part of the EOSCPilot project (\url{https://eoscpilot.eu/cryoem}), we are also improving  how \scipion exports workflows in order to ease reporting of the work done. The goal is to write a workflow file that fully describes the image processing steps, so that if the same input data is provided to the workflow the same results should be obtained (making the data more compliant with FAIR guidelines given by the \citet{eu2016:fair}). This file could go with the raw data acquired by CryoEM facilities and deposited at common EM databases like EMPIAR or EMDB \citep{Patwardhan2016:databasesEM}. To facilitate the visualization of the workflow file in any of these databases, we are also developing a  webcomponent (\scipion \emph{workflow viewer}) that will easily allow these repositories to visualize the workflow on their web pages (\url{https://github.com/I2PC/web-workflow-viewer}). Finally, we are implementing a workflow repository. This repository will contain ready-to-use workflows that support a range of use cases.

### 2018-04-02 5.conclusions.tex
go home
almost there switty

@@
 Cloud installation instruction are available at \url{https://github.com/I2PC/scipion/wiki/Scipion-in-the-Cloud}.
-Other public software mentioned in this article as \emadmin, \emadmin and \emph{wizard session} are available at \url{https://github.com/rmarabini/webservices/tree/master/EMadmin} and \url{\url{https://github.com/delarosatrevin/scipion-session}} respectively.
-If you would like more information about how to install \scipion or you have suggestions on how to improve it contact us at scipion@cnb.csic.es.
+Other public software mentioned in this article as \emadmin, \emph{wizard session} and the different scripts used at eBIC are available at \url{https://github.com/rmarabini/webservices/tree/master/EMadmin}, \url{\url{https://github.com/delarosatrevin/scipion-session}} and \url{https://github.com/I2PC/scipion/tree/release-1.1-headless-devel} respectively.
+If you would like to have more information about how to install \scipion or you have suggestions on how to improve it contact us at scipion@cnb.csic.es.
 

### 2018-04-02 6.acknowledgments.tex
almost there switty

@@
 
-----------------
+%\section{Project stability and funding}
 
-    AFTER THIS LINE I JUST SAVED SOME TEXT THAT MAY BE USEFUL IN THE FUTURE DISREGARD THIS MATERIAL
+%Scipion development is been done in the context of the Instruct Image Processing Cen-
+%ter (I2PC, http://i2pc.cnb.csic.es). Instruct is the European initiative for Strategic
+%Scientific Infrastructures (ESFRI) for Structural Biology. The I2PC started with the objective 
+%to provide efficient support for research projects in Europe demanding expertise in 
+%electron microscopy image processing. In order to support a large collection
+%of projects, the first activities of the I2PC were oriented to design and implement the
+%necessary software tools.
 
-    --------------------
+%At present the I2PC has this money and this people
 
-\section{Project stability and funding}
-
-Scipion development is been done in the context of the Instruct Image Processing Cen-
-ter (I2PC, http://i2pc.cnb.csic.es). Instruct is the European initiative for Strategic
-Scientific Infrastructures (ESFRI) for Structural Biology. The I2PC started with the objective 
-to provide efficient support for research projects in Europe demanding expertise in 
-electron microscopy image processing. In order to support a large collection
-of projects, the first activities of the I2PC were oriented to design and implement the
-necessary software tools.
-
-At present the I2PC has this money and this people
-
-ScipionBox is part of the Scipion package which is provided 
-freely as open source software. Online documentation
-describing Scipion download and installation is available
-at \url{http://scipion.cnb.csic.es/m/download_form/} and \url{https://github.com/I2PC/scipion/wiki/How-to-Install}, respectively.
+%ScipionBox is part of the Scipion package which is provided 
+%freely as open source software. Online documentation
+%describing Scipion download and installation is available
+%at \url{http://scipion.cnb.csic.es/m/download_form/} and \url{https://github.com/I2PC/scipion/wiki/How-to-Install}, respectively.
 
@@
 %assure standardization, traceability and reproducibility.
-\section{Things to mention if possible}
-\begin{itemize}
- \item Cloud
-\end{itemize}
+%\section{Things to mention if possible}
+%\begin{itemize}
+% \item Cloud
+%\end{itemize}

### 2018-04-02 scipionbox.tex
almost there switty

@@
 \def\epu{\textit{EPU}\xspace}
+\def\scilifelab{SciLifeLab\xspace}
 \newcommand{\protocol}[1]{\texttt{#1}}
@@
 \title{Using Scipion for Stream Image Processing at EM Facilities}
-\author{Scipion Boys$^{(a)}$,\\
-Jos\'e Miguel de la Rosa-Trev\'{i}n$^{(c)}$,\\
-Marta Carroni$^{(c)}$,\\
-Stefan Fleichman$^{(c)}$,\\
-Erik Lindahl$^{(c,d)}$\\
-Ashton AW$^{(e)}$,\\
-Basham M$^{(e)}$,\\
-Clare DK$^{(e)}$,\\
-Savage K$^{(e)}$,\\
-Siebert CA$^{(e)}$,\\
+Authors to be sorted later
+\author{C.O.S. Sorzano$^{(a)}$,
+A. Jim\'{e}nez$^{(a)}$,
+R. Melero$^{(a)}$,\\
+M. Mart\'{i}nez$^{(a)}$,
+J. G\'{o}mez-Blanco$^{(a)}$,
+L. del Cano$^{(a)}$,\\
+T. Majtner$^{(a)}$,
+E. Ram\'{i}rez$^{(a)}$,
+P. Conesa$^{(a)}$,\\
+J.L. Vilas$^{(a)}$,
+R. Marabini$^{(a)}$,
+Y. Rancel$^{(a)}$,
+J.M. Carazo$^{(a)}$,\\
+J.M. de la Rosa-Trev\'{i}n$^{(c)}$,
+M. Carroni$^{(c)}$,
+S. Fleichman $^{(c)}$,\\
+E. Lindahl$^{(c,d)}$
+A.W. Ashton$^{(e)}$,
+M. Basham$^{(e)}$,\\
+D.K. Clare$^{(e)}$,
+K. Savage$^{(e)}$,
+C.A. Siebert$^{(e)}$,\\
+D. Maluenda$^{(a)}$
         }
@@
 \input{3.workflow.tex}
-\input{4.results.tex}
+\input{4.future.tex}
 \input{5.conclusions.tex}

### 2018-04-11 0.abstract.tex
w
carazo's review

@@
 Three dimensional electron microscopy (3DEM) is becoming a very data-intensive field in which vast amounts of experimental images are acquired at high speed.
-To manage such large-scale projects, we have previously developed  a modular workflow system called \scipion \citep{delarosatrevin2016}.
+To manage such large-scale projects, we had previously developed  a modular workflow system called \scipion \citep{delarosatrevin2016}.
 We present here a major extension of \scipion that allows stream processing of EM images while the data is being acquired.
@@
 implemented their stream processing solution using custom-made scripts, but this script based approach is not flexible and present a slow response time to new feature requests.
-The advantages of our proposed solution are: (1) flexibility: users can customize their workflows from a  large collection of preloaded algorithms; (2) repeatability and traceability: workflows are stored and can be re-executed later, either with the same data or a different data set; (3) monitorization: there are utilities that constantly check how the execution of the algorithms is going and inform about possible anomalies; (4) reports: we have developed several graphical interfaces that are updated periodically producing a summary (e.g, CTF defocus values, system load, etc).
+The advantages of our proposed solution are: (1) flexibility: users can customize their workflows from a  large collection of preloaded algorithms; (2) repeatability and traceability: workflows are stored and can be re-executed later, either with the same data or with a different data set; (3) monitorization: there are utilities that constantly check how the execution of the algorithms is going on and inform about possible anomalies; (4) reports: we have developed several graphical interfaces that are updated periodically producing a summary (e.g, CTF defocus values, system load, etc).
 This summary may be generated in HTML format and  published though a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided.
-At present, \scipion has been installed and is in production mode in several Cryo-EM facilities and it is being considered in many more.
+At present, \scipion has been installed and is in production mode in seven Cryo-EM facilities and it is being considered in many more.
 

### 2018-04-11 1.introduction.tex
carazo's review

@@
 3D electron microscopy (3DEM)  can provide rich information about structural characteristics of macromolecular complexes.
-The field is under a major transformation due to the arrival of better microscopes, new large area detectors and automatization \citep{kuhlbrandt2014a, Kuhlbrandt2014b}.
+The field is under a major transformation due to the arrival of better microscopes, new large area direct electron detectors and automatization \citep{kuhlbrandt2014a, Kuhlbrandt2014b}.
 These improvements make possible that a single microscope generates high quality data sets on the order of terabytes per day \citep{Saibil2015} while working for several days without interruption.
@@
 
-Clearly, the future of projects that require high-end microscopes is through centralized microscopy facilities, in which automatic acquired data needs to be monitored by users and facility staff.
+For most cases, the future of projects that require high-end microscopes is through centralized microscopy facilities, in which automatic acquired data needs to be monitored by users and facility staff.
 \scipion has been redesigned with this scenario in mind.
@@
 Exceptions are UCSFImage4 \citep{Li2015} and Focus \citep{Biyani2017}.
-As we will show in this article, \scipion differs from these solutions in a larger flexibility both from the point of view of users and facility staff.
+As we will show in this article, \scipion differs from these solutions in a larger flexibility, both from the point of view of users and facility staff.
 

### 2018-04-11 2.overall.tex
carazo's review

--- 2018-03-21 2.overall.tex
@@
 %Each protocol execution in \scipion is an independent process that can create processing threads. 
-Each protocol that works in streaming mode, has one of its threads monitoring input and updating output  data while other threads take care of the data processing.
-This clear separation between data access and data processing marks an important difference between running a script or using \scipion since it facilitates a form of computing in which several jobs are executed  concurrently, instead of sequentially.
-Moreover, \scipion workflow engine, allows users to modify the workflow on the fly without modifying a single code line.
+Each protocol that works in streaming mode has one of its threads monitoring input and updating output  data while other threads take care of the data processing.
+This clear separation between data access and data processing marks an important difference between running a script or using \scipion, since it facilitates a form of computing in which several jobs are executed  concurrently, instead of sequentially.
+Moreover, \scipion workflow engine allows users to modify the workflow on the fly without modifying a single code line.
 %In practice, the latter follows  a more general and extensible approach that allows users to modify the workflow on the fly without modifying a single code line.
@@
 
-We would like to finish this section assuring that despite the modifications introduced to \scipion data and process model, the framework provides full data provenance.
+We would like to finish this section stressing that despite the modifications introduced to \scipion data and process model, the framework keeps providing full data provenance.
 That is, \scipion provides detailed information about the origin of the data  and the process to generate it.
@@
 In this work we do not recommend a particular workflow  over others since we realize that different specimens, microscopes or facilities may have different requirements.
-Nevertheless, since a use case may clarify \scipion capabilities, in the following we describe the pipeline offered by default at the CNB cryoEM Service.
+Nevertheless, since a use case may clarify \scipion capabilities, in the following we describe the pipeline offered by default at the \emph{Centro Nacional de Biotecnología} (CNB) cryoEM Facility.
 
@@
 The second protocol performs direct detector movie alignment.
-Currently \scipion offers  four  methods for movie alignment: motioncorr2 (Li et al., 2013), correlation (an unpublished CPU version that implements the global aligment method used by motioncorr2), unblur (Grant and Grigorieff, 2015) and optical flow (Abrishami et al., 2015).
+Currently \scipion offers  four  methods for movie alignment: motioncor2 \citep{Zheng2016:motioncor2}, correlation (an unpublished CPU version that implements the global alignment method used by motioncor \citep{Li2015}), unblur \citep{Grant2015:unblur} and optical flow \citep{Abrishami2015:optical-flow}.
 Then the CTF of the resulting micrographs are calculated.
@@
 Further automatic filtering based on CTF is available through the protocol \protocol{CTF filtering} that selects micrographs based on the astigmatism, resolution and defocus.
-Another criteria for CTF selection is iciness, that is,  the ratio of the intensity in the resolution band between 3.5 and 3.9 \AA, and between 6 and 30 \AA, which gives a good estimate of the ice crystal content of the vitreous specimen.
+%Another criteria for CTF selection is iciness, that is,  the ratio of the intensity in the resolution band between 3.5 and 3.9 \AA, and between 6 and 30 \AA, which gives a good estimate of the ice crystal content of the vitreous specimen.
 
@@
 The proposed pipeline is very flexible and can be easily modified by each user.
-At the CNB (and collaborating institutions) we collect protocol usage statistics and sort the different algorithms by popularity.
+At the CNB (and collaborating institutions) we collect protocol usage statistics within \scipion and sort the different algorithms by popularity.
 This information is available at  \url{http://scipion.i2pc.es/report_protocols/protocolTable/} and may help users to select between the different algorithms.

### 2018-04-11 3.workflow.tex
carazo's review

@@
 The data flow created by an electron microscope requires a careful design of the IT infrastructures.
-Different facilities has adopted different solutions and therefore, how \scipion  access data and interchange information may require minor adjustments.
+Different facilities have adopted different solutions and therefore, how \scipion  access data and interchange information may require minor adjustments.
 In this section we discuss  the different solutions applied in the different facilities in which \scipion is running.
@@
 There, \scipion was adapted to interface with this booking platform in order to produce better reports.
-Lately,  \scipion has been installed in  large synchrotrons  such as ESRF and Diamond  where projects are handled by ISPyB (a customized laboratory management information system -LIMBS- \citep{Delageniere2011}) and data are saved on a distributed file system that heavily penalizes access to lists of files.
-In this environment, \scipion needs to constantly interchange information with ISPyB, and it cannot longer check if there is a new file in a particular storage device but must wait to be informed by the system that a new file has been created.
+Lately,  \scipion has been installed in  large synchrotrons  such as ESRF and Diamond  where projects are handled by ISPyB (a customized laboratory information management system -LIMS- \citep{Delageniere2011}) and data are saved on a distributed file system.
+% that heavily penalizes access to lists of files.
+In this environment, \scipion needs to constantly interchange information with ISPyB and with other applications through a message queue system called \emph{ActiveQ}.% and it cannot longer check if there is a new file in a particular storage device but must wait to be informed by the system that a new file has been created.
 

### 2018-04-11 3.workflow_cnb.tex
go home
carazo's review

@@
 This section in divided in two parts.
-In the first one we discuss the network setup that allow us to move the data from the microscope to the user laboratory.
+In the first one we discuss the network setup that allows us to move the data from the microscope to the user laboratory.
 Then  we comment on an additional software (called  \emadmin) that connects \scipion with the microscope acquisition software (\epu) and records the facility activity.
@@
 
-\subsubsection{\emadmin}
+\subsubsection{Additional Software Developments: \emadmin}
 
 \emadmin was born as a script that created a normalized tree of directories  in which to store the movies acquired by the microscope, helped users to choose between a few predetermined image processing workflows and finally launched \scipion.
-At present it was grown in a client-server application that in addition to the old script capacities it has incorporate the following functions:
+At present it has grown in a client-server application that in addition to the old script capacities it has incorporate the following functions:
 
 \begin{enumerate}
- \item Transparent addition of new workflows. \scipion is able to export any executed workflow as a json file. \emadmin can import this files as use them to create new predetermined workflows.
+ \item Transparent addition of new workflows. \scipion is able to export any executed workflow as a json file. \emadmin can import this files and use them to create new predetermined workflows.
  \item Automatic creation of pdf user's report containing: microscope acquisition parameters and histograms summarizing data resolution and  defocus. 

### 2018-04-11 3.workflow_diamond.tex
carazo's review

@@
 In this environment a major problem is to achieve interaction and data exchange between diverse applications.
-One of the solutions to this problem has been to implement \emph{indirect communication} through a message queuing system.
+One of the solutions to this problem has been to implement \emph{indirect communication} through a message queuing system called \emph{ActiveQ}.
 Simply said; a message queuing system is a software where queues can be defined, applications may connect to the queue and send or retrieve  messages.
@@
 
-From the image processing point of view, in Diamond, a typical cryoEM session starts by launching a custom made script that performs two actions.
-First, it  allows users to specify values pertinent to their experiment, namely the session ID (or visit number), which microscope they are using, and several other key data collection parameters.
-Once set, a workflow file is created and then saved to disk.
-The file creation triggers the message system \emph{ActiveMQ} and a ``file creation'' message is created.
-The second action performed by the custom made script is to start a \scipion instance in the microscope control computer.
-The instance runs in headless mode. \scipion headless mode has been developed specifically for eBIC and allows \scipion to be executed without user interface in large clusters with no graphical board.
+From the image processing point of view, in Diamond, a typical cryoEM session starts by launching a custom made script that  allows users to specify values pertinent to their experiment, namely the session ID (or visit number), which microscope they are using, and several other key data collection parameters.
+Once set, a file describing a \scipion workflow is created and then saved to disk.
+A daemon that is running on the microscope control server notices the file creation and send a ``file creation'' message to \emph{ActiveMQ}.
+
+Another process listening to \emph{ActiveMQ} then launches \scipion from this workflow and creates a running ``headless'' project.
+\scipion headless mode has been developed specifically for eBIC and allows \scipion to be executed without user interface in large clusters with no graphical board.
 
@@
 
-After \scipion is launched the \protocol{import movie} protocol starts.
+%After \scipion is launched the \protocol{import movie} protocol starts.
 \protocol{Import movie} needs to look for a list of files on the file system with names matching a pattern but, in Diamond distributed file system, this is a very expensive task.
@@
 
-To monitor the progress of the automatic image processing CNB and \scilifelab cryo-EM services make use of \scipion default monitor system.
-In Diamond, the electronic notebook ISPyB is the standard tool to record and show metadata about the different experiments.
+To monitor the progress of the automatic image processing, CNB and \scilifelab cryo-EM services make use of \scipion default monitor system.
+However, in Diamond, the electronic notebook ISPyB is the standard tool to record and show metadata about the different experiments.
 In this way, a new monitor has been developed which, through a webservice, allows \scipion to directly interface with ISPyB.
@@
 
-Although most of developments described in this subsection will be integrated in the main \scipion code repository at present they are available in a developer's branch  at \url{https://github.com/I2PC/scipion/tree/release-1.1-headless-devel}.
+Although most of developments described in this subsection will be integrated in the main \scipion code repository, at present they are available in a developer's branch  at \url{https://github.com/I2PC/scipion/tree/release-1.1-headless-devel}.

### 2018-04-11 3.workflow_scilifelab.tex
carazo's review

@@
 
-At the beginning of 2016, the Cryo-EM facility at \scilifelab became one of the early adopters of \scipion for on the fly data processing.
+At the beginning of 2016, the Cryo-EM facility at \scilifelab became one of the early adopters of \scipion streaming processing.
 Before, there were some home-made scripts to perform motion correction (with motioncorr) and CTF estimation (with Gctf) but the microscope operators recognized the importance of a more general framework.
-In particular, they were interested in the possibility of modify easily the processing workflow as well as provide users and staff with  graphical tools for data analysis and quick feedback regarding the data collection.
+In particular, they were interested in the possibility of modifying easily the processing workflow as well as provide users and staff with  graphical tools for data analysis and quick feedback regarding the data collection.
 Initially, the same setup script used at CNB was adopted, but it did not fully satisfy the facility requirements and a new one has been developed.
@@
 
-In order to avoid interruption of data collection due to network issues each microscope has their own dedicated storage, that in the case of the Titan microscope is double since it may be connected to two different cameras, a K2 and a Falcon III.
+In order to avoid interruption of data collection due to network issues each microscope has its own dedicated storage, that in the case of the Titan microscope is double since it may be connected to two different cameras, a K2 and a Falcon III.
 In this way, after starting the data collection, users launch a script that continuously moves the produced data from the adequate storage to the staging server.
@@
 Unfortunately, it was difficult to reuse the existing code due to a different folder structure and, more important, a different logic about users and microscopes management.
-To address the new set of requirements, a new wizard is under development at \scilifelab.
+%To address the new set of requirements, a new wizard is under development at \scilifelab.
 For this new wizard, a basic data model was designed representing the current processing and data organization at \scilifelab.
@@
 
-When an user starts a data collection, the wizard is also executed using the command \textit{session-titan} (or \textit{session-talos}).
+When a user starts a data collection, the wizard is also executed using the command \textit{session-titan} (or \textit{session-talos}).
 The wizard will fetch the reservations for this day using the webservice API provided by \emph{Booked Scheduler} and will check which user is operating the microscope.

### 2018-04-11 4.future.tex
carazo's review

@@
 
-Currently, \scipion includes the appropriate code (python wrappers) to talk to the integrated EM packages (Xmipp, Eman, RELION, ...) at the time of the release.
-This implies that an update of any of the integrated EM packages immediately after the \scipion release, won't be available until next \scipion release (we are aiming for one release a year).
+Currently, \scipion includes the appropriate code (python wrappers) to talk to the integrated EM packages (Xmipp, Eman, Relion, ...) at the time of the release.
+Therefore, an update of any of the integrated EM packages immediately after the \scipion release won't be available until next \scipion release (we are aiming for one major release per year).
 To decouple EM-package releases from \scipion releases, we are working on making the wrappers for EM packages  independent from \scipion.
-Plans for ensuring this independence include: wrapper installation through pip (\url{https://pip.pypa.io/en/stable/}) and the reimplementation of the wrappers as plug-ins that will be automatically detected and added to the application menu.
+Plans for ensuring this independence include: (1) wrapper installation through pip (\url{https://pip.pypa.io/en/stable/}), (2) reimplementation of the wrappers as plug-ins that will be automatically detected and added to the application menu and (3) develop of standalone IO  and visualization libraries that make \scipion fully independent from other EM packages.
 
@@
 As part of the EOSCPilot project (\url{https://eoscpilot.eu/cryoem}), we are also improving  how \scipion exports workflows in order to ease reporting of the work done.
-The goal is to write a workflow file that fully describes the image processing steps, so that if the same input data is provided to the workflow the same results should be obtained (making the data more compliant with FAIR guidelines given by the \citet{eu2016:fair}).
-This file could go with the raw data acquired by CryoEM facilities and deposited at common EM databases like EMPIAR or EMDB \citep{Patwardhan2016:databasesEM}.
-To facilitate the visualization of the workflow file in any of these databases, we are also developing a  webcomponent (\scipion \emph{workflow viewer}) that will easily allow these repositories to visualize the workflow on their web pages (\url{https://github.com/I2PC/web-workflow-viewer}).
+The goal is to write a workflow file that fully describes the image processing steps, so that if the same input data is provided to the workflow the same results should be obtained.
+In this way we want to make  the data produced by \scipion more compliant with FAIR guidelines given by the \citet{eu2016:fair}.
+This workflow file could go with the raw data acquired by CryoEM facilities and deposited at common EM databases like EMPIAR or EMDB \citep{Patwardhan2016:databasesEM}.
+To facilitate the visualization of the workflow file in any of these databases, we are  developing a  webcomponent (\scipion \emph{workflow viewer}) that will easily allow these repositories to visualize the workflow on their web pages (\url{https://github.com/I2PC/web-workflow-viewer}).
 Finally, we are implementing a workflow repository.

### 2018-04-11 5.conclusions.tex
carazo's review

@@
 Therefore, microscopes tend to belong to centralized facilities where users are allocated microscope time (shifts) to perform a variety of tasks.
-Available shifts are a scarce commodity and therefore users (and staff) need to be able to monitor the quality of the acquisition process and obtain as soon as possible a first glimpse of the reconstructed specimen in order to provide an early detection of any issue.
+Available shifts are a scarce commodity and therefore users (and staff) need to be able to monitor the quality of the acquisition process and obtain, as soon as possible, a first glimpse of the reconstructed specimen in order to provide an early detection of any issue.
 
@@
 \scipion has been designed following a strict workflow paradigm.
-This approach is specially useful for tasks that need to go from a given start (movies) to a defined end (3D map) but there are many different paths (algorithms) and steps to get there.
+This approach is specially useful for tasks that need to go from a given start (movies) to a defined end (3D map), but where there are many different paths (algorithms) and steps to get there.
 
@@
 
- \scipion stream capabilities can handle large and continuously produced volumes of data.
+ \scipion stream capabilities can handle large and continuously produce volumes of data.
 The benefits of streaming are clear: (1) it helps to process data as soon as it is generated,  providing almost instant results on data quality; (2) it saves user processing time at their home institutions since part of the processing is done while data is being acquired; and (3)
  it further reduces the overall experimental time by allowing processing the data while it is transferred to the final user laboratory.
+
  \scipion has been deployed in cloud and there is a  a public image, with all the required preinstalled cryoEM software,  available at Amazon WebServices EC2 and EGI federated clouds.
@@
 Other public software mentioned in this article as \emadmin, \emph{wizard session} and the different scripts used at eBIC are available at \url{https://github.com/rmarabini/webservices/tree/master/EMadmin}, \url{\url{https://github.com/delarosatrevin/scipion-session}} and \url{https://github.com/I2PC/scipion/tree/release-1.1-headless-devel} respectively.
-If you would like to have more information about how to install \scipion or you have suggestions on how to improve it contact us at scipion@cnb.csic.es.
+If you would like to have more information about how to install \scipion or you have suggestions on how to improve it, contact us at scipion@cnb.csic.es.
 

### 2018-04-11 6.acknowledgments.tex
carazo's review

@@
 
-The Spanish Ministry of Economy and Competitiveness through Grants BIO2013-44647-R, BIO2016-76400-R(AEI/FEDER, UE) and AEI/FEDER BFU 2016 74868P, the Comunidad Aut\'{o}noma de Madrid through Grant: S2017/BMD-3817, European Union (EU) and Horizon 2020 through grant CORBEL (INFRADEV-1-2014-1, Proposal: 654248).
+The Spanish Ministry of Economy and Competitiveness through Grants BIO2016-76400-R(AEI/FEDER, UE) and AEI/FEDER BFU 2016 74868P, the Comunidad Aut\'{o}noma de Madrid through Grant: S2017/BMD-3817, European Union (EU) and Horizon 2020 through grant iNEXT (INFRADEV-1-2014-1, Proposal: 654248).
 The ``Knut \& Alice Wallenberg Foundation'', and ``A Pilot Facility development grant from Science for Life Laboratory''.

### 2018-04-18 2.overall.tex
to home

@@
 For example, as movies are imported to the system they are aligned and their CTF is estimated.
-In \scipion version 1.1, users may find  the first steps in the image processing workflow adapted to streaming, including movie alignment and CTF estimation.
-We have also implemented particle picking and extraction, that are going to be released soon in the new version 1.2.
+In \scipion version 1.2, users may find  the first steps in the image processing workflow adapted to streaming, including movie alignment,  CTF estimation and particle picking and extraction.
 In the developers version, several 2D classification algorithms are being adapted for this new way of processing.
@@
 This clear separation between data access and data processing marks an important difference between running a script or using \scipion, since it facilitates a form of computing in which several jobs are executed  concurrently, instead of sequentially.
-Moreover, \scipion workflow engine allows users to modify the workflow on the fly without modifying a single code line.
+Moreover, \scipion workflow engine allows users to modify the workflow on the fly without modifying a single line of code.
 %In practice, the latter follows  a more general and extensible approach that allows users to modify the workflow on the fly without modifying a single code line.
@@
 \subsection{Metaprotocols}
-Since \scipion original publication we have developed many \emph{metaprotocols}, that is, 
+Since the \scipion original publication we have developed many \emph{metaprotocols}, that is, 
 protocols that either check the progress of other protocols or compare the results of equivalent protocols.
@@
 \protocol{CTF-monitor} may raise an alert if the defocus values are above or below certain thresholds and continuously generate HTML files so that users and staff may easily follow the data acquisition and processing either in-house or remotely.
-An example of these reports may be seen at \url{http://nolan.cnb.csic.es/scipionbox/} [**ROB NOTE: select a nice one an complete URL, I need to ask to javi chichon]%At the project's end, a review of the processing history is produced.
+An example of these reports may be seen at \url{http://scipion.cnb.csic.es/scipionbox/lastHTMLReport/}% [**ROB NOTE: select a nice one an complete URL, I need to ask to javi chichon]%At the project's end, a review of the processing history is produced.
 [** URL to report example?]
@@
 For a given logical step (for example, particle picking) these protocols check if the datasets obtained from the same input data using a different algorithm are consistent.
-Continuing with the particle picking example, \protocol{consensus picking} will compare the particles selected for each one of the executed particle-picking algorithms and produce as output a particle set containing those particles selected by most of the algorithms.
+Continuing with the particle picking example, \protocol{consensus picking} will compare, at a given instant, the particles selected for each one of the executed particle-picking algorithms and produce as output a particle set containing those particles selected by most of the algorithms.
 
@@
 In this work we do not recommend a particular workflow  over others since we realize that different specimens, microscopes or facilities may have different requirements.
-Nevertheless, since a use case may clarify \scipion capabilities, in the following we describe the pipeline offered by default at the \emph{Centro Nacional de Biotecnología} (CNB) cryoEM Facility.
+Nevertheless, since a use case may clarify \scipion capabilities, in the following we describe the pipeline offered by default at the \emph{Centro Nacional de Biotecnolog\'{i}a} (CNB) cryoEM Facility.
 
@@
 The second protocol performs direct detector movie alignment.
-Currently \scipion offers  four  methods for movie alignment: motioncor2 \citep{Zheng2016:motioncor2}, correlation (an unpublished CPU version that implements the global alignment method used by motioncor \citep{Li2015}), unblur \citep{Grant2015:unblur} and optical flow \citep{Abrishami2015:optical-flow}.
+Currently \scipion offers  five  methods for movie alignment: motioncorr \cite{Li2015}, motioncor2 \citep{Zheng2016:motioncor2}, correlation (an unpublished CPU version that implements the global alignment method used by motioncor \citep{Li2013:motioncorr}), unblur \citep{Grant2015:unblur} and optical flow \citep{Abrishami2015:optical-flow}.
 Then the CTF of the resulting micrographs are calculated.

### 2018-04-18 3.workflow.tex
to home

@@
-\section{\scipion Setup at the Different Facilities}
+\section{\scipion Setup at Different Facilities}
 
@@
  \item \emph{CNB CryoEM Service} (\url{http://www.cnb.csic.es/index.php/en/research/core-facilities/microscopia-crioelectronica})
- \item \emph{Molecular Microscopy Consortium - NIH} (\url{https://www.niehs.nih.gov/research/atniehs/facilities/mmc/index.cfm})[**Pablo can you confirm this?]
- \item \emph{National Cancer Institute - NIH} (\url{https://cmm.nci.nih.gov/})[**Pablo can you confirm this?]
+ \item \emph{Molecular Microscopy Consortium - NIH} (\url{https://www.niehs.nih.gov/research/atniehs/facilities/mmc/index.cfm})
+ \item \emph{National Cancer Institute - NIH} (\url{https://www.cancer.gov/research/resources/cryoem})
 
@@
 There, \scipion was adapted to interface with this booking platform in order to produce better reports.
-Lately,  \scipion has been installed in  large synchrotrons  such as ESRF and Diamond  where projects are handled by ISPyB (a customized laboratory information management system -LIMS- \citep{Delageniere2011}) and data are saved on a distributed file system.
+Lately,  \scipion has been installed in  large synchrotrons  such as ESRF and Diamond  where projects are handled by ISPyB (a customized laboratory information management system -LIMS- \citep{Delageniere2011:ispb}) and data are saved on a distributed file system.
 % that heavily penalizes access to lists of files.

### 2018-04-18 3.workflow_cnb.tex
to home

@@
 Nevertheless, real-world is tough and it happens that two, or even three,  users try to download their data simultaneously.
-Fortunately, output-NAS works as front-end and isolates the data acquiring and processing infrastructure from the user downloading.
+Fortunately, output-NAS works as front-end and isolates the data acquiring and processing infrastructure from the user's downloading activities.
 Network users access to output-NAS using a chrooted common account.
@@
 
-\emadmin was born as a script that created a normalized tree of directories  in which to store the movies acquired by the microscope, helped users to choose between a few predetermined image processing workflows and finally launched \scipion.
-At present it has grown in a client-server application that in addition to the old script capacities it has incorporate the following functions:
+\emadmin was born as a script that created a normalized tree of directories  in which to store the movies acquired by the microscope, helping users to choose between a few predetermined image processing workflows and finally launched \scipion.
+At present it has grown into a web application that in addition to the old script capacities it has incorporate the following functions:
 
@@
  \item Automatic backup of the \scipion project
- \item \scipion is able to create live html reports on the image acquisition process. \emadmin keeps track of all of them and facilitates the access to this information
+ \item \scipion is able to create live HTML reports on the image acquisition process. \emadmin keeps track of all of them and facilitates the access to this information
 \end{enumerate}

### 2018-04-18 3.workflow_diamond.tex
to home

@@
-\subsection{Case 3: Scipion at eBIC facility (Diamond Light Source)}
+\subsection{Case 3: Scipion at eBIC (Diamond Light Source)}
 
@@
 Visits are structured such that all raw and processed data collected during an experiment are kept in one directory, and only the associated visitors have read and write permissions in that directory.
-Users may access their data through ISPyB \citep{ispb2011},  a Laboratory Information Management System (LIMS) combining sample tracking and experiment reporting  %The link between the user and their visits enables data management such as archiving data to a long-term tape storage system, or the Laboratory Information Management System ISPyB (REF.
+Users may access their data through ISPyB \citep{Delageniere2011:ispb},  a Laboratory Information Management System (LIMS) combining sample tracking and experiment reporting  %The link between the user and their visits enables data management such as archiving data to a long-term tape storage system, or the Laboratory Information Management System ISPyB (REF.
 Delageniere et al Bioinformatics 2011).
@@
 However, in Diamond, the electronic notebook ISPyB is the standard tool to record and show metadata about the different experiments.
-In this way, a new monitor has been developed which, through a webservice, allows \scipion to directly interface with ISPyB.
+In this way, a new monitor has been developed which, through a ISPyB python API, allows \scipion to directly interface with ISPyB.
 This setup provides the users with a web interface showing general metadata about their experiment as well as  real-time updates on steps of the process and relevant output such as motion corrected images, drift values, CTF images, etc.

### 2018-04-18 3.workflow_scilifelab.tex
to home

@@
-\subsection{Case 2: Scipion at The Swedish National Cryo-EM Facility}
+\subsection{Case 2: Scipion at the Swedish National Cryo-EM Facility}
 
@@
 At the beginning of 2016, the Cryo-EM facility at \scilifelab became one of the early adopters of \scipion streaming processing.
-Before, there were some home-made scripts to perform motion correction (with motioncorr) and CTF estimation (with Gctf) but the microscope operators recognized the importance of a more general framework.
+Before, there were some home-made scripts to perform motion correction (with motioncor2) and CTF estimation (with Gctf) but the microscope operators recognized the importance of a more general framework.
 In particular, they were interested in the possibility of modifying easily the processing workflow as well as provide users and staff with  graphical tools for data analysis and quick feedback regarding the data collection.

### 2018-04-18 4.future.tex
to home

@@
 
-Additionally, in collaboration with the Xmipp team, we are pushing the streaming workflow with the aim to obtain a first initial volume.
-We are also adding structure modeling capabilities, by integration of some functionality from programs such as Coot \citep{emsley2010:coot}, Refmac \citep{Murshudov1997:refmac}, etc.
+Additionally, in collaboration with the Xmipp team, we are pushing the streaming workflow with the aim of obtaining a first initial volume.
+We are also adding structure modeling capabilities, by integrating some functionality from programs such as Coot \citep{emsley2010:coot}, Refmac \citep{Murshudov1997:refmac}, etc.
 

### 2018-04-18 5.conclusions.tex
to home

@@
 
- \scipion stream capabilities can handle large and continuously produce volumes of data.
-The benefits of streaming are clear: (1) it helps to process data as soon as it is generated,  providing almost instant results on data quality; (2) it saves user processing time at their home institutions since part of the processing is done while data is being acquired; and (3)
+ \scipion stream capabilities can handle large and continuously produced volumes of data.
+The benefits of streaming are clear: (1) it helps to process data as soon as it is generated,  providing almost instant results on data quality; (2) it saves user processing time and resources at her home institutions since part of the processing is done while data is being acquired; and (3)
  it further reduces the overall experimental time by allowing processing the data while it is transferred to the final user laboratory.
 
- \scipion has been deployed in cloud and there is a  a public image, with all the required preinstalled cryoEM software,  available at Amazon WebServices EC2 and EGI federated clouds.
+ \scipion has been deployed in cloud and there is a public image, with all the required preinstalled cryoEM software,  available at Amazon WebServices EC2 and EGI federated clouds.
 These images make use of \scipion streaming capabilities, that is, they are able to process data as soon as it has been transfered to an Amazon storage unit, and can be used to handle  production peaks.

### 2018-04-18 0.abstract.tex
from home Marta (swedden) comments

@@
 
-Three dimensional electron microscopy (3DEM) is becoming a very data-intensive field in which vast amounts of experimental images are acquired at high speed.
+Three dimensional electron microscopy is becoming a very data-intensive field in which vast amounts of experimental images are acquired at high speed.
 To manage such large-scale projects, we had previously developed  a modular workflow system called \scipion \citep{delarosatrevin2016}.
 We present here a major extension of \scipion that allows stream processing of EM images while the data is being acquired.
-This approach helps to detect problems at early stages, saves computing time and provides users with an in deep evaluation of the data quality before the acquisition is finished.
+This approach helps to detect problems at early stages, saves computing time and provides users with a detailed evaluation of the data quality before the acquisition is finished.
 Different EM facilities have
@@
 
-\textbf{Keywords:} electron microscopy
+\textbf{Keywords:} electron microscopy, streaming, image processing, live processing, high throughput
 

### 2018-04-18 1.introduction.tex
to home
from home Marta (swedden) comments

@@
 3D electron microscopy (3DEM)  can provide rich information about structural characteristics of macromolecular complexes.
-The field is under a major transformation due to the arrival of better microscopes, new large area direct electron detectors and automatization \citep{kuhlbrandt2014a, Kuhlbrandt2014b}.
+The field is under a major transformation due to the arrival of better microscopes, new large area direct electron detectors and automation \citep{kuhlbrandt2014a, Kuhlbrandt2014b}.
 These improvements make possible that a single microscope generates high quality data sets on the order of terabytes per day \citep{Saibil2015} while working for several days without interruption.
@@
 For most cases, the future of projects that require high-end microscopes is through centralized microscopy facilities, in which automatic acquired data needs to be monitored by users and facility staff.
-\scipion has been redesigned with this scenario in mind.
+With this scenario in mind, we have redesigned \scipion which is a software framework for integrating several 3DEM software packages through a workflow-based approach \citep{delarosatrevin2016}.
 % in which the early detection of issues may save a data acquisition shift which otherwise needs to be repeated after a new application that may be granted many months later.
-In this way, \scipion starts the processing of the movies as they are being acquired with the double aim of checking for possible data collection errors and execute a first user-tailored image processing workflow that provides users with an accurate measurement of the acquired data quality.
+In this way, \scipion starts the processing of the movies as they are being acquired with the double aim of checking for possible data collection errors and execute a first user-tailored image processing workflow that provides users with an accurate estimation of the acquired data quality.
 

### 2018-04-18 3.workflow_cnb.tex
from home Marta (swedden) comments

--- 2018-04-11 3.workflow_cnb.tex
@@
 
-At the beginning CNB networking setup was a quite straight forward one, movies where stored in the  same machine that ran \scipion and processed data was saved to external USB hard disks by this same machine.
+At the beginning CNB networking setup was a quite straight forward one, movies where stored in the  same machine that ran \scipion and processed data were saved to external USB hard disks by this same machine.
 The situation changed when the new Falcon-III detector was installed.
-After that, data was produced at sustained speeds as high as 90 MB/sec (when acquiring four movies by hole in lineal mode) and the main problem became to move smoothly the data through the entire pipeline from the microscope to the final user laboratory.
+After that, data were produced at sustained speeds as high as 90 MB/sec (when acquiring four movies by hole in lineal mode) and the main problem became to move smoothly the data through the entire pipeline from the microscope to the final user laboratory.
 In the following we describe such a pipeline.
@@
 
-Although CNB staff try to discourage the use of external USB hard disks versus network as way of retrieving the raw and processing data, this is the preferred option for most users.
+Although CNB staff try to discourage the use of external USB hard disks versus network as way of retrieving the raw and processed data, this is the preferred option for most users.
 USB 3.0 drives can theoretically handle data flows of 300 MB/s, but CNB experience is that, in this busy environment, many of them  have trouble exceeding 50 MB/s.
@@
 
-\emadmin was born as a script that created a normalized tree of directories  in which to store the movies acquired by the microscope, helped users to choose between a few predetermined image processing workflows and finally launched \scipion.
-At present it has grown in a client-server application that in addition to the old script capacities it has incorporate the following functions:
+\emadmin was born as a script that created a normalized tree of directories  
+where movies acquired by the microscope could be stored, helped users to choose between a few predetermined image processing workflows and finally launched \scipion.
+At present it has grown in a client-server application that in addition to the old script capacities has incorporated the following functions:
 

### 2018-04-18 3.workflow_diamond.tex
from home Marta (swedden) comments

--- 2018-04-11 3.workflow_diamond.tex
@@
 One of the solutions to this problem has been to implement \emph{indirect communication} through a message queuing system called \emph{ActiveQ}.
-Simply said; a message queuing system is a software where queues can be defined, applications may connect to the queue and send or retrieve  messages.
+A message queue is a mechanism
+that allows a sender process and a receiver process to exchange information; instead of connecting directly, the sender posts messages in a queue and the receiver, when ready, takes these messages from the queue and process them appropriately.
+
+%In this environment any piece of information or data that needs to be interchanged among  two applications will be placed in a queue by the sender application until the destination application explicitly asks for it.
+
+%Simply said; a message queuing system is a software where queues can be defined, applications may connect to the queue and post or retrieve  messages.
 A message can include any kind of information, for example, about a process/task that should start on another application (that could be on another server).

### 2018-04-18 5.conclusions.tex
from home Marta (swedden) comments

--- 2018-04-11 5.conclusions.tex
@@
  In a workflow approach, users (including microscope staff) can reasonably be expected to edit the paths.
-They can potentially play with the pipeline in ways that aren't possible when having a programmer write code for them.
-Furthermore, not having the processing path hardcoded in the program means you don't have to recompile, so changes can be quickly tested   even at run time.
+They can potentially play with the pipeline in ways that are not possible when having a programmer writing the code for them.
+Furthermore, not having the processing path hardcoded in the program means you do not have to recompile, so changes can be quickly tested   even at run time.
 The \scipion workflow approach provides provenance, that is,  for any created object (micrograph, CTFs, class, etc) \scipion stores metadata such as which programs, which parameters, and which exact path was followed to create any output data.

### 2018-04-18 scipionbox.tex
carazo's review
to home
from home Marta (swedden) comments

@@
 \begin{document}
-\title{Using Scipion for Stream Image Processing at EM Facilities}
+\title{Using Scipion for Stream Image Processing at Electron Microscopy Facilities}
 Authors to be sorted later
@@
 \setcounter{figure}{2}%I do not know why but figure counter starts with -1
-\textbf{Corresponding Author:} %Roberto Marabini, Ph= 34 91 5854510, email: roberto@cnb.csic.es\\
+\textbf{Corresponding Author: Roberto Marabini, Ph = +34 91 5854510, email: roberto@cnb.csic.es\\} %Roberto Marabini, Ph= 34 91 5854510, email: roberto@cnb.csic.es\\
 %Escuela Polit\'{e}cnica Superior\\

### 2018-04-19 3.workflow_scilifelab.tex
from home Marta (swedden) comments
Stefan

@@
 \subsubsection{Network Setup and IT Infrastructure}
-At the heart of the \scilifelab computational infrastructure there is a storage and pre-processing server (the staging server).
+At the heart of cryo-EM National Facility computational infrastructure there is a storage and pre-processing server (the staging server).
 This machine has roughly 200 TB of storage (4 ZFS RAIDZ2 pools with 11 HDDs each), 2 NVIDIA GeForce GTX 1070 GPUs, 2 Intel Xeon E5-2630v4 CPUs (10 cores, 2.2 GHz, HT), 384 GB RAM, and a dual-port 10 GbE network card.
@@
 In order to avoid interruption of data collection due to network issues each microscope has its own dedicated storage, that in the case of the Titan microscope is double since it may be connected to two different cameras, a K2 and a Falcon III.
-In this way, after starting the data collection, users launch a script that continuously moves the produced data from the adequate storage to the staging server.
+Hence an script to continuously move the produced data from the adequate storage to the staging server has been created.
+
+%In this way, after starting the data collection, users launch a script that continuously moves the produced data from the adequate storage to the staging server.
 
@@
 After processing, users may  copy their data to external USB drives from two workstations in the microscope room that have access to the storage pool via NFS.
-There is also a dedicated download server which have access to the NFS export. This download server can be reached though the Internet and allow downloads via rsync (over SSH) or SFTP.
+There is also a dedicated download server which has access to the NFS export. This download server is accessible from  the Internet and allow downloads via rsync (over SSH) or SFTP.
 
-All these computers are connected to a 10 GbE network switch. This is a private
-network without external access. The only machine accessible from the Internet
-is the data download server, which has a dual-port 10 GbE network adapter (one
-used for the private network, one for the external network).
+All these computers are connected to a 10 GbE network switch.
+The only machine accessible from the Internet is the data download server, which has a dual-port 10 GbE network adapter (one used for the internal network, one for the external network).
 %Both the K2 computer and the Falcon-III dedicated storage server had one unused 10GbE interface that could be used to connect them to our private network, without having to interfere with the default setup.
@@
 \subsubsection{Additional Software Developments: \emph{Session Wizard}}
-In order to interface \scipion with the microscope infrastructure an upgraded version of the original script developed at the CNB was used.
-Unfortunately, it was difficult to reuse the existing code due to a different folder structure and, more important, a different logic about users and microscopes management.
+In order to interface \scipion with the microscope infrastructure an upgraded version of the original script developed at CNB was used.
+Unfortunately, it was difficult to reuse the existing code due to a different folder structure and, more important, a different approach to user and microscope management.
 %To address the new set of requirements, a new wizard is under development at \scilifelab.
@@
 
-When a user starts a data collection, the wizard is also executed using the command \textit{session-titan} (or \textit{session-talos}).
-The wizard will fetch the reservations for this day using the webservice API provided by \emph{Booked Scheduler} and will check which user is operating the microscope.
-From the reservation information, the wizard will also detect if it is an internal or a national facility project.
-If the latter, more information will be retrieved from the Application Portal (such as principal investigator, project code, contact person or invoice address).
-After gathering all this information, the wizard will store this information, create the processing environment and launch \scipion.
+When a user starts a data collection, the wizard is executed using the command \textit{session-titan} (or \textit{session-talos}).
+The wizard determines which user has booked the microscope by fetching today's booking information from \emph{Booked Scheduler} through their web API.
+From the reservation information the wizard will also detect if it is an internal or a national facility project.
+In case of the latter, additional information will be retrieved from the Application Portal (such as principal investigator, project code, contact person or invoice address).
+After gathering all this information the wizard will store it, create the processing environment and launch \scipion.
 %For example, the movies-import protocol will already point to the location where the files will be written for this session.
@@
 The current implementation of the wizard stores each session in a simple database file.
-This database, together with extra information from the Booking System and the Application Portal, is used to generate reports for a given period and the associated invoices.
+This database, together with extra information from the Booking System and the Application Portal, is used to generate invoices and reports for a given period.
 %The generation of a session report is under development, that will provide a summary and useful information for users to access the acquired data.
@@
 The described session wizard can be found at  \url{https://github.com/delarosatrevin/scipion-session}.
-It has been design to satisfy \scilifelab requirements and, although readers are welcomed to try it, no support is offered.
+It has been designed to satisfy the requirements for \scilifelab and, although readers are welcomed to try it, no support is offered.
+\emph{Booked Scheduler} is available at (\url{https://www.bookedscheduler.com/})
 

### 2018-04-19 0.abstract.tex
middle coss

@@
 The advantages of our proposed solution are: (1) flexibility: users can customize their workflows from a  large collection of preloaded algorithms; (2) repeatability and traceability: workflows are stored and can be re-executed later, either with the same data or with a different data set; (3) monitorization: there are utilities that constantly check how the execution of the algorithms is going on and inform about possible anomalies; (4) reports: we have developed several graphical interfaces that are updated periodically producing a summary (e.g, CTF defocus values, system load, etc).
-This summary may be generated in HTML format and  published though a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided.
-At present, \scipion has been installed and is in production mode in seven Cryo-EM facilities and it is being considered in many more.
+This summary may be generated in HTML format and  published through a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided.
+At present, \scipion has been installed and is in production mode in seven Cryo-EM facilities throughout the world and, it is being considered in many more.
 
-\textbf{Keywords:} electron microscopy, streaming, image processing, live processing, high throughput
+\textbf{Keywords:} electron microscopy, streaming, image processing, live processing, high throughput, Scipion
 

### 2018-04-19 1.introduction.tex
middle coss

@@
 
-For most cases, the future of projects that require high-end microscopes is through centralized microscopy facilities, in which automatic acquired data needs to be monitored by users and facility staff.
+For most cases, the future of projects that require high-end microscopes is through centralized microscopy facilities, in which automatically acquired data needs to be monitored by users and facility staff.
 With this scenario in mind, we have redesigned \scipion which is a software framework for integrating several 3DEM software packages through a workflow-based approach \citep{delarosatrevin2016}.
@@
 However, the next step,  which is to provide image processing while  data are being collected, is not so well addressed.
-In fact, in most places this is accomplished through home made scripts \citep[i.e.][]{Pichkur2018}.
+In fact, in most places this is accomplished through home made scripts \citep[e.g.][]{Pichkur2018}.
 %(i.e.
 \citet{Pichkur2018}).
-Exceptions are UCSFImage4 \citep{Li2015} and Focus \citep{Biyani2017}.
+Exceptions are UCSFImage4 \citep{Li2015}, Focus \citep{Biyani2017} and Relion \citep{fernandez2017:relion-pipeline}.
 As we will show in this article, \scipion differs from these solutions in a larger flexibility, both from the point of view of users and facility staff.

### 2018-04-19 2.overall.tex
middle coss

@@
 \subsection{Stream Processing}
-The main change introduced in \scipion is \emph{stream processing}, that is, computing on data directly as it is produced or received.
+The main change introduced in \scipion is \emph{stream processing}, that is, computing on data directly as they are produced or received.
 In this way, a new level of parallelization has been reached since the  data generated by the microscope continuously flows through a chain of protocols.
@@
 
-We would like to finish this section stressing that despite the modifications introduced to \scipion data and process model, the framework keeps providing full data provenance.
+We would like to finish this section stressing that despite the modifications introduced to \scipion data and process model, the framework keeps providing full data provenance and backward compatibility.
 That is, \scipion provides detailed information about the origin of the data  and the process to generate it.
@@
 In this work we do not recommend a particular workflow  over others since we realize that different specimens, microscopes or facilities may have different requirements.
-Nevertheless, since a use case may clarify \scipion capabilities, in the following we describe the pipeline offered by default at the \emph{Centro Nacional de Biotecnolog\'{i}a} (CNB) cryoEM Facility.
+Nevertheless, since a use case may clarify \scipion capabilities, in the following we describe the pipeline offered by default at the \emph{National Center for Biotechnology} (CNB) cryoEM Facility.
 
@@
 The second protocol performs direct detector movie alignment.
-Currently \scipion offers  five  methods for movie alignment: motioncorr \cite{Li2015}, motioncor2 \citep{Zheng2016:motioncor2}, correlation (an unpublished CPU version that implements the global alignment method used by motioncor \citep{Li2013:motioncorr}), unblur \citep{Grant2015:unblur} and optical flow \citep{Abrishami2015:optical-flow}.
+Currently \scipion offers  five  methods for movie alignment: motioncorr \citep{Li2015}, motioncor2 \citep{Zheng2016:motioncor2}, correlation (an unpublished CPU version that implements in Xmipp a global alignment method and solves it similarly to motioncorr but using robust statistics), unblur \citep{Grant2015:unblur} and optical flow \citep{Abrishami2015:optical-flow}.
 Then the CTF of the resulting micrographs are calculated.
@@
 Two new streaming protocols are under development and will join the described  pipeline in a near future.
-The first one will perform 2D classification and the second one will create a initial 3D map based on this classes.
+The first one will perform 2D classification and the second one will create a initial 3D map based on these classes.
 A first version of these protocols is available in the \scipion developers version.
 
-As commented in section \ref{overall}, to monitor, chart and create alerts  we have developed a set of metaprotocols.
-By default \protocol{system}, \protocol{CTF} and \protocol{report monitors} are executed.
-\protocol{System monitor} reports on  CPU and GPU usage, memory consumption and, disk I/O.
+As commented in Section \ref{overall}, to monitor, chart and create alerts  we have developed a set of metaprotocols.
+By default, \protocol{system}, \protocol{CTF} and \protocol{report monitors} are executed.
+\protocol{System monitor} reports on  CPU and GPU usage, memory consumption, and disk I/O.
 \protocol{CTF monitor} plots the CTF defocus and astigmatism  and it will raise an alarm if these parameters are not within a given range.

### 2018-04-19 3.workflow_cnb.tex
middle coss

@@
 The National Center for Biotechnology (CNB) forms part of the Spanish National Research Council (CSIC), the largest public research institution in Spain. 
-The CNB Cryo-Electron Microscopy Facility is a joint effort of the Centro Nacional de Biotecnolog\'{i}a (CNB-CSIC) and the Centro de Investigaciones Biol\'{o}gicas (CIB-CSIC).
+The CNB Cryo-Electron Microscopy Facility is a joint effort of the Nacional Center for Biotechnology (CNB-CSIC) and the Biological Research Centers (CIB-CSIC).
 The mission of the facility is to provide researchers access to the latest cryo-EM technology for high resolution imaging.
@@
 
-At the beginning CNB networking setup was a quite straight forward one, movies where stored in the  same machine that ran \scipion and processed data were saved to external USB hard disks by this same machine.
+At the beginning CNB networking setup was a quite straightforward one, movies where stored in the  same machine that ran \scipion and processed data were saved to external USB hard disks by this same machine.
 The situation changed when the new Falcon-III detector was installed.
@@
 \scipionbox has 32 CPUs at 2.40GHz, 62 Gb of RAM memory and 2 \textit{Quadro M4000} GPUs.
-Finally, a second NAS with 4 1Gb NICs is linked to the private, LAN and WAN networks (we will refer to this second NAS as output-NAS).
+Finally, a second NAS with four 1Gb NICs is linked to the private, LAN and WAN networks (we will refer to this second NAS as output-NAS).
 
 The data flow is a follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox will temporary read the movies from 
-the microscope-NAS but will not store them locally, (c) each hour, output-NAS will connect to \scipionbox and copy -or update- the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup \scipion project and the movies to a USB disk or serve this information through the Internet and, (e) in parallel, \scipionbox transfers to  output-NAS different html reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox}).
+the microscope-NAS but will not store them locally, (c) each hour, output-NAS will connect to \scipionbox and copy -or update- the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup \scipion project and the movies to a USB disk or serve this information through the Internet and, (e) in parallel, \scipionbox transfers to  output-NAS different HTML reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox}).
 
@@
 This account only allows rsync connections and make impossible to retrieve data if the name of the project to be downloaded is unknown.
-In order to speed up transfer weak ciphers (as \emph{arcfour}) has been activated in output-NAS.
+In order to speed up transfer, weak ciphers (as \emph{arcfour}) has been activated in output-NAS.
 
@@
  \item Automatic creation of pdf user's report containing: microscope acquisition parameters and histograms summarizing data resolution and  defocus. 
- \item The above mentioned information is stored in a database for further analysis. By default the system plots the  average resolution and astigmatism of the different acquisitions versus time
- \item Remote management. \emadmin is a client server application, the client maybe executed in a machine different from the one that executes \scipion
- \item Automatic backup of the \scipion project
- \item \scipion is able to create live html reports on the image acquisition process. \emadmin keeps track of all of them and facilitates the access to this information
+ \item The above mentioned information is stored in a database for further analysis. By default the system plots the  average resolution and astigmatism of the different acquisitions versus time.
+ \item Remote management. \emadmin is a client server application, the client maybe executed in a machine different from the one that executes \scipion.
+ \item Automatic backup of the \scipion project.
+ \item \scipion is able to create live HTML reports on the image acquisition process. \emadmin keeps track of all of them and facilitates the access to this information.
 \end{enumerate}

### 2018-04-19 3.workflow_diamond.tex
middle coss

@@
 Visits are structured such that all raw and processed data collected during an experiment are kept in one directory, and only the associated visitors have read and write permissions in that directory.
-Users may access their data through ISPyB \citep{Delageniere2011:ispb},  a Laboratory Information Management System (LIMS) combining sample tracking and experiment reporting  %The link between the user and their visits enables data management such as archiving data to a long-term tape storage system, or the Laboratory Information Management System ISPyB (REF.
+Users may access their data through ISPyB \citep{Delageniere2011:ispb},  a Laboratory Information Management System (LIMS) combining sample tracking and experiment reporting.
+%The link between the user and their visits enables data management such as archiving data to a long-term tape storage system, or the Laboratory Information Management System ISPyB (REF.
 Delageniere et al Bioinformatics 2011).

### 2018-04-19 scipionbox.tex
Stefan
middle coss

@@
 \begin{document}
-\title{Using Scipion for Stream Image Processing at Electron Microscopy Facilities}
+\title{Using Scipion for Stream Image Processing at Cryo-EM Facilities}
 Authors to be sorted later
-\author{C.O.S. Sorzano$^{(a)}$,
-A. Jim\'{e}nez$^{(a)}$,
-R. Melero$^{(a)}$,\\
-M. Mart\'{i}nez$^{(a)}$,
-J. G\'{o}mez-Blanco$^{(a)}$,
-L. del Cano$^{(a)}$,\\
-T. Majtner$^{(a)}$,
-E. Ram\'{i}rez$^{(a)}$,
-P. Conesa$^{(a)}$,\\
-J.L. Vilas$^{(a)}$,
+\author{C.O.S. Sorzano$^{(b)}$,
+A. Jim\'{e}nez$^{(b)}$,
+R. Melero$^{(b)}$,\\
+M. Mart\'{i}nez$^{(b)}$,
+J. G\'{o}mez-Blanco$^{(b)}$,
+L. del Cano$^{(b)}$,\\
+T. Majtner$^{(b)}$,
+E. Ram\'{i}rez$^{(b)}$,
+P. Conesa$^{(b)}$,\\
+J.L. Vilas$^{(b)}$,
 R. Marabini$^{(a)}$,
-Y. Rancel$^{(a)}$,
-J.M. Carazo$^{(a)}$,\\
+Y. Rancel$^{(b)}$,
+J.M. Carazo$^{(b)}$,\\
 J.M. de la Rosa-Trev\'{i}n$^{(c)}$,
 M. Carroni$^{(c)}$,
-S. Fleichman $^{(c)}$,\\
+S. Fleischmann $^{(c)}$,\\
 E. Lindahl$^{(c,d)}$
@@
 C.A. Siebert$^{(e)}$,\\
-D. Maluenda$^{(a)}$
+D. Maluenda$^{(a)}$,\\
+Grigory G. Sharov$^{(f)}$
         }
@@
 \lyxaddress{{\small$^{(d)}$ Swedish e-Science Research Center, KTH Royal Institute of Technology, Stockholm, Sweden}}
-\lyxaddress{{\small$^{(e)}$ Diamond Light Source, Harwell Science and Innovation Campus, Didcot OX11 0DE, England}}
+\lyxaddress{{\small$^{(e)}$ Diamond Light Source, Harwell Science and Innovation Campus, Didcot OX11 0DE, United Kingdom}}
+\lyxaddress{{\small$^{(f)}$ MRC Laboratory of Molecular Biology - Structural studies, Cambridge, England, United Kingdom}}
+
 %This document provides a preliminary analysis of the data collected by the CTF Challenge (http://i2pc.cnb.csic.es/3dembenchmark/LoadHome.htm).

### 2018-04-23 1.introduction.tex
I think this is done except for the cover letter

@@
 These improvements make possible that a single microscope generates high quality data sets on the order of terabytes per day \citep{Saibil2015} while working for several days without interruption.
-Major challenges faced in the field are: (1) an efficient management of these huge datasets and their corresponding image processing workflows; (2) a raising interest from scientists outside the field who might lack the skills of a experienced microscopist and; (3) the need for an effective data and metadata transfer in an increasingly distributed and collaborative environment.
+Major challenges faced in the field are: (1) an efficient management of these huge datasets and their corresponding image processing workflows; (2) the need for sharing data and metadata  in an increasingly distributed and collaborative environment and; (3) a raising interest from scientists outside the field who might lack the skills of a experienced microscopist.
 %This situation strongly demands the application of techniques that are highly automated to allow high throughput.
@@
 % in which the early detection of issues may save a data acquisition shift which otherwise needs to be repeated after a new application that may be granted many months later.
-In this way, \scipion starts the processing of the movies as they are being acquired with the double aim of checking for possible data collection errors and execute a first user-tailored image processing workflow that provides users with an accurate estimation of the acquired data quality.
+In this way, \scipion starts the processing of the movies as they are being acquired with the double aim of checking for possible data collection errors and executing a first user-tailored image processing workflow that provides users with an accurate estimation of the acquired data quality.
 
-Many software packages have been developed to provide automated data collection, such as SerialEM \citep{Mastronarde2005}, Leginon \citep{Suloway2009}, UCSF-Tomography \citep{Zheng2007}, Tom Toolbox \citep{Nickell2005}, EPU \citep{EPU}, Latitude in Digital Micrograph \citep{Latitude}, EMMenu \citep{emmenu}, etc.
+Many software packages have been developed to support automated data collection, such as SerialEM \citep{Mastronarde2005}, Leginon \citep{Suloway2009}, UCSF-Tomography \citep{Zheng2007}, Tom Toolbox \citep{Nickell2005}, EPU \citep{EPU}, Latitude in Digital Micrograph \citep{Latitude}, EMMenu \citep{emmenu}, etc.
 However, the next step,  which is to provide image processing while  data are being collected, is not so well addressed.

### 2018-04-23 2.overall.tex
I think this is done except for the cover letter

@@
 
-In the following we describe the most important modifications made to \scipion in order to facilitate its deployment and execution in centralized facilities.
+In the following we describe the most important modifications made to \scipion to facilitate its deployment and execution in centralized facilities.
 
 \subsection{Stream Processing}
-The main change introduced in \scipion is \emph{stream processing}, that is, computing on data directly as they are produced or received.
+
+Stream processing, i.e. computing on data directly as they are produced or received, is the main change introduced in \scipion. 
 In this way, a new level of parallelization has been reached since the  data generated by the microscope continuously flows through a chain of protocols.
@@
 \scipion stream implementation make extensive use of threads. %A thread is the smallest unit of processing that can be scheduled by an operating system. 
-Using threads, a programmer can effectively create two or more tasks that run at the same time and share the protocol resources (i.e. memory). 
+Using threads, a programmer can effectively create two or more tasks that run at the same time and share the protocol resources (e.g. memory). 
 %is based on periodic updates of the protocol input and output data during the image processing. 
@@
 
-The second set of metaprotocols are grouped under the name \emph{consensus}.
-For a given logical step (for example, particle picking) these protocols check if the datasets obtained from the same input data using a different algorithm are consistent.
+The second set of metaprotocols are grouped under the name \emph{consensus}. For a given logical step (for example, particle picking) these protocols check 
+the consistency of the output datasets obtained from the same input data using different algorithms.
 Continuing with the particle picking example, \protocol{consensus picking} will compare, at a given instant, the particles selected for each one of the executed particle-picking algorithms and produce as output a particle set containing those particles selected by most of the algorithms.
 
-We would like to finish this section stressing that despite the modifications introduced to \scipion data and process model, the framework keeps providing full data provenance and backward compatibility.
-That is, \scipion provides detailed information about the origin of the data  and the process to generate it.
+We would like to finish this section stressing that despite the modifications introduced to \scipion data and process model, the framework is backward compatible and keeps providing full data provenance.
+In other words, \scipion provides detailed information about the origin of the data  and the process to generate it.
 
@@
 \scipion is a flexible framework that allows to create different workflows by choosing among many algorithms at each step.
-In this work we do not recommend a particular workflow  over others since we realize that different specimens, microscopes or facilities may have different requirements.
-Nevertheless, since a use case may clarify \scipion capabilities, in the following we describe the pipeline offered by default at the \emph{National Center for Biotechnology} (CNB) cryoEM Facility.
+Although the best workflow for each case will depend on specific requirements of specimens, microscopes or facilities, we  described the pipeline offered by default at the \emph{National Center for Biotechnology} (CNB) cryoEM Facility as use case to clarify Scipion capabilities.
 
@@
 CTFFIND4 (Rohou and Grigorieff, 2015), gCTF (Zhang, 2016) and Xmipp CTF estimation (Sorzano et al., 2007; Vargas et al., 2013).
-At this point the user may compute the CTFs using several algorithms and compare them with the protocol \protocol{CTF consensus} that reports up to which frequency two estimation of the same micrograph are equivalent.
+At this point the user may compute the CTFs using several algorithms and compare them with the protocol \protocol{CTF consensus} that reports up to which frequency two estimations of the same micrograph are equivalent.
 Further automatic filtering based on CTF is available through the protocol \protocol{CTF filtering} that selects micrographs based on the astigmatism, resolution and defocus.
@@
 The next step is to select (pick) particles.
-We have recently adapted eight particle picking methods in \scipion  to streaming.
+We have recently adapted eight particle picking methods in \scipion  for streaming (see at \url{https://github.com/I2PC/scipion/wiki/Integrated-Protocols#particles} an exhaustive list).
 In this way,  particles may be selected as soon as any new micrograph is available.
@@
 For those users that execute several picking algorithms, the protocol \protocol{picking consensus} may be useful.
-\protocol{picking consensus}  calculates the agreement between different particle-picking algorithms and produces a list with the particle selected by most of them.
+\protocol{picking consensus}  calculates the agreement between different particle-picking algorithms and produces a list with the particles selected by most of them.
 
 Two new streaming protocols are under development and will join the described  pipeline in a near future.
-The first one will perform 2D classification and the second one will create a initial 3D map based on these classes.
+The first one will perform 2D classification and the second one will create an initial 3D map based on these classes.
 A first version of these protocols is available in the \scipion developers version.
@@
 \protocol{CTF monitor} plots the CTF defocus and astigmatism  and it will raise an alarm if these parameters are not within a given range.
-Finally the \protocol{report} protocol pushes all the created plots and summaries to a web server so this information can be accessed by users who are not physically in the microscope room.
+Finally, the \protocol{report} protocol pushes all the created plots and summaries to a web server.
+This information can be accessed by users who are not physically in the microscope room (a live example report is available at \url{http://nolan.cnb.csic.es/scipionbox/lastHTMLReport/}).
 

### 2018-04-23 3.workflow.tex
from home Marta (swedden) comments
I think this is done except for the cover letter

@@
 Different facilities have adopted different solutions and therefore, how \scipion  access data and interchange information may require minor adjustments.
-In this section we discuss  the different solutions applied in the different facilities in which \scipion is running.
+In this section we discuss  the  solutions applied by the different facilities in which \scipion is running.
 
-So far, \scipion is being used routinely in seven Cryo-EM facilities:
+\scipion is so far being used routinely in seven Cryo-EM facilities:
 \begin{itemize}
@@
  \item \emph{The Swedish National Cryo-EM Facility}.
-The Facility has two nodes: at \scilifelab in Stockholm and at Ume\aa\ University (\url{http://www.scilifelab.se}, \url{http://www.kbc.umu.se/english/ucem/cryo-em/})
+The Facility has two nodes, at \scilifelab in Stockholm and at Ume\aa\ University (\url{http://www.scilifelab.se}, \url{http://www.kbc.umu.se/english/ucem/cryo-em/})
  \item \emph{ESRF Cryo-Electron Microscope} (\url{http://www.esrf.eu/home/UsersAndScience/Experiments/MX/About_our_beamlines/CM01.html})
@@
 
-\scipion original design satisfied the requirements of the CNB, in which \scipion automatically fetches newly recorded movies from a network mounted disk.
-After this first installation, \scipion was deployed on SciLifeLab (a centralized facility located at Stockholm) where a rigorous booking system is followed.
-There, \scipion was adapted to interface with this booking platform in order to produce better reports.
-Lately,  \scipion has been installed in  large synchrotrons  such as ESRF and Diamond  where projects are handled by ISPyB (a customized laboratory information management system -LIMS- \citep{Delageniere2011:ispb}) and data are saved on a distributed file system.
+\scipion original design satisfied CNB requirements, in which \scipion automatically fetches newly recorded movies from a network mounted disk.
+After this first installation, \scipion was deployed on SciLifeLab (a centralized facility located at Stockholm), where a rigorous booking system is followed.
+There, \scipion was adapted to interface with this booking platform to produce better reports.
+Lately,  \scipion has been installed in  large synchrotrons  such as ESRF and Diamond,  where projects are handled by ISPyB (a customized laboratory information management system -LIMS- \citep{Delageniere2011:ispb}) and data are saved on a distributed file system.
 % that heavily penalizes access to lists of files.

### 2018-04-23 3.workflow_cnb.tex
I think this is done except for the cover letter

@@
 The situation changed when the new Falcon-III detector was installed.
-After that, data were produced at sustained speeds as high as 90 MB/sec (when acquiring four movies by hole in lineal mode) and the main problem became to move smoothly the data through the entire pipeline from the microscope to the final user laboratory.
+After that, data were produced at sustained speeds as high as 90 MB/sec (when acquiring four movies by hole in linear mode) and the main problem became to move smoothly the data through the entire pipeline from the microscope to the final user laboratory.
 In the following we describe such a pipeline.
@@
 
-The data flow is a follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox will temporary read the movies from 
-the microscope-NAS but will not store them locally, (c) each hour, output-NAS will connect to \scipionbox and copy -or update- the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup \scipion project and the movies to a USB disk or serve this information through the Internet and, (e) in parallel, \scipionbox transfers to  output-NAS different HTML reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox}).
+The data flow is a follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox reads the movies from the microscope-NAS but does not store them locally, (c) each hour, output-NAS connects to \scipionbox and copy -or update- the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup \scipion project and the movies to a USB disk or serve this information through the Internet and, (e) in parallel, \scipionbox transfers to  output-NAS different HTML reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox/lastHTMLReport/}).
 
@@
 USB 3.0 drives can theoretically handle data flows of 300 MB/s, but CNB experience is that, in this busy environment, many of them  have trouble exceeding 50 MB/s.
-In most cases, 50 MB/s is enough to transfer the data before the following microscope shift starts.
-Nevertheless, real-world is tough and it happens that two, or even three,  users try to download their data simultaneously.
-Fortunately, output-NAS works as front-end and isolates the data acquiring and processing infrastructure from the user downloading.
-Network users access to output-NAS using a chrooted common account.
+In most cases, 50 MB/s is enough to transfer the data before the following microscope  turn-shift starts.
+When two, or even three, users try to download their
+data simultaneously, output-NAS works as front-end and isolates the data
+acquiring and processing infrastructure from the user’s downloading activities.
+Network users access to output-NAS using a hardened common account.
 This account only allows rsync connections and make impossible to retrieve data if the name of the project to be downloaded is unknown.
-In order to speed up transfer, weak ciphers (as \emph{arcfour}) has been activated in output-NAS.
+Weak ciphers (as arcfour) has been activated in outputNAS to speed up transfer.
 
@@
 where movies acquired by the microscope could be stored, helped users to choose between a few predetermined image processing workflows and finally launched \scipion.
-At present it has grown in a client-server application that in addition to the old script capacities has incorporated the following functions:
+At present it has grown in a client-server application that, in addition to the old script capabilities, has incorporated the following functions:
 
@@
  \item Transparent addition of new workflows. \scipion is able to export any executed workflow as a json file. \emadmin can import this files and use them to create new predetermined workflows.
- \item Automatic creation of pdf user's report containing: microscope acquisition parameters and histograms summarizing data resolution and  defocus. 
- \item The above mentioned information is stored in a database for further analysis. By default the system plots the  average resolution and astigmatism of the different acquisitions versus time.
+ \item  Automatic creation of pdf user’s report that contains microscope acquisition parameters and histograms summarizing data resolution and defocus. 
+ \item The above mentioned information is stored in a database for further analysis. By default, the system plots the  average resolution and astigmatism of the different acquisitions versus time.
  \item Remote management. \emadmin is a client server application, the client maybe executed in a machine different from the one that executes \scipion.
@@
 \emadmin is free software available at github repository (\url{https://github.com/rmarabini/webservices/tree/master/EMadmin}).
-You are welcomed to try it but be aware that \emadmin is designed to be used at CNB and therefore, no support is provided for it.
+You are welcomed to download and customize it to meet your organization specific requirements.

### 2018-04-23 3.workflow_diamond.tex
I think this is done except for the cover letter

@@
 
-Diamond has thousands of users every year, both academic and commercial, and it is vital that their data is kept secure and confidential.
-A user management system has been developed which allocates each user with a unique federal identity (IDs), and associates those user IDs with the experiments in which they partake in.
+Diamond has thousands of users every year, both academic and commercial, and it is vital that their data is kept secure and confidential. A user management system
+has been developed to allocate each user with a unique federal identity (ID),
+and to associate this ID with the experiments in which the user partake in.
  Visits are structured such that all raw and processed data collected during an experiment are kept in one directory, and only the associated visitors have read and write permissions in that directory.
@@
 
-Diamond IT infrastructure is complex and requires to integrate a large collection of heterogeneous systems.
+ The complexity of Diamond IT infrastructure requires to integrate many heterogeneous systems.
 In this environment a major problem is to achieve interaction and data exchange between diverse applications.
@@
 
-\subsection{Additional Software Developments: ISyB monitor, Scipion Headless}
+\subsection{Additional Software Developments: ISPyB monitor, Scipion Headless}
 
@@
 Once set, a file describing a \scipion workflow is created and then saved to disk.
-A daemon that is running on the microscope control server notices the file creation and send a ``file creation'' message to \emph{ActiveMQ}.
+A daemon that is running on the microscope control server notices the file creation and sends a ``file creation'' message to \emph{ActiveMQ}.
 
 Another process listening to \emph{ActiveMQ} then launches \scipion from this workflow and creates a running ``headless'' project.
-\scipion headless mode has been developed specifically for eBIC and allows \scipion to be executed without user interface in large clusters with no graphical board.
+\scipion headless mode has been developed specifically for eBIC and allows \scipion to be executed without graphical user interface in large clusters with no graphical board.
 
@@
 
-To monitor the progress of the automatic image processing, CNB and \scilifelab cryo-EM services make use of \scipion default monitor system.
+Unlike CNB and SciLifeLab cryo-EM services, where the
+Scipion default monitor system displays the progress of the automatic image
+processing, in Diamond, the electronic notebook ISPyB is the natural framework
+to record and show metadata about the different experiments. A new monitor has thus been developed to allow Scipion to directly interface with ISPyB using the ISPyB python API.
+%To monitor the progress of the automatic image processing, CNB and \scilifelab cryo-EM services make use of \scipion default monitor system.
 However, in Diamond, the electronic notebook ISPyB is the standard tool to record and show metadata about the different experiments.
-In this way, a new monitor has been developed which, through a ISPyB python API, allows \scipion to directly interface with ISPyB.
-This setup provides the users with a web interface showing general metadata about their experiment as well as  real-time updates on steps of the process and relevant output such as motion corrected images, drift values, CTF images, etc.
+%In this way, a new monitor has been developed which, through a ISPyB python API, allows \scipion to directly interface with ISPyB.  
+This setup provides the users with a web interface showing general metadata about their experiment, as well as  real-time updates on steps of the process and relevant output, such as motion corrected images, drift values, CTF images, etc.
 

### 2018-04-23 3.workflow_scilifelab.tex
I think this is done except for the cover letter

@@
 %On the other side, some time of the facility is reserved for internal research groups at Stockholm University.
-Internally, a online booking system called \emph{Booked Scheduler} centralizes the reservations of the microscopes and other instruments.
+Internally, an online booking system called \emph{Booked Scheduler} centralizes the reservations of the microscopes and other instruments.
 
 At the beginning of 2016, the Cryo-EM facility at \scilifelab became one of the early adopters of \scipion streaming processing.
-Before, there were some home-made scripts to perform motion correction (with motioncor2) and CTF estimation (with Gctf) but the microscope operators recognized the importance of a more general framework.
+Although there already were using home-made scripts to perform motion correction (with motioncor2) and CTF estimation (with Gctf), the microscope operators recognized the importance of a more general
+framework.
 In particular, they were interested in the possibility of modifying easily the processing workflow as well as provide users and staff with  graphical tools for data analysis and quick feedback regarding the data collection.
-Initially, the same setup script used at CNB was adopted, but it did not fully satisfy the facility requirements and a new one has been developed.
+Although initially the same setup script used at
+CNB was adopted, a new one was developed to fully satisfy the facility
+requirements.
 In the following sections we will briefly describe the computational infrastructure at the \scilifelab node and the implementation of a Session Wizard to automate the initial setup while fetching information from other sources external to \scipion.
@@
 \subsubsection{Network Setup and IT Infrastructure}
-At the heart of cryo-EM National Facility computational infrastructure there is a storage and pre-processing server (the staging server).
+A storage and pre-processing server (the staging
+server) constitutes the core of the SciLifeLab computational infrastructure.
  This machine has roughly 200 TB of storage (4 ZFS RAIDZ2 pools with 11 HDDs each), 2 NVIDIA GeForce GTX 1070 GPUs, 2 Intel Xeon E5-2630v4 CPUs (10 cores, 2.2 GHz, HT), 384 GB RAM, and a dual-port 10 GbE network card.
+The storage pool is exported via NFS and Samba.
+The microscope computers can thus write data directly there and other machines can access it.
+
  The storage pool is exported via NFS and Samba, so the microscope computers can write data there directly and other machines can access it.
@@
 
-In order to avoid interruption of data collection due to network issues each microscope has its own dedicated storage, that in the case of the Titan microscope is double since it may be connected to two different cameras, a K2 and a Falcon III.
+To avoid interruption of data collection due to network issues each microscope has its own dedicated storage, that in the case of the Titan microscope is double since it may be connected to two different cameras, a K2 and a Falcon III.
 Hence an script to continuously move the produced data from the adequate storage to the staging server has been created.
@@
 After processing, users may  copy their data to external USB drives from two workstations in the microscope room that have access to the storage pool via NFS.
-There is also a dedicated download server which has access to the NFS export. This download server is accessible from  the Internet and allow downloads via rsync (over SSH) or SFTP.
+There is also a dedicated download server which has access to the NFS export. This download server is accessible from  the Internet and allows downloads via rsync (over SSH) or SFTP.
 
@@
 In case of the latter, additional information will be retrieved from the Application Portal (such as principal investigator, project code, contact person or invoice address).
-After gathering all this information the wizard will store it, create the processing environment and launch \scipion.
+After gathering all this information the wizard stores it, creates the processing environment and launches \scipion.
 %For example, the movies-import protocol will already point to the location where the files will be written for this session.
@@
 The described session wizard can be found at  \url{https://github.com/delarosatrevin/scipion-session}.
-It has been designed to satisfy the requirements for \scilifelab and, although readers are welcomed to try it, no support is offered.
+You are welcomed to download and customize it to meet your organization specific requirements.
 \emph{Booked Scheduler} is available at (\url{https://www.bookedscheduler.com/})

### 2018-04-23 4.future.tex
I think this is done except for the cover letter

@@
 Currently, \scipion includes the appropriate code (python wrappers) to talk to the integrated EM packages (Xmipp, Eman, Relion, ...) at the time of the release.
-Therefore, an update of any of the integrated EM packages immediately after the \scipion release won't be available until next \scipion release (we are aiming for one major release per year).
+Therefore, an update of any of the integrated EM packages immediately after the \scipion release will not be available until next \scipion release (we are aiming for one major release per year).
 To decouple EM-package releases from \scipion releases, we are working on making the wrappers for EM packages  independent from \scipion.

### 2018-04-23 5.conclusions.tex
I think this is done except for the cover letter

@@
  \scipion has been deployed in cloud and there is a public image, with all the required preinstalled cryoEM software,  available at Amazon WebServices EC2 and EGI federated clouds.
-These images make use of \scipion streaming capabilities, that is, they are able to process data as soon as it has been transfered to an Amazon storage unit, and can be used to handle  production peaks.
+This image makes use of \scipion streaming capabilities, that is, they are able to process data as soon as it has been transfered to an Amazon storage unit, and can be used to handle  production peaks.
 
@@
 at \url{http://scipion.cnb.csic.es/m/download_form/} and \url{https://github.com/I2PC/scipion/wiki/How-to-Install}, respectively.
-Cloud installation instruction are available at \url{https://github.com/I2PC/scipion/wiki/Scipion-in-the-Cloud}.
-Other public software mentioned in this article as \emadmin, \emph{wizard session} and the different scripts used at eBIC are available at \url{https://github.com/rmarabini/webservices/tree/master/EMadmin}, \url{\url{https://github.com/delarosatrevin/scipion-session}} and \url{https://github.com/I2PC/scipion/tree/release-1.1-headless-devel} respectively.
-If you would like to have more information about how to install \scipion or you have suggestions on how to improve it, contact us at scipion@cnb.csic.es.
+Cloud installation instructions are available at \url{https://github.com/I2PC/scipion/wiki/Scipion-in-the-Cloud}.
+Other public software mentioned in this article as \emadmin, \emph{wizard session} and the different scripts used at eBIC are available at \url{https://github.com/rmarabini/webservices/tree/master/EMadmin}, \url{https://github.com/delarosatrevin/scipion-session} and \url{https://github.com/I2PC/scipion/tree/release-1.1-headless-devel} respectively.
+\scipion download statistics, usage map and protocol usage ranking are available at \url{http://scipion.i2pc.es/downloadstats}, \url{http://scipion.i2pc.es/report_protocols/scipionUsage/} and \url{http://scipion.i2pc.es/report_protocols/protocolTable/} respectively.
+
+ For more information about how to install Scipion or
+suggestions to improve it, contact us at scipion@cnb.csic.es.
 

### 2018-04-25 2.overall.tex
final hopefully

@@
 The second protocol performs direct detector movie alignment.
-Currently \scipion offers  five  methods for movie alignment: motioncorr \citep{Li2015}, motioncor2 \citep{Zheng2016:motioncor2}, correlation (an unpublished CPU version that implements in Xmipp a global alignment method and solves it similarly to motioncorr but using robust statistics), unblur \citep{Grant2015:unblur} and optical flow \citep{Abrishami2015:optical-flow}.
-Then the CTF of the resulting micrographs are calculated.
+Currently, \scipion offers  five  methods for movie alignment: motioncorr \citep{Li2015}, motioncor2 \citep{Zheng2016:motioncor2}, correlation (an unpublished CPU version that implements in Xmipp a global alignment method and solves it similarly to motioncorr but using robust statistics), unblur \citep{Grant2015:unblur} and optical flow \citep{Abrishami2015:optical-flow}.
+Then the CTFs of the resulting micrographs are calculated.
 Scipion provides three CTF estimation methods:

### 2018-04-25 3.workflow_cnb.tex
final hopefully

@@
 
-At the beginning CNB networking setup was a quite straightforward one, movies where stored in the  same machine that ran \scipion and processed data were saved to external USB hard disks by this same machine.
+At the beginning CNB networking setup was a quite straightforward one, movies were stored in the  same machine that ran \scipion and processed data were saved to external USB hard disks by this same machine.
 The situation changed when the new Falcon-III detector was installed.
@@
 
-The data flow is a follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox reads the movies from the microscope-NAS but does not store them locally, (c) each hour, output-NAS connects to \scipionbox and copy -or update- the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup \scipion project and the movies to a USB disk or serve this information through the Internet and, (e) in parallel, \scipionbox transfers to  output-NAS different HTML reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox/lastHTMLReport/}).
+The data flow is as follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox reads the movies from the microscope-NAS but does not store them locally, (c) each hour, output-NAS connects to \scipionbox and copy -or update- the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup \scipion project and the movies to a USB disk or serve this information through the Internet and, (e) in parallel, \scipionbox transfers to  output-NAS different HTML reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox/lastHTMLReport/}).
 
@@
 data simultaneously, output-NAS works as front-end and isolates the data
-acquiring and processing infrastructure from the user’s downloading activities.
+acquiring and processing infrastructure from the user's downloading activities.
 Network users access to output-NAS using a hardened common account.
@@
 \begin{enumerate}
- \item Transparent addition of new workflows. \scipion is able to export any executed workflow as a json file. \emadmin can import this files and use them to create new predetermined workflows.
- \item  Automatic creation of pdf user’s report that contains microscope acquisition parameters and histograms summarizing data resolution and defocus. 
+ \item Transparent addition of new workflows. \scipion is able to export any executed workflow as a json file. \emadmin can import this file and use them to create new predetermined workflows.
+ \item  Automatic creation of pdf user's report that contains microscope acquisition parameters and histograms summarizing data resolution and defocus. 
  \item The above mentioned information is stored in a database for further analysis. By default, the system plots the  average resolution and astigmatism of the different acquisitions versus time.

### 2018-04-25 3.workflow_diamond.tex
final hopefully

@@
 has been developed to allocate each user with a unique federal identity (ID),
-and to associate this ID with the experiments in which the user partake in.
+and to associate this ID with the experiments in which the user partakes in.
  Visits are structured such that all raw and processed data collected during an experiment are kept in one directory, and only the associated visitors have read and write permissions in that directory.
-Users may access their data through ISPyB \citep{Delageniere2011:ispb},  a Laboratory Information Management System (LIMS) combining sample tracking and experiment reporting.
+Users may access their data through SynchWeb which fronts ISPyB, a Laboratory Information Management System combining sample tracking and experiment reporting \citep{Delageniere2011:ispb, Fisher2015:synchweb}.
+
+ %Users may access their data through ISPyB \citep{Delageniere2011:ispb},  a Laboratory Information Management System (LIMS) combining sample tracking and experiment reporting.
 %The link between the user and their visits enables data management such as archiving data to a long-term tape storage system, or the Laboratory Information Management System ISPyB (REF.

### 2018-04-25 3.workflow_scilifelab.tex
final hopefully

@@
 
- The storage pool is exported via NFS and Samba, so the microscope computers can write data there directly and other machines can access it.
+% The storage pool is exported via NFS and Samba, so the microscope computers can write data there directly and other machines can access it.
 %Unfortunately, the K2 [** Jose Miguel, what is K2?, do you have 2 cameras?
@@
 To avoid interruption of data collection due to network issues each microscope has its own dedicated storage, that in the case of the Titan microscope is double since it may be connected to two different cameras, a K2 and a Falcon III.
-Hence an script to continuously move the produced data from the adequate storage to the staging server has been created.
+Hence a script to continuously move the produced data from the adequate storage to the staging server has been created.
 

### 2018-04-25 scipionbox.tex
final hopefully

@@
 \title{Using Scipion for Stream Image Processing at Cryo-EM Facilities}
-Authors to be sorted later
-\author{C.O.S. Sorzano$^{(b)}$,
-A. Jim\'{e}nez$^{(b)}$,
-R. Melero$^{(b)}$,\\
-M. Mart\'{i}nez$^{(b)}$,
-J. G\'{o}mez-Blanco$^{(b)}$,
-L. del Cano$^{(b)}$,\\
-T. Majtner$^{(b)}$,
-E. Ram\'{i}rez$^{(b)}$,
-P. Conesa$^{(b)}$,\\
-J.L. Vilas$^{(b)}$,
-R. Marabini$^{(a)}$,
-Y. Rancel$^{(b)}$,
-J.M. Carazo$^{(b)}$,\\
-J.M. de la Rosa-Trev\'{i}n$^{(c)}$,
+
+\author{
+J. G\'{o}mez-Blanco$^{(a, g)}$,
+J.M. de la Rosa-Trev\'{i}n$^{(c, g)}$,
+R. Marabini$^{(b, g)}$,\\
+L. del Cano$^{(a)}$,
+A. Jim\'{e}nez$^{(a)}$,
+M. Mart\'{i}nez$^{(a)}$,\\
+R. Melero$^{(a)}$,
+T. Majtner$^{(a)}$,
+D. Maluenda$^{(a)}$,
+J. Mota$^{(a)}$,\\
+Y. Rancel$^{(a)}$,
+E. Ram\'{i}rez$^{(a)}$,
+J.L. Vilas$^{(a)}$,\\
 M. Carroni$^{(c)}$,
-S. Fleischmann $^{(c)}$,\\
-E. Lindahl$^{(c,d)}$
+S. Fleischmann $^{(c)}$,
+E. Lindahl$^{(c,d)}$\\
 A.W. Ashton$^{(e)}$,
-M. Basham$^{(e)}$,\\
-D.K. Clare$^{(e)}$,
+M. Basham$^{(e)}$,
+D.K. Clare$^{(e)}$,\\
 K. Savage$^{(e)}$,
 C.A. Siebert$^{(e)}$,\\
-D. Maluenda$^{(a)}$,\\
-Grigory G. Sharov$^{(f)}$
+Grigory G. Sharov$^{(f)}$,
+C.O.S. Sorzano$^{(a)}$,
+P. Conesa$^{(a)}$,
+J.M. Carazo$^{(a)}$,\\
 }
@@
 \lyxaddress{{\small \vspace{-0.5cm}$^{(a)}$Escuela Polit\'{e}cnica Superior, Universidad Aut\'{o}noma de Madrid, 28049 Cantoblanco, Madrid, Spain.}}
-\lyxaddress{{\small$^{(b)}$ Biocomputing Unit, National Center for Biotechnology (CSIC), C/ Darwin, 3, Campus Universidad Aut\'{o}noma, 28049 Cantoblanco, Madrid, Spain.}}
+\lyxaddress{{\small$^{(a)}$ Biocomputing Unit, National Center for Biotechnology (CSIC), C/ Darwin, 3, Campus Universidad Aut\'{o}noma, 28049 Cantoblanco, Madrid, Spain.}}
 \lyxaddress{{\small$^{(c)}$ Department of Biochemistry and Biophysics, Science for Life Laboratory, Stockholm University, Stockholm, Sweden}} 
@@
 \lyxaddress{{\small$^{(e)}$ Diamond Light Source, Harwell Science and Innovation Campus, Didcot OX11 0DE, United Kingdom}}
-\lyxaddress{{\small$^{(f)}$ MRC Laboratory of Molecular Biology - Structural studies, Cambridge, England, United Kingdom}}
+\lyxaddress{{\small$^{(f)}$ Medical Research Council Laboratory of Molecular Biology, Francis Crick Avenue, Cambridge CB2 OQH, UK}}
+\lyxaddress{{\small$^{(g)}$ These authors  contributed equally to this work.}}
 
@@
 \setcounter{figure}{2}%I do not know why but figure counter starts with -1
-\textbf{Corresponding Author: Roberto Marabini, Ph = +34 91 5854510, email: roberto@cnb.csic.es\\} %Roberto Marabini, Ph= 34 91 5854510, email: roberto@cnb.csic.es\\
+\textbf{Corresponding Author: Roberto Marabini, Phone = +34 91 5854510, email: roberto@cnb.csic.es\\} %Roberto Marabini, Ph= 34 91 5854510, email: roberto@cnb.csic.es\\
 %Escuela Polit\'{e}cnica Superior\\

### 2018-05-07 0.abstract.tex
final version

@@
 This summary may be generated in HTML format and  published through a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided.
-At present, \scipion has been installed and is in production mode in seven Cryo-EM facilities throughout the world and, it is being considered in many more.
+At present, \scipion has been installed and is in production mode in seven Cryo-EM facilities throughout the world, and it is being considered in many more.
 

### 2018-05-07 3.workflow_cnb.tex
final version

@@
 \scipionbox has 32 CPUs at 2.40GHz, 62 Gb of RAM memory and 2 \textit{Quadro M4000} GPUs.
-Finally, a second NAS with four 1Gb NICs is linked to the private, LAN and WAN networks (we will refer to this second NAS as output-NAS).
+Finally, a second NAS with four 1Gb NICs is linked to the private LAN and WAN networks (we will refer to this second NAS as output-NAS).
 
-The data flow is as follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox reads the movies from the microscope-NAS but does not store them locally, (c) each hour, output-NAS connects to \scipionbox and copy -or update- the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup \scipion project and the movies to a USB disk or serve this information through the Internet and, (e) in parallel, \scipionbox transfers to  output-NAS different HTML reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox/lastHTMLReport/}).
+The data flow is as follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox reads the movies from the microscope-NAS but does not store them locally, (c) each hour, output-NAS connects to \scipionbox and copies -or updates- the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup \scipion project and the movies to a USB disk or serve this information through the Internet and, (e) in parallel, \scipionbox transfers to  output-NAS different HTML reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox/lastHTMLReport/}).
 
@@
 \begin{enumerate}
- \item Transparent addition of new workflows. \scipion is able to export any executed workflow as a json file. \emadmin can import this file and use them to create new predetermined workflows.
+ \item Transparent addition of new workflows. \scipion is able to export any executed workflow as a json file. \emadmin can import this file and use it to create new predetermined workflows.
  \item  Automatic creation of pdf user's report that contains microscope acquisition parameters and histograms summarizing data resolution and defocus. 

### 2018-05-07 3.workflow_scilifelab.tex
final version

@@
 At the beginning of 2016, the Cryo-EM facility at \scilifelab became one of the early adopters of \scipion streaming processing.
-Although there already were using home-made scripts to perform motion correction (with motioncor2) and CTF estimation (with Gctf), the microscope operators recognized the importance of a more general
+Although the microscope operators were using home-made scripts to perform motion correction (with motioncor2) and CTF estimation (with Gctf), they recognized the importance of a more general
 framework.

### 2018-05-07 4.future.tex
final version

@@
 In this way we want to make  the data produced by \scipion more compliant with FAIR guidelines given by the \citet{eu2016:fair}.
-This workflow file could go with the raw data acquired by CryoEM facilities and deposited at common EM databases like EMPIAR or EMDB \citep{Patwardhan2016:databasesEM}.
+This workflow file could go with the raw data acquired by CryoEM facilities and deposited at common EM databases such as EMPIAR or EMDB \citep{Patwardhan2016:databasesEM}.
 To facilitate the visualization of the workflow file in any of these databases, we are  developing a  webcomponent (\scipion \emph{workflow viewer}) that will easily allow these repositories to visualize the workflow on their web pages (\url{https://github.com/I2PC/web-workflow-viewer}).

### 2018-09-14 Review/0.abstract.tex
version2

--- 2018-05-07 0.abstract.tex
@@
 This approach helps to detect problems at early stages, saves computing time and provides users with a detailed evaluation of the data quality before the acquisition is finished.
-Different EM facilities have
-implemented their stream processing solution using custom-made scripts, but this script based approach is not flexible and present a slow response time to new feature requests.
+%Different EM facilities have
+%implemented their stream processing solution using custom-made scripts, but this script based approach is not flexible and present a slow response time to new feature requests.
 The advantages of our proposed solution are: (1) flexibility: users can customize their workflows from a  large collection of preloaded algorithms; (2) repeatability and traceability: workflows are stored and can be re-executed later, either with the same data or with a different data set; (3) monitorization: there are utilities that constantly check how the execution of the algorithms is going on and inform about possible anomalies; (4) reports: we have developed several graphical interfaces that are updated periodically producing a summary (e.g, CTF defocus values, system load, etc).

### 2018-09-14 Review/1.introduction.tex
version2

--- 2018-04-23 1.introduction.tex
@@
 % in which the early detection of issues may save a data acquisition shift which otherwise needs to be repeated after a new application that may be granted many months later.
-In this way, \scipion starts the processing of the movies as they are being acquired with the double aim of checking for possible data collection errors and executing a first user-tailored image processing workflow that provides users with an accurate estimation of the acquired data quality.
+In this way, \scipion starts the processing of the movies as they are being acquired with the double aim of checking for possible data collection errors and executing a user-tailored image processing workflow that provides users with an accurate estimation of the acquired data quality.
 

### 2018-09-17 Review/answer_referee.tex
w

\documentclass[a4paper,12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{fontenc}
\usepackage{graphicx}
\usepackage{scrextend}
\usepackage{tikz}
\usepackage{pgfplotstable}%color table

\def\groel{GroEL \emph{in silico}}

\newcommand{\ffigure}[1]{Figure \ref{#1}}
\newcommand{\ttable}[1]{Table \ref{#1}}
\newcommand{\ffigures}{Figures}
\newcommand{\ttables}{Tables}

\date{06/09/18}
\input{response_macros}
\initresponses

\begin{document}

\begin{reviewer}
\section*{Reviewer 1}
\point{The graphical abstract is likely too detailed 
and would not show well when reduced to the image size that the final version has.}

\reply  Both reviewers complain that the graphical abstract need to be redesigned. The first reviewer complains that is ``...too detailed...'' while the second one claims that ``...is oversimplified...''. Since the graphical abstract is not compulsory  and both reviewers dislike it for completely opposite and irreconcilable reasons we have decided to drop it.

\point{The manuscript would benefit from a figure that outlines the layout of a typical Scipion system, showing computers, data connections, data flows in workflow scenarios, and how streaming is now implemented. At least for one of the three examples described, a figure would be helpful and make the manuscript more appealing to readers. }

\reply tut-tuc-tuc

\point{The English language at a few locations would benefit from further proof reading, but these are minor problems. }

\reply tut-tuc-tuc

% The English language at a few locations would benefit from further proof reading, but these are minor problems. 
% 
% E.g.:
% There are two different abstracts provided. One in the PDF for reviewing, and a second, shorter Abstract at the beginning of the manuscript text. 
% For the first Abstract: "… how the execution of the algorithms is going on" would be better phrased "… how the execution of the algorithms is progressing".
% For the second Abstract: "and it is being considered in many more." is a claim, not a provable fact. It should not be stated here. Even though, this statement might be true, the same would apply to all other software systems, which are likely being considered somewhere. 
% 
% Page 4: "Stream processing, i.e. computing" should be "Stream processing, i.e., computing" (comma missing). 
% 
% Page 6: "we described the pipeline offered…" should be "we describe the pipeline offered…"
% 
% Page 9: "by any user which…" should be "by any user, who…"
% Same page: Explain what EPU is.
% 
% Page 10: "Weak ciphers (as arcfour) has been activated…" Should this be "Weak ciphers have been activated…"?   Please explain, what ciphers are.
% 
% Page 11 and in a few locations thereafter: "You are welcomed to download…" is awkwardly phrased. It should be "You are welcome to download…", but even then, it should be better stated as "The software can be freely downloaded and customized to meet the specific requirements of client organizations."
% 
% Page 13:  "Thermofisher Titan Krios" should probably be "Thermo Fischer Scientific Titan Krios" or "Thermo Scientific Titan Krios". 
% 
% Page 16: The sentence "They can potentially play with the pipeline in ways that are not possible when having a programmer writing the code for them." Is not understandable on first reading. Do you mean "This allows users to adjust the pipeline, which would not be possible with hard-coded workflows."
% 
% Page 17: "her home institution" should be "his/her home institution".

\point{Minor corrections in pages: Abstract, 4, 6, 9, 10, 11, 13, 16, 17}

\reply All grammar, spelling typos and other minor corrections suggested by the reviewer have been corrected

\end{reviewer}

\begin{reviewer}
\section*{Reviewer 2}

\point{The authors claim that the strength of their software is in its ability to process data while being collected; a point that is not well-addresses by other software. At the same time, they acknowledge that UCSFImage4, Focus, and Relion are software packages similar to Scipion capable of processing data in a high-throughput manner while new data is being acquired at the microscope. A clear differentiation between Scipion and these software packages was lacking in this manuscript. What advantages Scipion provide over the abovementioned software packages?}

\reply At the software architectural level, perhaps the greatest differ-
ence is that since Scipion was developed later, special attention
was paid to defining abstraction layers to simplify maintenance
and extensibility. For example, to create a new protocol in Scipion,
only a Python script needs to be developed. From this script, the
system will discover the new protocol and will automatically build
a form and store the protocol-related information in the database.
In Appion, however, the developer must develop the script, the cor-
responding form, and an extra table in the database; Appion devel-
opers might thus need to know the underlying framework in more
detail.

It is difficult to predict how EM software will evolve in the
future. Our view is that software developers will continue to add
algorithms to the different EM packages, but that the burden of
many operations will be shifted from packages to frameworks.
Bookkeeping will require special attention to provide real tracking
and reproducibility. Workflows will have a key role when explor-
ing processing alternatives. Most algorithms will need to use dis-
tributed computing through clusters or the Cloud. Scipion is our
first step in implementing an integrative framework to address
important problems in the field simply and effectively

\point{The authors mentioned that Scipion has been installed and in production mode in seven facilities. Is there a way to know how Scipion is being used and to get statistical data on its usage?}

\reply tut-tuc-tuc

\point{Section 4 on Scipion setup at different facilities was confusing and unnecessarily lengthy. It wasn't clear to me if the "Additional software developments" made in 4.1.2 and 4.2.2 were general to Scipion or only made as a case-by-case scenario?}

\reply tut-tuc-tuc

\point{The graphical abstract is oversimplified and does not convey any of the useful information presented in the manuscript. The background is distracting, the cartoons are redundant with the labels, and the image describing the reporting feature is not legible and improperly sized (laterally stretched).}

\reply tut-tuc-tuc

\point{Finally, the manuscript could benefit greatly from more visual summaries and schematics to guide the readers especially the non-experts that the authors refer to as part of their target consumer base.}

\reply tut-tuc-tuc

\end{reviewer}
\end{document}

### 2018-09-18 Review/3.workflow_cnb.tex
tohome

--- 2018-05-07 3.workflow_cnb.tex
@@
 After that, data were produced at sustained speeds as high as 90 MB/sec (when acquiring four movies by hole in linear mode) and the main problem became to move smoothly the data through the entire pipeline from the microscope to the final user laboratory.
-In the following we describe such a pipeline.
+In the following we describe such a pipeline (see \ffigure{fig:cnbpipeline} for details).
 
-From a network point of view there are three different levels of accessibility: (1) a private network that cannot be accessed by regular users, (2) the CNB LAN (local area network)  that can be accessed by any user which works inside the hosting  institution and (3) finally, there is a single computer that may be accessed from outside the institution and therefore, it can see the WAN (wide area network).
+From a network point of view there are three different levels of accessibility (see \ffig): (1) a private network that cannot be accessed by regular users -green lines in \ffigure{fig:cnbpipeline}-, (2) the CNB LAN (local area network)  that can be accessed by any user which works inside the hosting  institution -blue lines in \ffigure{fig:cnbpipeline}- and (3) finally, there are two computers that may be accessed from outside the institution -red lines in \ffigure{fig:cnbpipeline}-, this computer give access to data and a wes site respectively.
 

### 2018-09-21 Review/3.workflow_cnb.tex
start

--- 2018-05-07 3.workflow_cnb.tex
@@
 \subsection{Case 1: Scipion at CNB}
-The National Center for Biotechnology (CNB) forms part of the Spanish National Research Council (CSIC), the largest public research institution in Spain. 
-The CNB Cryo-Electron Microscopy Facility is a joint effort of the Nacional Center for Biotechnology (CNB-CSIC) and the Biological Research Centers (CIB-CSIC).
+The National Center for Biotechnology (\cnb) forms part of the Spanish National Research Council (CSIC), the largest public research institution in Spain. 
+The \cnb Cryo-Electron Microscopy Facility is a joint effort of the Nacional Center for Biotechnology and the Biological Research Centers.
 The mission of the facility is to provide researchers access to the latest cryo-EM technology for high resolution imaging.
@@
 
-At the beginning CNB networking setup was a quite straightforward one, movies were stored in the  same machine that ran \scipion and processed data were saved to external USB hard disks by this same machine.
+At the beginning \cnb networking setup was a quite straightforward one, movies were stored in the same machine that ran \scipion and processed data were saved to external USB hard disks by this same machine.
 The situation changed when the new Falcon-III detector was installed.
@@
 
-From a network point of view there are three different levels of accessibility: (1) a private network that cannot be accessed by regular users, (2) the CNB LAN (local area network)  that can be accessed by any user which works inside the hosting  institution and (3) finally, there is a single computer that may be accessed from outside the institution and therefore, it can see the WAN (wide area network).
+From a network point of view there are three different levels of accessibility (see  \ffigure{fig:cnbpipeline}): (1) a private network that cannot be accessed by regular users -green lines in \ffigure{fig:cnbpipeline}-, (2) the \cnb LAN (local area network)  that can be accessed by any user which works inside the hosting  institution -blue lines in \ffigure{fig:cnbpipeline}- and (3) finally, there are two computers that may be accessed from outside the institution -red lines in \ffigure{fig:cnbpipeline}-, these computers give access to data and a web site respectively.
+s
+
+\begin{figure}
+  \centering
+      \includegraphics[width=0.75\textwidth]{images/diagram.pdf}
+  \caption{Graphical representation of the data flow through the CNB facility.
+Green, blue and red lines represent the microscope private network, the \cnb local area network and the public network respectively}
+  \label{fig:cnbpipeline}
+
+\end{figure}
 
@@
 
-The data flow is as follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox reads the movies from the microscope-NAS but does not store them locally, (c) each hour, output-NAS connects to \scipionbox and copies -or updates- the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup \scipion project and the movies to a USB disk or serve this information through the Internet and, (e) in parallel, \scipionbox transfers to  output-NAS different HTML reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox/lastHTMLReport/}).
+The data flow is as follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox reads the movies from the microscope-NAS but does not store them locally, (c) output-NAS has access to Microscope-NAS and \scipionbox hard disks (d) output-NAS has a collection of scripts than may be used to incrementally backup \scipion project and the movies to a local or remote disk, or serve this information through the Internet and, (e) in parallel, periodically \scipionbox transfers to  HTML Server reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox/lastHTMLReport/}).
 

### 2018-09-24 Review/answer_referee.tex
to home

@@
 \usepackage{pgfplotstable}%color table
+\usepackage{xspace}
+\usepackage{url}
 
-\def\groel{GroEL \emph{in silico}}
+\def\cnb{CNB\xspace}
 
@@
 \newcommand{\ttables}{Tables}
+\def\scipionbox{\textit{ScipionBox}\xspace}
+\def\mnas{\textit{microscope-NAS}\xspace}
+\def\onas{\textit{Output-NAS}\xspace}
+\def\hserver{\textit{HTML-Server}\xspace}
+\def\scipion{\textit{Scipion}\xspace}
+\def\emadmin{\textit{EMAdmin}\xspace}
+
+\def\epu{\textit{EPU}\xspace}
+\def\scilifelab{SciLifeLab\xspace}
+\def\cnbcsic{CNB-CSIC\xspace}
+\def\cnb{CNB\xspace}
 
@@
 
-\reply tut-tuc-tuc
+\reply The following figure has been included in the manuscript.
 
-\point{The English language at a few locations would benefit from further proof reading, but these are minor problems. }
+\begin{figure}
+  \centering
+      \includegraphics[width=0.75\textwidth]{images/diagram.pdf}
+  \caption{Graphical representation of the data flow through the CNB facility.
+Green, blue and red lines represent the microscope private network, the \cnb local area network and the public network respectively.}
+  \label{fig:cnbpipeline}
 
-\reply tut-tuc-tuc
+\end{figure}
+\clearpage
+
+%\point{The English language at a few locations would benefit from further proof reading, but these are minor problems. }
+
+%\reply tut-tuc-tuc
 
@@
 
-\reply All grammar, spelling typos and other minor corrections suggested by the reviewer have been corrected
+\reply All grammar, spelling typos, rephrasing and other minor corrections suggested by the reviewer have been followed except one.
+Below we comment on the non followed suggestions (page 9 and 10) and the modifications introduced in page 16 since involve the rewriting of a paragraph.
+
+Q \the\numexpr\value{pointcounter}-1\relax.1: Page 9: ``...Explain what EPU is...''. EPU has already been introduced by the following two paragraphs.
+
+\begin{quote}
+ Many software packages have been developed to support automated data collection, such as 
+ .., EPU (Thermo Fisher Scientific, 2018), ..., etc.
+\end{quote}
+
+\begin{quote}
+...we comment on an additional software (called  \emadmin) that connects \scipion with the microscope acquisition software (\epu) and records the facility activity.
+\end{quote}
+
+Q \the\numexpr\value{pointcounter}-1\relax.1: Page 10: ``Please explain, what ciphers are''.
+Following reviewer 2 recommendations this section has been simplified and the sentence containing the word ``ciphers'' has been deleted.
+
+Q \the\numexpr\value{pointcounter}-1\relax.2: Page 16: ``The sentence -They can potentially play with the pipeline in ways that are not possible when having a programmer writing the code for them.- Is not understandable on first reading.''.
+We have rewriting the paragraph as:
+
+\begin{quote}
+ Workflows implement pipelines and describe them as a set of interconnected steps.
+In a workflow approach, users (including microscope staff) can reasonably be expected to modify the paths and incorporate new steps using a GUI.
+Alternate solutions as scripts do not clearly separate each step definition from the connections between the different steps and, for no programmers, are scary and difficult to modify.
+\end{quote}
 
@@
 A clear differentiation between Scipion and these software packages was lacking in this manuscript.
-What advantages Scipion provide over the abovementioned software packages?}
+What advantages Scipion provide over the above mentioned software packages?}
 
-\reply At the software architectural level, perhaps the greatest differ-
-ence is that since Scipion was developed later, special attention
-was paid to defining abstraction layers to simplify maintenance
-and extensibility. For example, to create a new protocol in Scipion,
+[[[[[[[[[[[[[[[[[[[THIS ANSWER NEED TO BE POLISHED]]]]]]]]]]]]]]]]]]]]]]
+
+\reply Since there is already a published manuscript describing \scipion, in the present manuscript  we decided to focus on \scipion features that have been specifically developed for working at EM facilities (e.g.
+streaming).
+The main reasons why \scipion is better suited than other packages for EM work is rooted in \scipion main design rather than the new facilities related features.
+Nevertheless, the reviewer is right in pointing out than a ``A clear differentiation between Scipion and these software packages was lacking in this manuscript''.
+Therefore, we have decided to add the following paragraphs to the manuscript
+
+\begin{quote}
+
+At the software architectural level,  the greatest difference 
+between \scipion and the above mentioned software is that  \scipion pays 
+special attention to define abstraction layers to simplify maintenance
+and extensibility, and that follows a strict workflow approach. A few examples will illustrate the main advantages.
+
+To create a new protocol in Scipion,
 only a Python script needs to be developed. From this script, the
 system will discover the new protocol and will automatically build
-a form and store the protocol-related information in the database.
-In Appion, however, the developer must develop the script, the cor-
-responding form, and an extra table in the database; Appion devel-
-opers might thus need to know the underlying framework in more
-detail.
+a form and store the protocol related information in a database.
+\scipion protocol developers must certainly know \scipion conventions
+but they do not need to modify or even understand the existing code.
+In  \scipion next version, the abstraction will be increased
+since protocols will be implemented as plug-ins,
+that is, independent small python scripts  that add functionality to the main
+\scipion application. The plug-in approach allows parallel development. That is,  
+since protocols are implemented independently from the main application
+they can be developed and released in parallel by different teams. In this way, support for the new version of \textit{Relion} does not require a new \scipion release but
+a release of the independent \textit{Relion} plug-in
 
-It is difficult to predict how EM software will evolve in the
-future. Our view is that software developers will continue to add
-algorithms to the different EM packages, but that the burden of
-many operations will be shifted from packages to frameworks.
-Bookkeeping will require special attention to provide real tracking
-and reproducibility. Workflows will have a key role when explor-
-ing processing alternatives. Most algorithms will need to use dis-
-tributed computing through clusters or the Cloud. Scipion is our
-first step in implementing an integrative framework to address
-important problems in the field simply and effectively
+\scipion follows a pure workflows approach and therefore implement pipelines and describe them as a set of interconnected steps (or protocols in \scipion).
+In a workflow approach, users (including microscope staff) can reasonably be expected to modify the paths and incorporate new steps using a GUI.
+The previously mentioned software packages do not clearly separate each step definition from the connections between the different steps and, for no programmers, create or modify workflows may be challenging.
+
+Another advantage of a clear separation between steps and relationships 
+is the improvement of load valance, reproducibility and  provenance (process of tracing and recording the origins of data and its movement between protocols).
+The first step in a \scipion workflow execution
+is to store in a database each of the protocols to be executed as well as how 
+outputs of one protocol are attached to the input of another protocol.
+Then, a workflow engine analyzes the stored data and using as constrain the number of CPUs assigned by the user to the workflow optimizes the execution, that is, try to execute as many protocols in parallel as possible.
+Obviously, the data stored in the database  have enough information to execute each step and therefore guaranties reproducibility and provenance.
+\end{quote}
+
+%In Appion, however, the developer must develop the script, the cor-
+%responding form, and an extra table in the database; 
+%Appion developers might thus need to know the underlying framework in more
+%detail.
+
+%It is difficult to predict how EM software will evolve in the
+%future. Our view is that software developers will continue to add
+%algorithms to the different EM packages, but that the burden of
+%many operations will be shifted from packages to frameworks.
+%Bookkeeping will require special attention to provide real tracking
+%and reproducibility. Workflows will have a key role when explor-
+%ing processing alternatives. Most algorithms will need to use dis-
+%tributed computing through clusters or the Cloud. Scipion is our
+%first step in implementing an integrative framework to address
+%important problems in the field simply and effectively
 
@@
 
-\reply tut-tuc-tuc
+\reply The following paragraphs have been added in the manuscript.
+
+\begin{quote}
+
+In the context of free and open source projects it is difficult to create usage statistics that are accurate. But certainly, if we want \scipion to succeed, we need to 
+know how frequent the different protocols are used.
+This is the reason why we have recently developed our \textit{Scipion Usage Data Collector} that monitors -if activated by the user- the usage of the different protocols and send -via HTTP protocol- the information to the developer's team.
+This information is anonymous and cannot be used to identify the original computer in which \scipion has been executed.
+
+Using the HTTP header it is possible to guess the country in which \scipion has been installed.
+With these information we have created the usage map and table available at URL \url{http://scipion.i2pc.es/report_protocols/scipionUsage/} and  \url{http://scipion.i2pc.es/report_protocols/protocolTable/}.
+The data corresponds on ``global'' \scipion usage rather than ``facility'' usage but there
+is a one to one relationship between the  countries with highest number of \scipion projects and the countries with \scipion based EM facilities.
+The only exception to the rule is Canada where two of the original \scipion developers are setting up a new facility (McGill University).
+
+Note that, in a effort to report real use and skip test projects, the number of projects in the map refers to the number of non empty \scipion projects updated more than once. That is, 
+\scipion projects that have been opened in at least two different days. Another constraint is that developers computers are filtered out.
+\end{quote}
 
@@
 
-\reply tut-tuc-tuc
+\reply Section 4 has been extensively rewritten to take into account the reviewers' suggestions.
+In summary, we have included a new image, delete several paragraphs, modify others and move those paragraphs related with future work to a different section.
+
+Please confer the file ``Revised Manuscript(marked)'' attached to this review in which 
+all significant differences between the original manuscript and the revised one are marked up.
+
+Q \the\numexpr\value{pointcounter}-1\relax.1: ``It wasn't clear to me if the Additional software developments made in 4.1.2 and 4.2.2 were general to Scipion or only made as a case-by-case scenario?''.
+The following paragraph have been modified.
+
+\begin{quote}
+ This section in divided in two parts.
+In the first one we discuss the network setup that allows us to move the data from the microscope to the user laboratory.
+Then,  we comment on an additional software called  \emadmin.
+\emadmin has been specifically designed for the \cnb facility, it connects \scipion with the microscope acquisition software (\epu) and records the facility activity.
+\end{quote}
 
@@
 
-\reply tut-tuc-tuc
+\reply See answer to reviewer 1 question 1.
 
@@
 
-\reply tut-tuc-tuc
+\reply We think the reviewer refer here to section 4. As commented previously this section has been extensively rewritten to take into account the reviewers' suggestions and it includes a new figure.
 

### 2018-09-24 Review/0.abstract.tex
to home

@@
 To manage such large-scale projects, we had previously developed  a modular workflow system called \scipion \citep{delarosatrevin2016}.
-We present here a major extension of \scipion that allows stream processing of EM images while the data is being acquired.
+We present here a major extension of \scipion that allows processing of EM images while the data is being acquired.
 This approach helps to detect problems at early stages, saves computing time and provides users with a detailed evaluation of the data quality before the acquisition is finished.
@@
 This summary may be generated in HTML format and  published through a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided.
-At present, \scipion has been installed and is in production mode in seven Cryo-EM facilities throughout the world, and it is being considered in many more.
+At present, \scipion has been deployed and is in production mode in seven Cryo-EM facilities throughout the world.%, and it is being considered in many more.
 

### 2018-09-24 Review/2.overall.tex
to home

--- 2018-04-25 2.overall.tex
@@
 
-Stream processing, i.e. computing on data directly as they are produced or received, is the main change introduced in \scipion. 
+Stream processing, i.e., computing on data directly as they are produced or received, is the main change introduced in \scipion. 
 In this way, a new level of parallelization has been reached since the  data generated by the microscope continuously flows through a chain of protocols.
@@
 \scipion is a flexible framework that allows to create different workflows by choosing among many algorithms at each step.
-Although the best workflow for each case will depend on specific requirements of specimens, microscopes or facilities, we  described the pipeline offered by default at the \emph{National Center for Biotechnology} (CNB) cryoEM Facility as use case to clarify Scipion capabilities.
+Although the best workflow for each case will depend on specific requirements of specimens, microscopes or facilities, we  describe the pipeline offered by default at the \emph{National Center for Biotechnology} (\cnb) cryoEM Facility as use case to clarify \scipion capabilities.
 

### 2018-09-24 Review/3.workflow.tex
version2
to home

--- 2018-04-23 3.workflow.tex
@@
 Different facilities have adopted different solutions and therefore, how \scipion  access data and interchange information may require minor adjustments.
-In this section we discuss  the  solutions applied by the different facilities in which \scipion is running.
+In this section we discuss  the  solutions applied by some facilities in which \scipion is running.
 
-\scipion is so far being used routinely in seven Cryo-EM facilities:
+So far \scipion is being used routinely in seven Cryo-EM facilities:
 \begin{itemize}
@@
  \item \emph{National Cancer Institute - NIH} (\url{https://www.cancer.gov/research/resources/cryoem})
+ \item \emph{University of Virginia Health System University Hospital} (\url{https://med.virginia.edu/molecular-electron-microscopy-core/})
 
 \end{itemize}
-and it is being considered in many more.
+%and it is being considered in many more.
 
@@
 After this first installation, \scipion was deployed on SciLifeLab (a centralized facility located at Stockholm), where a rigorous booking system is followed.
-There, \scipion was adapted to interface with this booking platform to produce better reports.
+There, middleware (software that acts as a bridge between applications) was developed to interface with this booking platform to produce better reports.
 Lately,  \scipion has been installed in  large synchrotrons  such as ESRF and Diamond,  where projects are handled by ISPyB (a customized laboratory information management system -LIMS- \citep{Delageniere2011:ispb}) and data are saved on a distributed file system.
@@
 
-In the following, through three use cases, we describe in detail how \scipion has been adapted to the different environments.
+%In the following, through three use cases, we describe in detail how \scipion has been adapted to the different environments.
+
+In the following, through the three above mentioned use cases, we describe in detail how \scipion is been used in these different environments.
+%Notice that EM facilities are designed and implemented to support high-throughput data acquisition and require complex data-collection workflows.
+These workflows need access to experimental parameters such as microscope settings, sample and user information, as well as channels for feeding processing results back to the control system.
+For each use case we will describe first the facility setup and then in the subsection entitled ``Aditional Software Developments'' we will comment on specific  software developed for each facility that connects \scipion with the facility and microscope control software.
 

### 2018-09-24 Review/3.workflow_cnb.tex
to home

@@
 The \cnb Cryo-Electron Microscopy Facility is a joint effort of the Nacional Center for Biotechnology and the Biological Research Centers.
-The mission of the facility is to provide researchers access to the latest cryo-EM technology for high resolution imaging.
+%The mission of the facility is to provide researchers access to the latest cryo-EM technology for high resolution imaging.
 
@@
 In the first one we discuss the network setup that allows us to move the data from the microscope to the user laboratory.
-Then  we comment on an additional software (called  \emadmin) that connects \scipion with the microscope acquisition software (\epu) and records the facility activity.
+Then,  we comment on an additional software called  \emadmin.
+\emadmin has been specifically designed for the \cnb facility, it connects \scipion with the microscope acquisition software (\epu) and records the facility activity.
 
@@
 After that, data were produced at sustained speeds as high as 90 MB/sec (when acquiring four movies by hole in linear mode) and the main problem became to move smoothly the data through the entire pipeline from the microscope to the final user laboratory.
-In the following we describe such a pipeline.
-
-From a network point of view there are three different levels of accessibility (see  \ffigure{fig:cnbpipeline}): (1) a private network that cannot be accessed by regular users -green lines in \ffigure{fig:cnbpipeline}-, (2) the \cnb LAN (local area network)  that can be accessed by any user which works inside the hosting  institution -blue lines in \ffigure{fig:cnbpipeline}- and (3) finally, there are two computers that may be accessed from outside the institution -red lines in \ffigure{fig:cnbpipeline}-, these computers give access to data and a web site respectively.
-s
+In the following we describe such a pipeline (see \ffigure{fig:cnbpipeline} for details.).
 
@@
   \caption{Graphical representation of the data flow through the CNB facility.
-Green, blue and red lines represent the microscope private network, the \cnb local area network and the public network respectively}
+Green, blue and red lines represent the microscope private network, the \cnb local area network and the public network respectively.}
   \label{fig:cnbpipeline}
@@
 
+%From a network point of view there are three different levels of accessibility (see  \ffigure{fig:cnbpipeline}): (1) a private network that cannot be accessed by regular users -green lines in \ffigure{fig:cnbpipeline}-, (2) the \cnb LAN (local area network)  that can be accessed by any user who works inside the hosting  institution -blue lines in \ffigure{fig:cnbpipeline}- and (3) finally, there are two computers that may be accessed from outside the institution -red lines in \ffigure{fig:cnbpipeline}-, these computers give access to data and a web site respectively.
+
 Data sets are recorded using \epu.
-The microscope control system has a 70 TB NAS (Network-attached storage) unit for saving movies as they are obtained (we will refer to this NAS as microscope-NAS).
+The microscope control system has a 70 TB NAS (Network-attached storage) unit for saving movies as they are obtained (we will refer to this NAS as \mnas).
 This storage unit is shared through a Samba share and connected to the private network using a 10Gb connexion.
-\scipion is executed on a Linux server which has two 1Gb NICs (network interface controllers) that connect it to the private and the LAN networks respectively (we will refer to this machine as \scipionbox).
+\scipion is executed on a Linux server which has two 1Gb NICs (network interface controllers) that connect it to the private network and the LAN (local area network) respectively (we will refer to this machine as \scipionbox).
 \scipionbox has 32 CPUs at 2.40GHz, 62 Gb of RAM memory and 2 \textit{Quadro M4000} GPUs.
-Finally, a second NAS with four 1Gb NICs is linked to the private LAN and WAN networks (we will refer to this second NAS as output-NAS).
+Finally, a second NAS with four 1Gb NICs is linked to the private, LAN and WAN networks (we will refer to this second NAS as \onas) so it can retrieve data from any computer and serve it to external users.
 
-The data flow is as follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox reads the movies from the microscope-NAS but does not store them locally, (c) output-NAS has access to Microscope-NAS and \scipionbox hard disks (d) output-NAS has a collection of scripts than may be used to incrementally backup \scipion project and the movies to a local or remote disk, or serve this information through the Internet and, (e) in parallel, periodically \scipionbox transfers to  HTML Server reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox/lastHTMLReport/}).
+The data flow is as follows: (a) the microscope system stores movies in the \mnas, (b) when executing movie-alignment protocols \scipionbox reads the movies from the \mnas but does not store them locally, (c) \onas has access to \mnas and \scipionbox hard disks (d) \onas has a collection of scripts than may be used to incrementally backup \scipion project and the movies to a local or remote disk, or serve this information through the Internet and, (e) in parallel, periodically \scipionbox transfers to  \hserver reports that may be checked by users or staff with an HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox/lastHTMLReport/}).
 
-Although CNB staff try to discourage the use of external USB hard disks versus network as way of retrieving the raw and processed data, this is the preferred option for most users.
-USB 3.0 drives can theoretically handle data flows of 300 MB/s, but CNB experience is that, in this busy environment, many of them  have trouble exceeding 50 MB/s.
+Although \cnb staff try to discourage the use of external USB hard disks versus network as way of retrieving the raw and processed data, this is the preferred option for most users.
+USB 3.0 drives can theoretically handle data flows of 300 MB/s, but \cnb experience is that, in this busy environment, many of them  have trouble exceeding 50 MB/s.
 In most cases, 50 MB/s is enough to transfer the data before the following microscope  turn-shift starts.
-When two, or even three, users try to download their
-data simultaneously, output-NAS works as front-end and isolates the data
-acquiring and processing infrastructure from the user's downloading activities.
-Network users access to output-NAS using a hardened common account.
+%When two, or even three, users try to download their
+%data simultaneously, \onas works as front-end and isolates the data
+%acquiring and processing infrastructure from the user's downloading activities.
+Network users access to \onas using a hardened common account.
 This account only allows rsync connections and make impossible to retrieve data if the name of the project to be downloaded is unknown.
-Weak ciphers (as arcfour) has been activated in outputNAS to speed up transfer.
+%The achievable speed of copying a file depends on quite a few different factors
+%including the selected ciphers (algorithms for performing data encryption or decryption). Weak ciphers (as arcfour) have been activated in \onas to speed up transfer.
 
@@
 \emadmin is free software available at github repository (\url{https://github.com/rmarabini/webservices/tree/master/EMadmin}).
-You are welcomed to download and customize it to meet your organization specific requirements.
+The software can be freely downloaded and customized to meet the specific requirements of client organizations.

### 2018-09-24 Review/3.workflow_diamond.tex
to home

--- 2018-04-25 3.workflow_diamond.tex
@@
 Recently, the electron Bio-Imaging Centre (eBIC) has been set up at Diamond~\citep{diamond2017}.
-The centre houses four Thermofisher Titan Krios, a Talos Arctica and a Scios dual-beam.
+The centre houses four Thermo Fischer Scientific Titan Krios, a Talos Arctica and a Scios dual-beam.
 All of the transmission electron microscopes are equipped with counting-mode direct electron detectors and Volta phase plates.
@@
 and to associate this ID with the experiments in which the user partakes in.
- Visits are structured such that all raw and processed data collected during an experiment are kept in one directory, and only the associated visitors have read and write permissions in that directory.
+ %Visits are structured such that all raw and processed data collected during an experiment are kept in one directory, and only the associated visitors have read and write permissions in that directory.
  Users may access their data through SynchWeb which fronts ISPyB, a Laboratory Information Management System combining sample tracking and experiment reporting \citep{Delageniere2011:ispb, Fisher2015:synchweb}.
@@
 
-Although most of developments described in this subsection will be integrated in the main \scipion code repository, at present they are available in a developer's branch  at \url{https://github.com/I2PC/scipion/tree/release-1.1-headless-devel}.
+At present, most of developments described in this subsection are available in a developer's branch  at \url{https://github.com/I2PC/scipion/tree/release-1.1-headless-devel}.
+Some of them (e.g.
+the monitor protocol) will be integrated in the main \scipion repository while the more Diamond specific software will be published in an independent repository.
+[** Ask Pablo about headless]

### 2018-09-24 Review/3.workflow_scilifelab.tex
to home

--- 2018-05-07 3.workflow_scilifelab.tex
@@
 
-The Swedish National Cryo-EM Facility offers access to state-of-the-art equipment and expertise in single particle cryo-EM and cryo electron tomography (cryo-ET).
-The Facility has two nodes: at \scilifelab in Stockholm and at Ume\aa\ University.
-\scilifelab in Stockholm offers single-particle service with a
+The Swedish National Cryo-EM Facility %%offers access to state-of-the-art equipment and expertise in single particle cryo-EM and cryo electron tomography (cryo-ET). The Facility 
+has two nodes: at \scilifelab in Stockholm and at Ume\aa\ University. \scilifelab in Stockholm offers single-particle service with a 
 Talos Arctica for sample optimisation and a Titan Krios for high-resolution data collection.
 The Ume\aa\ node is expected to become operational during 2018, and will offer cryo-ET with a Titan Krios and a Scios DualBeam SEM.
-
-The facilities can be accessed by Swedish researchers through 
-a peer-reviewed process.
+%The facilities can be accessed by Swedish researchers through 
+%a peer-reviewed process.
 Applications can be submitted through an application portal (\url{https://cryoem.scilifelab.se/}).
@@
 
-To avoid interruption of data collection due to network issues each microscope has its own dedicated storage, that in the case of the Titan microscope is double since it may be connected to two different cameras, a K2 and a Falcon III.
+%To avoid interruption of data collection due to network issues each microscope has its own dedicated storage, that in the case of the Titan microscope is double since it may be connected to two different cameras, a K2 and a Falcon III.
 Hence a script to continuously move the produced data from the adequate storage to the staging server has been created.
@@
 Unfortunately, it was difficult to reuse the existing code due to a different folder structure and, more important, a different approach to user and microscope management.
-%To address the new set of requirements, a new wizard is under development at \scilifelab.
-For this new wizard, a basic data model was designed representing the current processing and data organization at \scilifelab.
+To address the new set of requirements, a new wizard has been development at \scilifelab.
+(In this contest wizard refers to a helper script which guide users through a certain process.) %For this new wizard, a basic data model was designed representing the current processing and data organization at \scilifelab.
 %The model is composed by the following entities:
@@
 When a user starts a data collection, the wizard is executed using the command \textit{session-titan} (or \textit{session-talos}).
-The wizard determines which user has booked the microscope by fetching today's booking information from \emph{Booked Scheduler} through their web API.
-From the reservation information the wizard will also detect if it is an internal or a national facility project.
-In case of the latter, additional information will be retrieved from the Application Portal (such as principal investigator, project code, contact person or invoice address).
-After gathering all this information the wizard stores it, creates the processing environment and launches \scipion.
+The wizard determines which user has booked the microscope by fetching today's booking information from \emph{Booked Scheduler} through their web API, %From the reservation information the wizard will also detect if it is an internal or a national facility project.
+%In case of the latter, additional information will be retrieved from the Application Portal (such as principal investigator, project code, contact person or invoice address). 
+%After gathering all this information the wizard 
+stores it, creates the processing environment and launches \scipion.
 %For example, the movies-import protocol will already point to the location where the files will be written for this session.
 Moreover, taking into account which microscope-camera is being  used, the files pattern can be inferred.
-
 The current implementation of the wizard stores each session in a simple database file.
-This database, together with extra information from the Booking System and the Application Portal, is used to generate invoices and reports for a given period.
+This database, together with extra information from the Booking System and an Application Portal, is used to generate invoices and reports for a given period.
 %The generation of a session report is under development, that will provide a summary and useful information for users to access the acquired data.
@@
 
-The described session wizard can be found at  \url{https://github.com/delarosatrevin/scipion-session}.
-You are welcomed to download and customize it to meet your organization specific requirements.
+The described session wizard can be found at  
+\url{https://github.com/delarosatrevin/scipion-session}.
+The software can be freely downloaded and customized to meet the specific requirements of client organizations.
 \emph{Booked Scheduler} is available at (\url{https://www.bookedscheduler.com/})

### 2018-09-24 Review/4.future.tex
to home

--- 2018-05-07 4.future.tex
@@
+\section{Future Developments and Usage}
+
 \subsection{Future Developments}
-
 Currently, \scipion includes the appropriate code (python wrappers) to talk to the integrated EM packages (Xmipp, Eman, Relion, ...) at the time of the release.
@@
 This repository will contain ready-to-use workflows that support a range of use cases.
+
+\subsection{Usage}
+In the context of free and open source projects it is difficult to create usage statistics that are accurate. But certainly, if we want \scipion to succeed, we need to 
+know how frequent the different protocols are used.
+This is the reason why we have recently developed our \textit{Scipion Usage Data Collector} that monitors -if activated by the user- the usage of the different protocols and send -via HTTP protocol- the information to the developer's team.
+This information is anonymous and cannot be used to identify the original computer in which \scipion has been executed.
+
+Using the HTTP header it is possible to guess the country in which \scipion has been installed.
+With these information we have created the usage map and table available at URL \url{http://scipion.i2pc.es/report_protocols/scipionUsage/} and  \url{http://scipion.i2pc.es/report_protocols/protocolTable/}.
+The data corresponds on ``global'' \scipion usage rather than ``facility'' usage but there
+is a one to one relationship between the  countries with highest number of \scipion projects and the countries with \scipion based EM facilities.
+The only exception to the rule is Canada where two of the original \scipion developers are setting up a new facility (McGill University).
+
+Note that, in a effort to report real use and skip test projects, the number of projects in the map refers to the number of non empty \scipion projects updated more than once. That is, 
+\scipion projects that have been opened in at least two different days. Another constraint is that developers computers are filtered out.

### 2018-09-24 Review/5.conclusions.tex
to home

--- 2018-04-23 5.conclusions.tex
@@
 
- Workflow systems have been explored and used extensively in bioinformatics but there is relatively little progress in that direction in the cryo-EM community.
+ Workflow systems have been explored and used extensively in bioinformatics but they have raised little interest in the cryo-EM community.
 \scipion has been designed following a strict workflow paradigm.
@@
 
- In a workflow approach, users (including microscope staff) can reasonably be expected to edit the paths.
-They can potentially play with the pipeline in ways that are not possible when having a programmer writing the code for them.
-Furthermore, not having the processing path hardcoded in the program means you do not have to recompile, so changes can be quickly tested   even at run time.
+ Workflows implement pipelines and describe them as a set of interconnected steps.
+In a workflow approach, users (including microscope staff) can reasonably be expected to modify the paths and incorporate new steps using a GUI.
+Alternate solutions as scripts do not clearly separate each step definition from the connections between the different steps and, for no programmers, are scary and difficult to modify.
 The \scipion workflow approach provides provenance, that is,  for any created object (micrograph, CTFs, class, etc) \scipion stores metadata such as which programs, which parameters, and which exact path was followed to create any output data.
@@
  \scipion stream capabilities can handle large and continuously produced volumes of data.
-The benefits of streaming are clear: (1) it helps to process data as soon as it is generated,  providing almost instant results on data quality; (2) it saves user processing time and resources at her home institutions since part of the processing is done while data is being acquired; and (3)
+The benefits of streaming are clear: (1) it helps to process data as soon as it is generated,  providing almost instant results on data quality; (2) it saves user processing time and resources at the user home institution since part of the processing is done while data is being acquired; and (3)
  it further reduces the overall experimental time by allowing processing the data while it is transferred to the final user laboratory.

### 2018-09-25 Review/answer_referee.tex
to home

@@
 
-[[[[[[[[[[[[[[[[[[[THIS ANSWER NEED TO BE POLISHED]]]]]]]]]]]]]]]]]]]]]]
-
 \reply Since there is already a published manuscript describing \scipion, in the present manuscript  we decided to focus on \scipion features that have been specifically developed for working at EM facilities (e.g.
@@
 Nevertheless, the reviewer is right in pointing out than a ``A clear differentiation between Scipion and these software packages was lacking in this manuscript''.
-Therefore, we have decided to add the following paragraphs to the manuscript
+Therefore, we have decided to add the following subsection to the manuscript
 
@@
 
+\textbf{\scipion Design and Architecture}
+
+We would like to finish this section commenting on some of \scipion architecture design decisions that differentiate \scipion from another image processing frameworks.
 At the software architectural level,  the greatest difference
-between \scipion and the above mentioned software is that  \scipion pays 
-special attention to define abstraction layers to simplify maintenance
-and extensibility, and that follows a strict workflow approach. A few examples will illustrate the main advantages.
+between \scipion and the other EM image processing packages mentioned in the introduction 
+is that  \scipion pays special attention to define abstraction layers to simplify maintenance
+and extensibility, and that follows a strict workflow approach. A few examples will illustrate the main advantages of this design.
 
-To create a new protocol in Scipion,
-only a Python script needs to be developed. From this script, the
-system will discover the new protocol and will automatically build
+To create a new protocol in Scipion, only a Python script needs to be developed. From this script, the system will discover the new protocol and will automatically build
 a form and store the protocol related information in a database.
@@
 since protocols are implemented independently from the main application
-they can be developed and released in parallel by different teams. In this way, support for the new version of \textit{Relion} does not require a new \scipion release but
+they can be developed and released in parallel by different teams. In this way, support for a new version of \textit{Relion} does not require a new \scipion release but
 a release of the independent \textit{Relion} plug-in
 
-\scipion follows a pure workflows approach and therefore implement pipelines and describe them as a set of interconnected steps (or protocols in \scipion).
+\scipion follows a pure workflow approach and therefore, implements pipelines and describes them as a set of interconnected steps (or protocols).
 In a workflow approach, users (including microscope staff) can reasonably be expected to modify the paths and incorporate new steps using a GUI.
-The previously mentioned software packages do not clearly separate each step definition from the connections between the different steps and, for no programmers, create or modify workflows may be challenging.
+On the contrary, software packages that do not make this separation produce workflows that, in general, are more challenging to create or modify for non programmers.
 
@@
 is to store in a database each of the protocols to be executed as well as how 
-outputs of one protocol are attached to the input of another protocol.
-Then, a workflow engine analyzes the stored data and using as constrain the number of CPUs assigned by the user to the workflow optimizes the execution, that is, try to execute as many protocols in parallel as possible.
+outputs of one protocol are attached to the inputs of another protocol.
+Then, a workflow engine analyzes the stored data and using as constrain the number of CPUs assigned by the user to the workflow, optimizes the execution, that is, try to execute as many protocols in parallel as possible.
 Obviously, the data stored in the database  have enough information to execute each step and therefore guaranties reproducibility and provenance.
+
 \end{quote}

### 2018-09-25 Review/2.overall.tex
to home

@@
 In the following we describe the most important modifications made to \scipion to facilitate its deployment and execution in centralized facilities.
+The section ends with a very brief description of \scipion design and architecture that highlights the most interesting features of this image processing framework.
 
@@
 This clear separation between data access and data processing marks an important difference between running a script or using \scipion, since it facilitates a form of computing in which several jobs are executed  concurrently, instead of sequentially.
-Moreover, \scipion workflow engine allows users to modify the workflow on the fly without modifying a single line of code.
+%Moreover, \scipion workflow engine allows users to modify the workflow on the fly without modifying a single line of code.
 %In practice, the latter follows  a more general and extensible approach that allows users to modify the workflow on the fly without modifying a single code line.
@@
 
-We would like to finish this section stressing that despite the modifications introduced to \scipion data and process model, the framework is backward compatible and keeps providing full data provenance.
-In other words, \scipion provides detailed information about the origin of the data  and the process to generate it.
+%It is worth mentioning that despite the modifications introduced to \scipion data and process model, the framework is backward compatible and keeps providing full data provenance.% In other words, \scipion provides detailed information about the origin of the data  and the process to generate it.
+
+\subsection{\scipion Design and Architecture}
+
+We would like to finish this section commenting on some of \scipion architecture design decisions that differentiate \scipion from another image processing frameworks.
+At the software architectural level,  the greatest difference
+between \scipion and the other EM image processing packages mentioned in the introduction 
+is that  \scipion pays special attention to define abstraction layers to simplify maintenance
+and extensibility, and that follows a strict workflow approach. A few examples will illustrate the main advantages of this design.
+
+To create a new protocol in Scipion, only a Python script needs to be developed. From this script, the system will discover the new protocol and will automatically build
+a form and store the protocol related information in a database.
+\scipion protocol developers must certainly know \scipion conventions
+but they do not need to modify or even understand the existing code.
+In  \scipion next version, the abstraction will be increased
+since protocols will be implemented as plug-ins,
+that is, independent small python scripts  that add functionality to the main
+\scipion application. The plug-in approach allows parallel development. That is,  
+since protocols are implemented independently from the main application
+they can be developed and released in parallel by different teams. In this way, support for a new version of \textit{Relion} does not require a new \scipion release but
+a release of the independent \textit{Relion} plug-in
+
+\scipion follows a pure workflow approach and therefore, implements pipelines and describes them as a set of interconnected steps (or protocols).
+In a workflow approach, users (including microscope staff) can reasonably be expected to modify the paths and incorporate new steps using a GUI.
+On the contrary, software packages that do not make this separation produce workflows that, in general, are more challenging to create or modify for non programmers.
+
+Another advantage of a clear separation between steps and relationships 
+is the improvement of load valance, reproducibility and  provenance (process of tracing and recording the origins of data and its movement between protocols).
+The first step in a \scipion workflow execution
+is to store in a database each of the protocols to be executed as well as how 
+outputs of one protocol are attached to the inputs of another protocol.
+Then, a workflow engine analyzes the stored data and using as constrain the number of CPUs assigned by the user to the workflow, optimizes the execution, that is, try to execute as many protocols in parallel as possible.
+Obviously, the data stored in the database  have enough information to execute each step and therefore guaranties reproducibility and provenance.
 

### 2018-09-25 Review/3.workflow_cnb.tex
to home

@@
   \caption{Graphical representation of the data flow through the CNB facility.
-Green, blue and red lines represent the microscope private network, the \cnb local area network and the public network respectively.}
+Green, blue and red lines represent the microscope private network, the \cnb local area network and the public (WAN) network respectively.}
   \label{fig:cnbpipeline}
@@
 The microscope control system has a 70 TB NAS (Network-attached storage) unit for saving movies as they are obtained (we will refer to this NAS as \mnas).
-This storage unit is shared through a Samba share and connected to the private network using a 10Gb connexion.
+This storage unit is shared through a Samba share and connected to the microscope private network using a 10Gb connexion.
 \scipion is executed on a Linux server which has two 1Gb NICs (network interface controllers) that connect it to the private network and the LAN (local area network) respectively (we will refer to this machine as \scipionbox).
@@
 
-The data flow is as follows: (a) the microscope system stores movies in the \mnas, (b) when executing movie-alignment protocols \scipionbox reads the movies from the \mnas but does not store them locally, (c) \onas has access to \mnas and \scipionbox hard disks (d) \onas has a collection of scripts than may be used to incrementally backup \scipion project and the movies to a local or remote disk, or serve this information through the Internet and, (e) in parallel, periodically \scipionbox transfers to  \hserver reports that may be checked by users or staff with an HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox/lastHTMLReport/}).
+The data flow is as follows: (a) the microscope system stores movies in the \mnas, (b) when executing movie-alignment protocols \scipionbox reads the movies from the \mnas but does not store them locally, (c) \onas has access to \mnas and \scipionbox hard disks (d) \onas has a collection of scripts than may be used to incrementally backup \scipion projects and the associated movies to a local or remote disk, or serve this information through the Internet, and (e) in parallel, periodically \scipionbox transfers to  \hserver reports that may be checked by users or staff with an HTML browser (see a report example at \url{http://nolan.cnb.csic.es/scipionbox/lastHTMLReport/}).
 

### 2018-09-25 Review/3.workflow_diamond.tex
to home

@@
 
-At present, most of developments described in this subsection are available in a developer's branch  at \url{https://github.com/I2PC/scipion/tree/release-1.1-headless-devel}.
+At present, most of the developments described in this subsection are available in a developer's branch  at \url{https://github.com/I2PC/scipion/tree/release-1.1-headless-devel}.
 Some of them (e.g.
 the monitor protocol) will be integrated in the main \scipion repository while the more Diamond specific software will be published in an independent repository.
-[** Ask Pablo about headless]
+\scipion headless mode is already integrated in the main \scipion branch.

### 2018-09-25 Review/4.future.tex
to home

@@
-\section{Future Developments and Usage}
+\section{Future Developments and Usage Statistics}
 
@@
 
-\subsection{Usage}
+\subsection{Usage Statistics}
 In the context of free and open source projects it is difficult to create usage statistics that are accurate. But certainly, if we want \scipion to succeed, we need to 
@@
 Using the HTTP header it is possible to guess the country in which \scipion has been installed.
-With these information we have created the usage map and table available at URL \url{http://scipion.i2pc.es/report_protocols/scipionUsage/} and  \url{http://scipion.i2pc.es/report_protocols/protocolTable/}.
-The data corresponds on ``global'' \scipion usage rather than ``facility'' usage but there
+With these information we have created the usage map and table available at URLs \url{http://scipion.i2pc.es/report_protocols/scipionUsage/} and  \url{http://scipion.i2pc.es/report_protocols/protocolTable/}.
+The data shows ``global'' \scipion usage rather than ``facility'' usage but there
 is a one to one relationship between the  countries with highest number of \scipion projects and the countries with \scipion based EM facilities.
@@
 Note that, in a effort to report real use and skip test projects, the number of projects in the map refers to the number of non empty \scipion projects updated more than once. That is, 
-\scipion projects that have been opened in at least two different days. Another constraint is that developers computers are filtered out.
+\scipion projects that have been opened in at least two different days. Another constraint is that developers' computers are filtered out.

### 2018-09-26 Review/answer_referee.tex
to class

@@
 \def\emadmin{\textit{EMAdmin}\xspace}
+\def\python{\textit{Python}\xspace}
 
@@
 Q \the\numexpr\value{pointcounter}-1\relax.2: Page 16: ``The sentence -They can potentially play with the pipeline in ways that are not possible when having a programmer writing the code for them.- Is not understandable on first reading.''.
-We have rewriting the paragraph as:
+We have rewritten the paragraph as:
 
@@
 
-\reply Since there is already a published manuscript describing \scipion, in the present manuscript  we decided to focus on \scipion features that have been specifically developed for working at EM facilities (e.g.
+\reply %Since there is already a published manuscript describing \scipion, in the present manuscript  we decided to focus on \scipion features that have been specifically developed for working at EM facilities (e.g.
 streaming).
-The main reasons why \scipion is better suited than other packages for EM work is rooted in \scipion main design rather than the new facilities related features.
-Nevertheless, the reviewer is right in pointing out than a ``A clear differentiation between Scipion and these software packages was lacking in this manuscript''.
+The main reasons why \scipion is better suited than other packages for EM work is rooted in \scipion main design, complimented with the new facilities related features.
+Nevertheless,
+The reviewer is right in pointing out than a ``A clear differentiation between \scipion and these software packages was lacking in this manuscript''.
 Therefore, we have decided to add the following subsection to the manuscript
@@
 
-We would like to finish this section commenting on some of \scipion architecture design decisions that differentiate \scipion from another image processing frameworks.
+We would like to finish this section commenting on some of \scipion  design decisions that differentiate \scipion from other image processing frameworks. \scipion
+is fully oriented to integrate existing
+EM software and therefore the versatility and number of possible steps
+combination is huge. Additionally, \scipion takes care of keeping a 
+rigorous traceability and reproducibility as well as to provide an API (application programming interface) to allow the integration  with already existing LIMS systems.
 At the software architectural level,  the greatest difference
@@
 is that  \scipion pays special attention to define abstraction layers to simplify maintenance
-and extensibility, and that follows a strict workflow approach. A few examples will illustrate the main advantages of this design.
+and extensibility and follows a strict workflow approach. A few examples will illustrate the main advantages of this approach.
 
-To create a new protocol in Scipion, only a Python script needs to be developed. From this script, the system will discover the new protocol and will automatically build
+To create a new protocol in \scipion, a single \python file needs to be created. From this file, the system will discover the new protocol and will automatically build
 a form and store the protocol related information in a database.
@@
 since protocols will be implemented as plug-ins,
-that is, independent small python scripts  that add functionality to the main
+that is, independent small \python files  that add functionality to the main
 \scipion application. The plug-in approach allows parallel development. That is,  
@@
 \scipion follows a pure workflow approach and therefore, implements pipelines and describes them as a set of interconnected steps (or protocols).
-In a workflow approach, users (including microscope staff) can reasonably be expected to modify the paths and incorporate new steps using a GUI.
+In a workflow approach, users (including microscope staff) can reasonably be expected to modify the paths and incorporate new steps using a GUI (graphic user interface).
 On the contrary, software packages that do not make this separation produce workflows that, in general, are more challenging to create or modify for non programmers.
@@
 Another advantage of a clear separation between steps and relationships 
-is the improvement of load valance, reproducibility and  provenance (process of tracing and recording the origins of data and its movement between protocols).
+is the improvement of load balance, reproducibility and  provenance (process of tracing and recording the origins of data and its movement between protocols).
 The first step in a \scipion workflow execution
@@
 outputs of one protocol are attached to the inputs of another protocol.
-Then, a workflow engine analyzes the stored data and using as constrain the number of CPUs assigned by the user to the workflow, optimizes the execution, that is, try to execute as many protocols in parallel as possible.
-Obviously, the data stored in the database  have enough information to execute each step and therefore guaranties reproducibility and provenance.
+Then, a workflow engine analyzes the stored data and, using as constraint the number of CPUs assigned by the user to the workflow, optimizes the execution, that is, tries to execute as many protocols in parallel as possible.
+Obviously, the data stored in the database  have enough information to execute each step and therefore guarantee reproducibility and provenance.
+Additionally, \scipion currently exports a JSON (JavaScript Object Notation) file describing the complete preprocessing workflow used, so that it can be fully reproduced and visualized elsewhere.
 
@@
 Note that, in a effort to report real use and skip test projects, the number of projects in the map refers to the number of non empty \scipion projects updated more than once. That is, 
-\scipion projects that have been opened in at least two different days. Another constraint is that developers computers are filtered out.
+\scipion projects that have been opened in at least two different days. Another constraint is that developers' computers are filtered out.
 \end{quote}

### 2018-09-26 Review/1.introduction.tex
to class

@@
 Exceptions are UCSFImage4 \citep{Li2015}, Focus \citep{Biyani2017} and Relion \citep{fernandez2017:relion-pipeline}.
-As we will show in this article, \scipion differs from these solutions in a larger flexibility, both from the point of view of users and facility staff.
+As we will show in this article, \scipion differs from these solutions in its larger flexibility, both from the point of view of users and facility staff.
 

### 2018-09-26 Review/2.overall.tex
to class

@@
 outputs of one protocol are attached to the inputs of another protocol.
-Then, a workflow engine analyzes the stored data and using as constrain the number of CPUs assigned by the user to the workflow, optimizes the execution, that is, try to execute as many protocols in parallel as possible.
-Obviously, the data stored in the database  have enough information to execute each step and therefore guaranties reproducibility and provenance.
-
-%for any object created by a  protocol, \scipion always annotate how this object has been generated by detailing: the steps, input data and parameters used in the computation.
-Data provenance guaranties reproducibility and traceability.
-%Since provenance is solved by Scipion and not Scipionbox we will not discuss here how it is been achieved.
-We just note that, before choosing which data has to be stored, it is necessary to define how these data have to be structured so that they can be later recovered and understood .
-%This abstract modeling of the data also simplifies the software interoperability offered by Scipion.
+Then, a workflow engine analyzes the stored data and, using as constraint the number of CPUs assigned by the user to the workflow, optimizes the execution, that is, tries to execute as many protocols in parallel as possible.
+Obviously, the data stored in the database  have enough information to execute each step and therefore guarantee reproducibility and provenance.
+Additionally, \scipion currently exports a JSON (JavaScript Object Notation) file describing the complete preprocessing workflow used, so that it can be fully reproduced and visualized elsewhere.
 
@@
 The second protocol performs direct detector movie alignment.
-Currently, \scipion offers  five  methods for movie alignment: motioncorr \citep{Li2015}, motioncor2 \citep{Zheng2016:motioncor2}, correlation (an unpublished CPU version that implements in Xmipp a global alignment method and solves it similarly to motioncorr but using robust statistics), unblur \citep{Grant2015:unblur} and optical flow \citep{Abrishami2015:optical-flow}.
+Currently, \scipion offers  five  methods for movie alignment: motioncorr \citep{Li2015}, motioncor2 \citep{Zheng2016:motioncor2}, correlation (an unpublished CPU version that implements in Xmipp a global alignment method and solves it similarly to motioncorr but using more robust statistics), unblur \citep{Grant2015:unblur} and optical flow \citep{Abrishami2015:optical-flow}.
 Then the CTFs of the resulting micrographs are calculated.
-Scipion provides three CTF estimation methods:
-CTFFIND4 (Rohou and Grigorieff, 2015), gCTF (Zhang, 2016) and Xmipp CTF estimation (Sorzano et al., 2007; Vargas et al., 2013).
+\scipion provides three CTF estimation methods:
+CTFFIND4 \citep{Rohou2015}, gCTF \citep{Zhang2016} and Xmipp CTF estimation \citep{sorzano2007, Vargas2013}.
 At this point the user may compute the CTFs using several algorithms and compare them with the protocol \protocol{CTF consensus} that reports up to which frequency two estimations of the same micrograph are equivalent.

### 2018-09-26 Review/3.workflow.tex
to class

@@
 After this first installation, \scipion was deployed on SciLifeLab (a centralized facility located at Stockholm), where a rigorous booking system is followed.
-There, middleware (software that acts as a bridge between applications) was developed to interface with this booking platform to produce better reports.
+There, middleware (software that acts as a bridge between applications) was developed to interface with this booking platform to produce more precise reports.
 Lately,  \scipion has been installed in  large synchrotrons  such as ESRF and Diamond,  where projects are handled by ISPyB (a customized laboratory information management system -LIMS- \citep{Delageniere2011:ispb}) and data are saved on a distributed file system.

### 2018-09-26 Review/3.workflow_diamond.tex
to class

@@
-\subsection{Case 3: Scipion at eBIC (Diamond Light Source)}
+\subsection{Case 3: \scipion at eBIC (Diamond Light Source)}
 
@@
 
-\subsection{Additional Software Developments: ISPyB monitor, Scipion Headless}
+\subsection{Additional Software Developments: ISPyB monitor, \scipion Headless}
 
@@
 processing, in Diamond, the electronic notebook ISPyB is the natural framework
-to record and show metadata about the different experiments. A new monitor has thus been developed to allow Scipion to directly interface with ISPyB using the ISPyB python API.
+to record and show metadata about the different experiments. A new monitor has thus been developed to allow \scipion to directly interface with ISPyB using the ISPyB python API.
 %To monitor the progress of the automatic image processing, CNB and \scilifelab cryo-EM services make use of \scipion default monitor system.

### 2018-09-26 Review/3.workflow_scilifelab.tex
to home
to class

@@
-\subsection{Case 2: Scipion at the Swedish National Cryo-EM Facility}
+\subsection{Case 2: \scipion at the Swedish National Cryo-EM Facility}
 
@@
 In the following sections we will briefly describe the computational infrastructure at the \scilifelab node and the implementation of a Session Wizard to automate the initial setup while fetching information from other sources external to \scipion.
+%(In this contest wizard refers to a helper script which guide users through a certain process.)
 %[** thalos falcon III, titan k2 and falcon III]
@@
 \subsubsection{Additional Software Developments: \emph{Session Wizard}}
-In order to interface \scipion with the microscope infrastructure an upgraded version of the original script developed at CNB was used.
+In order to interface \scipion with the microscope infrastructure, an upgraded version of the original script developed at CNB was used.
 Unfortunately, it was difficult to reuse the existing code due to a different folder structure and, more important, a different approach to user and microscope management.
 To address the new set of requirements, a new wizard has been development at \scilifelab.
-(In this contest wizard refers to a helper script which guide users through a certain process.) %For this new wizard, a basic data model was designed representing the current processing and data organization at \scilifelab.
+%(In this contest wizard refers to a helper script which guide users through a certain process.) %For this new wizard, a basic data model was designed representing the current processing and data organization at \scilifelab.
 %The model is composed by the following entities:

### 2018-09-26 Review/4.future.tex
to class

@@
 
-As part of the EOSCPilot project (\url{https://eoscpilot.eu/cryoem}), we are also improving  how \scipion exports workflows in order to ease reporting of the work done.
+As part of the EOSCPilot project (\url{https://eoscpilot.eu/cryoem}), we have improved 
+how \scipion exports workflows in order to ease reporting of the work done.
+In this way we want to make  the data produced by \scipion more compliant with the FAIR guidelines given by the \citet{eu2016:fair}.
+This workflow file could go with the raw data acquired by CryoEM facilities and deposited at common EM databases such as EMPIAR or EMDB \citep{Patwardhan2016:databasesEM}.
+%As part of the EOSCPilot project (\url{https://eoscpilot.eu/cryoem}), we are also improving  how \scipion exports workflows in order to ease reporting of the work done.
 The goal is to write a workflow file that fully describes the image processing steps, so that if the same input data is provided to the workflow the same results should be obtained.
 In this way we want to make  the data produced by \scipion more compliant with FAIR guidelines given by the \citet{eu2016:fair}.
-This workflow file could go with the raw data acquired by CryoEM facilities and deposited at common EM databases such as EMPIAR or EMDB \citep{Patwardhan2016:databasesEM}.
 To facilitate the visualization of the workflow file in any of these databases, we are  developing a  webcomponent (\scipion \emph{workflow viewer}) that will easily allow these repositories to visualize the workflow on their web pages (\url{https://github.com/I2PC/web-workflow-viewer}).
@@
 This repository will contain ready-to-use workflows that support a range of use cases.
+A preliminary neta version of the workflows repository is available at \url{http://workflows.scipion.i2pc.es/}
 

### 2018-09-26 Review/6.acknowledgments.tex
to class

--- 2018-04-11 6.acknowledgments.tex
@@
 
-The Spanish Ministry of Economy and Competitiveness through Grants BIO2016-76400-R(AEI/FEDER, UE) and AEI/FEDER BFU 2016 74868P, the Comunidad Aut\'{o}noma de Madrid through Grant: S2017/BMD-3817, European Union (EU) and Horizon 2020 through grant iNEXT (INFRADEV-1-2014-1, Proposal: 654248).
+The Spanish Ministry of Economy and Competitiveness through Grants BIO2016-76400-R(AEI/FEDER, UE) and AEI/FEDER BFU 2016 74868P, the Comunidad Aut\'{o}noma de Madrid through Grant: S2017/BMD-3817, European Union (EU) and Horizon 2020 through grant Corbel (INFRADEV-1-2014-1, Proposal: 654248).
 The ``Knut \& Alice Wallenberg Foundation'', and ``A Pilot Facility development grant from Science for Life Laboratory''.
+European Union (EU) and Horizon 2020 through grant EOSCpilot (INFRADEV-04-2016, Proposal: 739563).
 

### 2018-09-27 Review/answer_referee.tex
to home

@@
 
-Another advantage of a clear separation between steps and relationships 
-is the improvement of load balance, reproducibility and  provenance (process of tracing and recording the origins of data and its movement between protocols).
+Another advantage of the workflow approach 
+is the improvement of balance load, reproducibility and  provenance (process of tracing and recording the origins of data and its movement between protocols).
 The first step in a \scipion workflow execution

### 2018-09-27 Review/2.overall.tex
to home

@@
 
-Another advantage of a clear separation between steps and relationships 
+Another advantage of the workflow approach 
 is the improvement of load balance, reproducibility and  provenance (process of tracing and recording the origins of data and its movement between protocols).

### 2018-09-27 Review/extract.sh

@@ -1,0 +1,11 @@
+#!/bin/bash
+export INFILE=scipionbox.pdf
+pages=("20")
+figures=("01")
+for ((i=0;i<${#pages[@]};++i)); do
+    pdftk $INFILE cat ${pages[i]} output figure_${figures[i]}.pdf
+done
+
+#RECOMPILE document with table caption OK.
+pdftk $INFILE cat 1-18 output manuscript.pdf
+pdftk $INFILE cat 19-19 output listFigures.pdf

### 2018-10-01 Review/answer_referee.tex
to class

@@
 
-\reply The following figure has been included in the manuscript.
+\reply The following two figures have been included in the manuscript.
 
@@
   \label{fig:cnbpipeline}
+
+\end{figure}
+\begin{figure}
+%\hspace*{-2cm}                                                           
+\includegraphics[width=1.0\textwidth]{images/scilifelab.png}
+  \caption{Overview of the Session Wizard  developed at \scilifelab.
+A) The user triggers the wizard for any of the current microscopes.
+B) The wizard retrieves project information from the \textit{Application Portal} and reservation details from the \textit{Booking System}.
+C) The wizard creates data folders, executes the image copying script, configures the Scipion streaming project and logs the current session into a database.
+D) The \scipion project is used to generate on the fly web reports and invoice documents later on.}
+  \label{fig:wizard}
 
@@
 
-\reply We think the reviewer refer here to section 4. As commented previously this section has been extensively rewritten to take into account the reviewers' suggestions and it includes a new figure.
+\reply We think the reviewer refer here to section 4.
+As commented previously this section has been extensively rewritten to take into account the reviewers' suggestions and it includes a two new figures.
 

### 2018-10-01 Review/3.workflow_scilifelab.tex
to class

@@
 has two nodes: at \scilifelab in Stockholm and at Ume\aa\ University. \scilifelab in Stockholm offers single-particle service with a 
-Talos Arctica for sample optimisation and a Titan Krios for high-resolution data collection.
-The Ume\aa\ node is expected to become operational during 2018, and will offer cryo-ET with a Titan Krios and a Scios DualBeam SEM.
+Talos Arctica for sample optimisation and a Titan Krios for high-resolution data collection. The Ume\aa\ node offers cryo-ET with a Titan Krios and a Scios DualBeam SEM.
 %The facilities can be accessed by Swedish researchers through 
@@
 Unfortunately, it was difficult to reuse the existing code due to a different folder structure and, more important, a different approach to user and microscope management.
-To address the new set of requirements, a new wizard has been development at \scilifelab.
+To address the new set of requirements, a new wizard has been developed at \scilifelab (see \ffigure{fig:wizard}).
 %(In this contest wizard refers to a helper script which guide users through a certain process.) %For this new wizard, a basic data model was designed representing the current processing and data organization at \scilifelab.
 %The model is composed by the following entities:
+\begin{figure}
+\hspace*{-2cm}                                                           
+\includegraphics[width=1.25\textwidth]{images/scilifelab.pdf}
+  \caption{Overview of the Session Wizard  developed at \scilifelab.
+A) The user triggers the wizard for any of the current microscopes.
+B) The wizard retrieves project information from the \textit{Application Portal} and reservation details from the \textit{Booking System}.
+C) The wizard creates data folders, executes the image copying script, configures the Scipion streaming project and logs the current session into a database.
+D) The \scipion project is used to generate on the fly web reports and invoice documents later on.}
+  \label{fig:wizard}
+
+\end{figure}
 

### 2018-10-01 Review/extract.sh

@@ -1,11 +1,14 @@
 #!/bin/bash
 export INFILE=scipionbox.pdf
 pages=("20")
-figures=("01")
+figures=("01" "02")
 for ((i=0;i<${#pages[@]};++i)); do
     pdftk $INFILE cat ${pages[i]} output figure_${figures[i]}.pdf
 done
 
 #RECOMPILE document with table caption OK.
-pdftk $INFILE cat 1-18 output manuscript.pdf
-pdftk $INFILE cat 19-19 output listFigures.pdf
+pdftk $INFILE cat 1-17 output manuscript_17.pdf
+pdftk $INFILE cat 1-18 output manuscript_18.pdf
+pdftk $INFILE cat 18-18 output listFigures.pdf
+#pdftk $INFILE cat 20-20 output Figure1.pdf
+#pdftk $INFILE cat 21-21 output Figure2.pdf

### 2018-10-01 Review/answer_referee.tex
to class

@@
 %\hspace*{-2cm}                                                           
-\includegraphics[width=1.0\textwidth]{images/scilifelab.png}
+\includegraphics[width=1.0\textwidth]{images/scilifelab.pdf}
   \caption{Overview of the Session Wizard  developed at \scilifelab.
@@
 \reply Section 4 has been extensively rewritten to take into account the reviewers' suggestions.
-In summary, we have included a new image, delete several paragraphs, modify others and move those paragraphs related with future work to a different section.
+In summary, we have included a two new images, delete several paragraphs, modify others and move those paragraphs related with future work to a different section.
 

### 2018-10-01 Review/extract.sh

@@ -2,13 +2,13 @@
 export INFILE=scipionbox.pdf
 pages=("20")
 figures=("01" "02")
-for ((i=0;i<${#pages[@]};++i)); do
-    pdftk $INFILE cat ${pages[i]} output figure_${figures[i]}.pdf
-done
+#for ((i=0;i<${#pages[@]};++i)); do
+#    pdftk $INFILE cat ${pages[i]} output figure_${figures[i]}.pdf
+#done
 
 #RECOMPILE document with table caption OK.
 pdftk $INFILE cat 1-17 output manuscript_17.pdf
 pdftk $INFILE cat 1-18 output manuscript_18.pdf
 pdftk $INFILE cat 18-18 output listFigures.pdf
-#pdftk $INFILE cat 20-20 output Figure1.pdf
-#pdftk $INFILE cat 21-21 output Figure2.pdf
+pdftk $INFILE cat 19-19 output Figure1.pdf
+pdftk $INFILE cat 20-20 output Figure2.pdf
